---
title: "FRI / BLT / FRZ mismatches between nominal and georeferenced IOTC data"
output: html_document
---------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(lubridate)
library(purrr)
```

## Context

This note focuses on the species-group inconsistencies involving `FRI`, `BLT`, and `FRZ` between the nominal IOTC dataset (`iotc_nom`) and the georeferenced IOTC dataset (`iotc`).

The issue is restricted to strata where:

* `fishing_fleet == "NEI"`
* `gear_type == "99.9"`

This means the problem is limited to very imprecise strata and does not affect the more detailed fleet/gear combinations.

## Diagnostic code

```{r}
prepare_species_strata <- function(data,
                                   dataset_name,
                                   grouping_keys = c("year", "fishing_fleet", "gear_type"),
                                   date_col = "time_start",
                                   species_col = "species") {
  
  data <- data %>%
    dplyr::ungroup() %>%
    dplyr::mutate(year = lubridate::year(.data[[date_col]]))
  
  missing_cols <- setdiff(c(grouping_keys, species_col), names(data))
  if (length(missing_cols) > 0) {
    stop("Missing columns in data: ", paste(missing_cols, collapse = ", "))
  }
  
  data %>%
    dplyr::select(dplyr::all_of(c(grouping_keys, species_col))) %>%
    dplyr::distinct() %>%
    dplyr::mutate(dataset = dataset_name)
}

inspect_species_group_mismatches <- function(nominal_data,
                                             georef_data,
                                             group_species = c("FRI", "BLT", "FRZ"),
                                             group_label = "FRZ_group",
                                             grouping_keys = c("year", "fishing_fleet", "gear_type"),
                                             date_col_nom = "time_start",
                                             date_col_geo = "time_start",
                                             species_col = "species") {
  
  nom <- prepare_species_strata(
    data = nominal_data,
    dataset_name = "nominal",
    grouping_keys = grouping_keys,
    date_col = date_col_nom,
    species_col = species_col
  ) %>%
    dplyr::filter(.data[[species_col]] %in% group_species)
  
  geo <- prepare_species_strata(
    data = georef_data,
    dataset_name = "georef",
    grouping_keys = grouping_keys,
    date_col = date_col_geo,
    species_col = species_col
  ) %>%
    dplyr::filter(.data[[species_col]] %in% group_species)
  
  join_keys <- c(grouping_keys, species_col)
  
  raw_compare <- dplyr::full_join(
    geo %>% dplyr::mutate(in_georef = TRUE),
    nom %>% dplyr::mutate(in_nominal = TRUE),
    by = join_keys
  ) %>%
    dplyr::mutate(
      in_georef = dplyr::coalesce(in_georef, FALSE),
      in_nominal = dplyr::coalesce(in_nominal, FALSE),
      match_status = dplyr::case_when(
        in_georef & in_nominal ~ "common",
        in_georef & !in_nominal ~ "only_georef",
        !in_georef & in_nominal ~ "only_nominal"
      )
    )
  
  mismatched_strata <- raw_compare %>%
    dplyr::filter(match_status != "common") %>%
    dplyr::distinct(dplyr::across(dplyr::all_of(grouping_keys)))
  
  if (nrow(mismatched_strata) == 0) {
    return(list(
      raw_compare = raw_compare,
      mismatched_species = raw_compare %>% dplyr::filter(match_status != "common"),
      grouped_diagnosis = tibble::tibble(),
      warnings = tibble::tibble()
    ))
  }
  
  grouped_diagnosis <- mismatched_strata %>%
    dplyr::left_join(
      geo %>%
        dplyr::group_by(dplyr::across(dplyr::all_of(grouping_keys))) %>%
        dplyr::summarise(
          species_georef = list(sort(unique(.data[[species_col]]))),
          n_georef_species = dplyr::n_distinct(.data[[species_col]]),
          .groups = "drop"
        ),
      by = grouping_keys
    ) %>%
    dplyr::left_join(
      nom %>%
        dplyr::group_by(dplyr::across(dplyr::all_of(grouping_keys))) %>%
        dplyr::summarise(
          species_nominal = list(sort(unique(.data[[species_col]]))),
          n_nominal_species = dplyr::n_distinct(.data[[species_col]]),
          .groups = "drop"
        ),
      by = grouping_keys
    ) %>%
    dplyr::mutate(
      species_georef = purrr::map(species_georef, ~ if (is.null(.x)) character(0) else .x),
      species_nominal = purrr::map(species_nominal, ~ if (is.null(.x)) character(0) else .x),
      common_species = purrr::map2(species_georef, species_nominal, intersect),
      only_georef_species = purrr::map2(species_georef, species_nominal, setdiff),
      only_nominal_species = purrr::map2(species_nominal, species_georef, setdiff),
      n_common_species = purrr::map_int(common_species, length),
      n_only_georef_species = purrr::map_int(only_georef_species, length),
      n_only_nominal_species = purrr::map_int(only_nominal_species, length),
      group_present_in_both =
        purrr::map_int(species_georef, length) > 0 &
        purrr::map_int(species_nominal, length) > 0,
      regrouping_possible =
        group_present_in_both &
        (n_only_georef_species > 0 | n_only_nominal_species > 0),
      partial_warning =
        n_common_species > 0 &
        (n_only_georef_species > 0 | n_only_nominal_species > 0),
      warning_message = dplyr::case_when(
        partial_warning ~ purrr::pmap_chr(
          list(common_species, only_georef_species, only_nominal_species),
          function(common_species, only_georef_species, only_nominal_species) {
            paste0(
              "Partial match inside ", group_label, ": common = {",
              paste(common_species, collapse = ", "),
              "}; only_georef = {",
              paste(only_georef_species, collapse = ", "),
              "}; only_nominal = {",
              paste(only_nominal_species, collapse = ", "),
              "}."
            )
          }
        ),
        regrouping_possible ~ "Mismatch may be resolved by grouping species.",
        TRUE ~ NA_character_
      )
    )
  
  warnings_tbl <- grouped_diagnosis %>%
    dplyr::filter(!is.na(warning_message))
  
  list(
    raw_compare = raw_compare,
    mismatched_species = raw_compare %>% dplyr::filter(match_status != "common"),
    grouped_diagnosis = grouped_diagnosis,
    warnings = warnings_tbl
  )
}
```

## Run the diagnostic

```{r}
res_frz <- inspect_species_group_mismatches(
  nominal_data = iotc_nom,
  georef_data = iotc,
  group_species = c("FRI", "BLT", "FRZ"),
  group_label = "FRI_BLT_FRZ",
  grouping_keys = c("year", "fishing_fleet", "gear_type"),
  date_col_nom = "time_start",
  date_col_geo = "time_start",
  species_col = "species"
)
```

## Relevant problematic strata

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible)
```

## Restriction to the problematic fleet and gear strata

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible) %>%
  dplyr::count(fishing_fleet, gear_type, sort = TRUE)
```

The output shows that these mismatches concern only:

* `fishing_fleet = "NEI"`
* `gear_type = "99.9"`

## Interpretation by period

### 1970 to 1983

For 1970 to 1983, the georeferenced dataset contains only `FRI`, whereas the nominal dataset contains `BLT` and `FRI`.

This suggests two possible interpretations:

1. only `BLT` is missing in the georeferenced dataset;
2. or the georeferenced `FRI` was in fact intended to represent the broader grouped category later represented by `FRZ`.

In practice, this discrepancy remains limited because the mismatch occurs only in `NEI` / `99.9` strata.

### 1984 to 2021

From 1984 to 2021, the georeferenced dataset contains `BLT`, `FRI`, and `FRZ`, while the nominal dataset does not contain `FRZ`.

This is not a major issue, but a methodological choice is needed:

* either aggregate everything into `FRZ`;
* or redistribute / raise `BLT`, `FRI`, and `FRZ` in the georeferenced data against the nominal sum of `BLT + FRI`.

The second option preserves more detail, but the first is simpler and may be more robust if the goal is only to ensure compatibility between datasets.

### 1992 to 1993 and 2007 to 2013

For 1992 to 1993 and 2007 to 2013, the situation is less clear.

In these years, the georeferenced dataset contains `FRI` and `FRZ`, whereas the nominal dataset contains only `BLT` and `FRI`.

The key question is whether georeferenced `FRZ` corresponds mainly to nominal `BLT`, or whether both datasets should instead be harmonised at the aggregated `FRZ` level.

A cautious solution would be to aggregate over the full group (`FRI`, `BLT`, `FRZ`) for these strata rather than trying to infer a one-to-one correspondence.

## Useful extracts

### Problematic strata only

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible) %>%
  dplyr::select(
    year, fishing_fleet, gear_type,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species,
    warning_message
  )
```

### Focus on 1970 to 1983

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible, dplyr::between(year, 1970, 1983)) %>%
  dplyr::select(
    year, fishing_fleet, gear_type,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species
  )
```

### Focus on 1984 to 2021

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible, dplyr::between(year, 1984, 2021)) %>%
  dplyr::select(
    year, fishing_fleet, gear_type,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species
  )
```

### Focus on 1992 to 1993 and 2007 to 2013

```{r}
res_frz$grouped_diagnosis %>%
  dplyr::filter(
    regrouping_possible,
    year %in% c(1992, 1993, 2007:2013)
  ) %>%
  dplyr::select(
    year, fishing_fleet, gear_type,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species,
    warning_message
  )
```

## Short conclusion

These `FRI` / `BLT` / `FRZ` inconsistencies are limited to the very coarse strata `fishing_fleet = NEI` and `gear_type = 99.9`.

They do not suggest a broad incompatibility between nominal and georeferenced datasets, but rather a coding difference within a small tuna group.

A pragmatic decision is still needed for harmonisation:

* aggregate all these strata to `FRZ`; or
* keep detail in the georeferenced data and compare / raise on the nominal total of `BLT + FRI`.

For the ambiguous years (1992-1993 and 2007-2013), aggregation at the group level is likely the safest approach.

## Same diagnostic for ICCAT

```{r}
iccat_nom <- nominal_catch_for_raising %>%
  dplyr::filter(source_authority == "ICCAT")

iccat <- georef_dataset %>%
  dplyr::filter(source_authority == "ICCAT")
```

### Diagnostic at `year x fishing_fleet x gear_type`

```{r}
res_frz_iccat_detail <- inspect_species_group_mismatches(
  nominal_data = iccat_nom,
  georef_data = iccat,
  group_species = c("FRI", "BLT", "FRZ"),
  group_label = "FRI_BLT_FRZ",
  grouping_keys = c("year", "fishing_fleet", "gear_type"),
  date_col_nom = "time_start",
  date_col_geo = "time_start",
  species_col = "species"
)
```

```{r}
res_frz_iccat_detail$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible)
```

For ICCAT, the detailed comparison suggests that some mismatches come from nominal strata being less precise than georeferenced strata rather than from a true contradiction between datasets.

For example, `FRI` is missing in nominal USA data in 1970, and this likely corresponds to a broader `NEI` category. This appears to happen in several years. In such cases, georeferenced data may simply be more precise than nominal data.

A specific point is that for `GG`, the issue seems to concern only `gear_type = 99.9`.

### Diagnostic at year level only

Because the `year x fishing_fleet x gear_type` comparison can produce many georeferenced-only strata with no nominal counterpart, it is useful to also inspect the same issue at the year level only.

```{r}
res_frz_iccat_year <- inspect_species_group_mismatches(
  nominal_data = iccat_nom,
  georef_data = iccat,
  group_species = c("FRI", "BLT", "FRZ"),
  group_label = "FRI_BLT_FRZ",
  grouping_keys = c("year"),
  date_col_nom = "time_start",
  date_col_geo = "time_start",
  species_col = "species"
)
```

```{r}
res_frz_iccat_year$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible)
```

### Interpretation for ICCAT

Before 1967, there are only nominal data, which is not problematic in itself.

After 1967, the main issue is that georeferenced data contain `FRI` while nominal data contain `BLT + FRI`.

This leaves an open question:

* either georeferenced data are missing `BLT`;
* or georeferenced `FRI` is in fact an aggregated category, but coded as `FRI` rather than `FRZ`.

For 2022 and 2023, nominal data do not contain `BLT` or `FRI`, which should be treated separately from the earlier pattern.

### Useful extracts for ICCAT

```{r}
res_frz_iccat_detail$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible) %>%
  dplyr::select(
    year, fishing_fleet, gear_type,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species,
    warning_message
  )
```

```{r}
res_frz_iccat_year$grouped_diagnosis %>%
  dplyr::filter(regrouping_possible) %>%
  dplyr::select(
    year,
    species_georef, species_nominal,
    common_species,
    only_georef_species,
    only_nominal_species,
    warning_message
  )
```
