---
title: "Harmonize IATTC PS Shark ByFlag Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'iattc', 'catch', 'data', 'PublicPSSharkFlag.csv')
```

Harmonize IATTC PS Shark ByFlag Catch Datasets

This function harmonizes the structure of IATTC PS (Purse Seine) shark catch datasets by flag,
preparing them for integration into the Tuna Atlas database. It ensures that only the essential
fields are retained and that metadata is included if the dataset is destined for the database.


@return None; this function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details The function processes the data based on flag categories specifically for shark species.
         It involves cleaning, restructuring, and harmonizing data fields to meet specified
         standards. Metadata integration is optional and contingent on the destination of the dataset.

@importFrom dplyr select mutate
@importFrom readr read_csv write_csv
@seealso \code{\link{FUN_catches_IATTC_CE_Flag_or_SetType_Shark}} for the specific processing of shark catches by flag,
         \code{\link{IATTC_CE_catches_pivotDSD_to_harmonizedDSD}} for general data structuring.
@export
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@keywords IATTC, tuna, fisheries, data harmonization, shark catch
'# This script works with any data that has the first 5 columns named and ordered as follow: {Year|Month|Flag|LatC1|LonC1|NumSets}
packages
Historical name for the dataset at source  PublicPSSharkFlag.csv

```{r}
opts <- options()
options(encoding = "UTF-8")
```

# Catches

```{r}
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/FUN_catches_IATTC_CE_Flag_or_SetType_Shark.R")
catches_pivot_IATTC <-FUN_catches_IATTC_CE_Flag_or_SetType_Shark(path_to_raw_dataset,"Flag","PS")
catches_pivot_IATTC$NumSets<-NULL

colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/IATTC_CE_catches_pivotDSD_to_harmonizedDSD.R")
catches<-IATTC_CE_catches_pivotDSD_to_harmonizedDSD(catches_pivot_IATTC,colToKeep_captures)

colnames(catches)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
catches$source_authority<-"IATTC"
catches$measurement_type <- "RC" # Retained catches
catches$measurement <- "catch"
```


```{r}
catches$time_start <- as.Date(catches$time_start)
catches$time_end <- as.Date(catches$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(catches$time_start), "%Y"), "-01-01"),
	paste0(format(max(catches$time_end), "%Y"), "-12-31"),
	sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
```

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

