# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is automatically created from the function: https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/refs/heads/master/R/tunaatlas_scripts/pre-harmonization/southern_hemisphere_oceans_effort_1deg_1m_tunaatlasccsbt_level0__surface.R, the documentation keeps the format of roxygen2 skeleton.

A summary of the mapping process is provided. The path to the dataset is specified. You will find on this same repository on GitHub the first line of each dataset.
The datasets are named after the historical name provided by tRFMOs while exporting and may change.
The information provided in the Rmd allows understanding correctly which dataset should be used in this markdown.

Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.

If you are interested in further details, the results and codes are available for review.

*Each `.Rmd` script requires the user to knit the dataset at the beginning of the script in order to execute the harmonization process correctly.
It is also possible to run the code chunk by chunk but be sure to be in the correct working directory (i.e., the one of the `.Rmd`).*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


````{r}
path_to_raw_dataset <- 
````

Harmonize CCSBT Surface Effort Datasets

Harmonizes the structure of CCSBT catch-and-effort datasets specifically for 'Surface' effort data.
This involves reading the raw dataset, performing various data transformations including renaming,
mutating, and selecting columns, and calculating new fields. The output is a harmonized dataset suitable
for integration into the Tuna Atlas database.


@details
The function performs the following key steps:
- Reads the raw dataset and metadata file from specified paths.
- Applies renaming and transformation logic to align with the database schema.
- Calculates conversion factors and standardizes time format.
- Generates harmonized dataset and metadata files, which are then saved to CSV.

The harmonization process is tailored to CCSBT 'Surface' datasets and assumes specific column names
date range and exports the harmonized data, metadata, and code lists as CSV files.


@examples
\dontrun{
}

@importFrom dplyr filter mutate group_by summarise
@importFrom readxl read_excel
@importFrom lubridate as_date ceiling_date
@export



````{r}
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_spatial_5.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_time_2.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/format_time_db_format.R")
  
  if(!(require(dplyr))){ 
    install.packages(dplyr) 
    (require(dplyr))} 
  
  
  if(!require(readxl)){
    install.packages("readxl")
    require(readxl)
  }
````

Input data sample (after importing as data.frame in R):
YEAR MONTH COUNTRY_CODE GEAR_CODE CCSBT_STATISTICAL_AREA LATITUDE LONGITUDE NUMBER_OF_HOURS_SEARCHED WEIGHT_(Kg)_OF_SBT_RETAINED
1975    12           AU        PS                      4      -37       150                       17                       59800
1976     8           AU        BB                      4      -36       150                        8                           0
1976     9           AU        BB                      4      -37       150                       14                       36200
1976     9           AU        BB                      4      -35       150                        4                           0
1976     9           AU        BB                      4      -34       151                       21                           0
1976    10           AU        BB                      4      -37       150                       51                       41000
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
AU   BB 1976-08-01 1976-09-01  6336150    ALL       HRSRH      8
AU   BB 1976-09-01 1976-10-01  6334151    ALL       HRSRH     21
AU   BB 1976-09-01 1976-10-01  6335150    ALL       HRSRH      4
AU   BB 1976-09-01 1976-10-01  6337150    ALL       HRSRH     14
AU   BB 1976-10-01 1976-11-01  6334151    ALL       HRSRH     21
AU   BB 1976-10-01 1976-11-01  6336150    ALL       HRSRH     29

````{r}
  #----------------------------------------------------------------------------------------------------------------------------
````

Historical name for the dataset at source  CEData_Surface.xlsx

````{r}
  opts <- options()
  options(encoding = "UTF-8")
  
  
  RFMO_CE<-readxl::read_excel(path_to_raw_dataset, sheet = "CEData_Surface", col_names = TRUE, col_types = NULL,na = "")
  
  colnames(RFMO_CE)<-gsub("\r\n", "_", colnames(RFMO_CE))
  colnames(RFMO_CE)<-gsub(" ", "_", colnames(RFMO_CE))
  colnames(RFMO_CE)<-gsub("\\(", "", colnames(RFMO_CE))
  colnames(RFMO_CE)<-gsub("\\)", "", colnames(RFMO_CE))
  RFMO_CE<-as.data.frame(RFMO_CE)
  
  #Remove lines that are read in the Excel but that are not real
  RFMO_CE<- RFMO_CE[!is.na(RFMO_CE$YEAR),] 
  RFMO_CE$WEIGHT_Kg_OF_SBT_RETAINED<-as.numeric(RFMO_CE$WEIGHT_Kg_OF_SBT_RETAINED)
  RFMO_CE$NUMBER_OF_HOURS_SEARCHED<-as.numeric(RFMO_CE$NUMBER_OF_HOURS_SEARCHED)
  
  #FishingFleet
  RFMO_CE$FishingFleet<-RFMO_CE$COUNTRY_CODE
  #Gear
  RFMO_CE$Gear<-RFMO_CE$GEAR_CODE
  
  # replace PS by Purse Seine and BB by Pole and Line
  
  RFMO_CE$Gear[RFMO_CE$Gear == "PS"] <- "Purse Seine"
  RFMO_CE$Gear[RFMO_CE$Gear == "BB"] <- "Pole and Line"
  
  #Year and period
  RFMO_CE<-harmo_time_2(RFMO_CE, "YEAR", "MONTH")
  #Format inputDataset time to have the time format of the DB, which is one column time_start and one time_end
  RFMO_CE<-format_time_db_format(RFMO_CE)
  
  # Area 
  RFMO_CE<-harmo_spatial_5(RFMO_CE,"LATITUDE","LONGITUDE",1,5)
  
  #School
  RFMO_CE$School<-"UNK"
  
  #Species
  RFMO_CE$Species<-"SBF"
  
  #CatchType
  RFMO_CE$CatchType<-"UNK" #not used later as it is no catch
  
  
  efforts<-RFMO_CE
  
  efforts$EffortUnits<-"NUMBER_OF_HOURS_SEARCHED"
  efforts$Effort<-efforts$NUMBER_OF_HOURS_SEARCHED
  colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
  efforts <-efforts[colToKeep_efforts]
  
  
  #remove whitespaces on columns that should not have withespace
  
  efforts[,c("AreaName","FishingFleet")]<-as.data.frame(apply(efforts[,c("AreaName","FishingFleet")],2,function(x){gsub(" *$","",x)}),stringsAsFactors=FALSE)
  
  # remove 0 and NA values 
  
  efforts <- efforts  %>% 
    filter( ! Effort %in% 0 ) %>%
    filter( ! is.na(Effort)) 
  
  efforts <- efforts %>% 
    dplyr::group_by(FishingFleet,Gear,time_start,time_end,AreaName,School,EffortUnits) %>% 
    dplyr::summarise(Effort = sum(Effort))  
  efforts<-as.data.frame(efforts)
  
  
  colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
  efforts$source_authority<-"CCSBT"
  efforts$measurement <- "effort" 
  
  #----------------------------------------------------------------------------------------------------------------------------
  efforts$time_start <- as.Date(efforts$time_start)
  efforts$time_end <- as.Date(efforts$time_end)
  dataset_temporal_extent <- paste(
    paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
    paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
    sep = "/"
  )
  efforts$measurement_processing_level <- "unknown" 
  base1 <- tools::file_path_sans_ext(basename(filename1))
  # sorties same folder as path_to_raw_dataset 
  output_name_dataset <- "Dataset_harmonized.csv"
  
  write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
  
  #----------------------------------------------------------------------------------------------------------------------------
````

@ Load pre-harmonization scripts and apply mappings

````{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
````

@ Handle unmapped values and save the results

````{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
````

Display the first few rows of the mapping summaries

````{r}
print(head(mapping_codelist$recap_mapping))
````

