---
title: "Harmonize Data Structure of ICCAT Effort Dataset"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'iccat', 'effort', 'data', 't2ce_noSchool.csv')
```

Harmonize Data Structure of ICCAT Effort Dataset

This function harmonizes the data structure of ICCAT effort datasets from voluminous Microsoft Access databases.
It handles datasets by potentially including metadata and code lists for integration within the Tuna Atlas database.
Depending on the parameter settings, either 'fleet' or 'flag' column will be kept in the output.

@param keep_fleet_instead_of_flag Boolean parameter to decide whether to keep 'fleet' instead of 'flag'.
@return Writes a harmonized dataset, metadata file, and code lists to CSV files.
@export

@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}

@keywords International Commission for the Conservation of Atlantic Tunas (ICCAT), RFMO, Sardara Global Database

@seealso \code{\link{convertDSD_iccat_ce_task2_ByOperationMode}}, \code{\link{convertDSD_iccat_nc}} for other ICCAT data structure conversions.

@examples



```{r}
keep_fleet_instead_of_flag=FALSE
```

packages

```{r}
if(!require(readr)){
  install.packages("readr")
  require(readr)
}
if(!require(dplyr)){
  install.packages("dplyr")
  require(dplyr)
}
```

Input data sample: No sample. Miscrosoft Acces DB. However after the commands that read the input DB the sample is the following:
StrataID DSetID FleetID GearGrpCode GearCode FileTypeCode YearC TimePeriodID SquareTypeCode QuadID Lat Lon Eff1 Eff1Type Eff2 Eff2Type DSetTypeID CatchUnit ALB BET     BFT BUM
1      1 021ES00          TP     TRAP       OF-REP  1950           17            1x1      4  36   5    4 NO.TRAPS   NA                  nw        kg   0   0 6725000   0
2      1 021ES00          TP     TRAP       OF-REP  1950           17            1x1      4  36   5    4 NO.TRAPS   NA                  nw        nr   0   0   52928   0
3      2 026YU00          PS       PS       OF-REP  1950           17            5x5      1  40  15   14 NO.BOATS   NA                  -w        kg   0   0  657000   0
4      3 021ES00          TP     TRAP       OF-REP  1951           17            1x1      4  36   5    4 NO.TRAPS   NA                  nw        kg   0   0 3072000   0
5      3 021ES00          TP     TRAP       OF-REP  1951           17            1x1      4  36   5    4 NO.TRAPS   NA                  nw        nr   0   0   28654   0
6      4 026YU00          PS       PS       OF-REP  1951           17            5x5      1  40  15   14 NO.BOATS   NA                  -w        kg   0   0  531000   0
SAI SKJ SWO WHM YFT BLF BLT BON BOP BRS CER FRI KGM KGX LTA MAW SLT SSM WAH oSmt BIL BLM MLS SBF SPF oTun BSH POR SMA MAK oSks FleetCode       FleetName FlagID FlagCode
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0    EU.ESP       EU.España     21   EU.ESP
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0    EU.ESP       EU.España     21   EU.ESP
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0       YUG Yugoslavia Fed.     26      YUG
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0    EU.ESP       EU.España     21   EU.ESP
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0    EU.ESP       EU.España     21   EU.ESP
0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0    0   0   0   0   0   0    0   0   0   0   0    0       YUG Yugoslavia Fed.     26      YUG
FlagName StatusCode
EU.España         CP
EU.España         CP
Yugoslavia Fed.        NCO
EU.España         CP
EU.España         CP
Yugoslavia Fed.        NCO
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
ARG   LL 1967-02-01 1967-03-01  7330040    ALL    NO.HOOKS  13000
ARG   LL 1967-03-01 1967-04-01  7330040    ALL    NO.HOOKS  67000
ARG   LL 1967-04-01 1967-05-01  7330040    ALL    NO.HOOKS 107000
ARG   LL 1967-05-01 1967-06-01  7330040    ALL    NO.HOOKS  88000
ARG   LL 1967-06-01 1967-07-01  7330040    ALL    NO.HOOKS  66000
ARG   LL 1967-07-01 1967-08-01  7330040    ALL    NO.HOOKS  50000
# download database
Historical name for the dataset at source  t2ce_noSchool.csv
----------------------------------------------------------------------------------------------------------------------------

```{r}
keep_fleet_instead_of_flag=FALSE  

t2ce <- as.data.frame(readr::read_csv(path_to_raw_dataset))
ICCAT_CE_species_colnames<-setdiff(colnames(t2ce),c("StrataID","DSetID","FleetID","GearGrpCode","GearCode","FileTypeCode","YearC","TimePeriodID","SquareTypeCode","QuadID","Lat","Lon","Eff1","Eff1Type","Eff2","Eff2Type","DSetTypeID","CatchUnit", "FleetCode", "FleetName", "FlagID", "FlagCode"))
```

Flags<-mdb.get(paste0(working_directory_init,"/db.mdb"),tables='Flags',stringsAsFactors=FALSE,strip.white=TRUE)
data_pivot_ICCAT<-left_join(t2ce,Flags,by="FleetID")  # equivalent to "select FlagCode,FlagID,t2ce.* from t2ce, Flags where t2ce.FleetID=Flags.FleetID"
# Efforts
# Reach the efforts pivot DSD using a function in ICCAT_functions.R
# If we want in the output dataset the column 'FleetCode' instead of 'flag'

```{r}
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/FUN_efforts_ICCAT_CE_without_schooltype.R")
efforts_pivot_ICCAT<-FUN_efforts_ICCAT_CE_without_schooltype(RFMO_CE = t2ce,ICCAT_CE_species_colnames)

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/FUN_efforts_ICCAT_CE_keep_all_efforts.R")
efforts_pivot_ICCAT<-FUN_efforts_ICCAT_CE_keep_all_efforts(efforts_pivot_ICCAT,c("Eff1","Eff2"),c("Eff1Type","Eff2Type"))
```

School

```{r}
efforts_pivot_ICCAT$School<-"ALL"
```

Flag

```{r}
efforts_pivot_ICCAT$Flag<-efforts_pivot_ICCAT$FlagCode

names(efforts_pivot_ICCAT)[names(efforts_pivot_ICCAT) == 'FleetCode'] <- 'FishingFleet'
```

Reach the efforts harmonized DSD using a function in ICCAT_functions.R

```{r}
colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/ICCAT_CE_effort_pivotDSD_to_harmonizedDSD.R")
efforts<-ICCAT_CE_effort_pivotDSD_to_harmonizedDSD(efforts_pivot_ICCAT,colToKeep_efforts)
efforts$CatchType <- "C" #bastien adding as it is not in effort function but it is in chatch function
colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value","measurement_type")
efforts$source_authority<-"ICCAT"
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)


output_name_dataset <- "Dataset_harmonized.csv"
write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
```

----------------------------------------------------------------------------------------------------------------------------
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

