---
title: "Harmonize data structure of ICCAT by operation mode effort datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is automatically created from the function: https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/refs/heads/feature/update-ICCAT-efforts-dataset/no-ref/R/tunaatlas_scripts/pre-harmonization/atlantic_ocean_effort_1deg_1m_ps_tunaatlasiccat_level0__byschool.R, the documentation keeps the format of roxygen2 skeleton.

A summary of the mapping process is provided. The path to the dataset is specified. You will find on this same repository on GitHub the first line of each dataset.
The datasets are named after the historical name provided by tRFMOs while exporting and may change.
The information provided in the Rmd allows understanding correctly which dataset should be used in this markdown.

Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.

If you are interested in further details, the results and codes are available for review.

*Each `.Rmd` script requires the user to knit the dataset at the beginning of the script in order to execute the harmonization process correctly.
It is also possible to run the code chunk by chunk but be sure to be in the correct working directory (i.e., the one of the `.Rmd`).*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


````{r}
path_to_raw_dataset <- here::here('R/tunaatlas_scripts/pre-harmonization', 'iccat', 'effort', 'data', 't2ce_ETRO-PS1991-2024_bySchool_v1.xlsx')
````

Harmonize data structure of ICCAT by operation mode effort datasets

This function harmonizes the structure of ICCAT catch-and-effort datasets
based on operation modes. It adapts the dataset structure for compatibility with
the Tuna Atlas database, considering only specific fields as mandatory. The
function also handles optional metadata integration if provided.

@param keep_fleet_instead_of_flag Logical, defaults to FALSE. Determines whether to
       replace the 'flag' column with the 'fleet' column in the output dataset.

@return None; this function performs data manipulation and outputs files directly.

@import dplyr
@importFrom stringr str_detect str_replace
@importFrom readr read_csv write_csv
@seealso \code{\link{FUN_efforts_ICCAT_CE_keep_all_efforts}} for converting ICCAT task 2,
@export
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
@keywords ICCAT, tuna, fisheries, data harmonization

````{r}
keep_fleet_instead_of_flag=FALSE
````

packages

````{r}
if(!require(dplyr)){
  install.packages("dplyr")
  require(dplyr)
}
````

Historical name for the dataset at source  t2ce_bySchool.csv

````{r}
opts <- options()
options(encoding = "UTF-8")
````

Input data sample (after importing as data.frame in R): 
A tibble: 6 × 33
DSetID StrataID FlagName  FleetCode      GearCode YearC Decade TimePeriodID GeoStrataCode QuadID   Lat   Lon  xLon  yLat FishMode  Eff1 Eff1Type   Eff2 Eff2Type   Eff3 Eff3Type   Eff4 Eff4Type   Eff5 Eff5Type    YFT   ALB   BET   BLF   LTA   SKJ   FRI TOTAL
<dbl>    <dbl> <chr>     <chr>          <chr>    <dbl>  <dbl>        <dbl> <chr>          <dbl> <dbl> <dbl> <dbl> <dbl> <chr>    <dbl> <chr>     <dbl> <chr>     <dbl> <chr>     <dbl> <chr>     <dbl> <chr>     <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
1   2797   509067 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                1     0     1   1.5   0.5 FAD       13   FISH.HOUR  25.9 HOURS.SEA  13   Hours.STD     0 Hours.FAD  0    Hours.FSC     0     0     0     0     0     0     0     0
2   2797   509068 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                1     1     0   0.5   1.5 FAD       13   FISH.HOUR  25.9 HOURS.SEA  13   Hours.STD     0 Hours.FAD  0    Hours.FSC     0     0     0     0     0     0     0     0
3   2797   509069 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                1     1     2   2.5   1.5 FAD       13   FISH.HOUR  25.9 HOURS.SEA  13   Hours.STD     0 Hours.FAD  0    Hours.FSC     0     0     0     0     0     0     0     0
4   2797   509070 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                2     0     0   0.5  -0.5 FAD       13.1 FISH.HOUR  25.9 HOURS.SEA  13.1 Hours.STD     0 Hours.FAD  0    Hours.FSC     0     0     0     0     0     0     0     0
5   2797   509071 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                2     0     1   1.5  -0.5 FAD       26.1 FISH.HOUR  51.8 HOURS.SEA  26.1 Hours.STD     0 Hours.FAD  2.21 Hours.FSC     0     0     0     0     0     0     0     0
6   2797   509072 EU-España EU.ESP-ES-ETRO PS        1994   1990            1 1x1                2     0     2   2.5  -0.5 FAD       12.1 FISH.HOUR  24   HOURS.SEA  12.1 Hours.STD     0 Hours.FAD  0    Hours.FSC     0     0     0     0     0     0     0     0
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
Belize   PS 2009-08-01 2009-09-01  5402000     FS   FISH.HOUR  36.50
Belize   PS 2009-08-01 2009-09-01  5402000     FS   Hours.FSC   3.12
Belize   PS 2009-08-01 2009-09-01  5402000     FS   HOURS.SEA  72.00
Belize   PS 2009-08-01 2009-09-01  5402000     FS   Hours.STD  37.10
Belize   PS 2009-09-01 2009-10-01  5202006     LS   FISH.HOUR  12.10
# Catches
RFMO_CE<-read.csv(path_to_raw_dataset,stringsAsFactors = F)

````{r}
RFMO_CE <- read_excel(path_to_raw_dataset,
                sheet = "Data")

RFMO_CE$FleetCode_short <- sub("-.*", "", RFMO_CE$FleetCode)

names(RFMO_CE)[names(RFMO_CE) == 'FleetCode_short'] <- 'FishingFleet'
RFMO_CE <- RFMO_CE[, c("FishingFleet", setdiff(names(RFMO_CE), "FishingFleet"))] # oput flag in first position
````

# # If we want in the output dataset the column 'FleetCode' instead of 'flag'

if(keep_fleet_instead_of_flag==TRUE){
RFMO_CE$Flag<-NULL
names(RFMO_CE)[names(RFMO_CE) == 'Fishingfleet'] <- 'Flag'
}
# Efforts

````{r}
last_column_not_catch_value=27
RFMO_CE<-RFMO_CE[,-(last_column_not_catch_value:ncol(RFMO_CE))] 

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/FUN_efforts_ICCAT_CE_keep_all_efforts.R")
efforts_pivot_ICCAT<-FUN_efforts_ICCAT_CE_keep_all_efforts(RFMO_CE,c("Eff1","Eff2","Eff3","Eff4","Eff5"),c("Eff1Type","Eff2Type","Eff3Type","Eff4Type","Eff5Type"))
````

School
The format changed, the school is now in the FishMode column

````{r}
efforts_pivot_ICCAT <- efforts_pivot_ICCAT %>% dplyr::mutate(FishMode = ifelse(FishMode == "n/a", "UNK", FishMode)) 
efforts_pivot_ICCAT<- efforts_pivot_ICCAT %>% dplyr::rename("School" = "FishMode")
colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/ICCAT_CE_effort_pivotDSD_to_harmonizedDSD.R")
efforts_pivot_ICCAT <- efforts_pivot_ICCAT %>% dplyr::rename(SquareTypeCode = GeoStrataCode) # to match definition in ICCAT_CE_effort_pivotDSD_to_harmonizedDSD

efforts<-ICCAT_CE_effort_pivotDSD_to_harmonizedDSD(efforts_pivot_ICCAT,colToKeep_efforts)
efforts$AreaName <- as.character(as.integer(efforts$AreaName))
colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"ICCAT"
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)
efforts$geographic_identifier <- format(as.integer(efforts$geographic_identifier), scientific = FALSE, trim = TRUE)

````

sorties same folder as path_to_raw_dataset 

````{r}
output_name_dataset <- "Dataset_harmonized.csv"

write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
````

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

````{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
````

@ Handle unmapped values and save the results

````{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
````

Display the first few rows of the mapping summaries

````{r}
print(head(mapping_codelist$recap_mapping))
````

