---
title: "Harmonize CCSBT Nominal Catch Dataset"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'ccsbt', 'nominal', 'data', 'CCSBT_Global_Catch.xlsx')
```

Harmonize CCSBT Nominal Catch Dataset

This function harmonizes the nominal catch dataset provided by the Commission for the Conservation of Southern Bluefin Tuna (CCSBT), preparing it for integration into the Tuna Atlas database.


@return None; the function outputs files directly, including a harmonized dataset,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details The function processes input datasets to match the standardized format required for integration into the Tuna Atlas,
         including adjustments to column names, units conversion, and data aggregation.
         Metadata integration is conditional, based on whether it will be loaded into the Tuna Atlas database.

@importFrom dplyr %>% filter select mutate group_by summarise
@importFrom readxl read_excel
@importFrom reshape melt
@seealso \code{\link{format_time_db_format}} for converting CCSBT Longline data structure.
@export
@keywords data harmonization, fisheries, CCSBT, tuna
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}

# Catches
Input data sample (after importing as data.frame in R):
A tibble: 6 × 6
Calendar_Year Flag_Code Flag         Ocean    Gear      Catch_mt
<dbl> <chr>     <chr>        <chr>    <chr>        <dbl>
1          1965 AU        Australia    Indian   Unspecif…   4675. 
2          1965 AU        Australia    Pacific  Unspecif…   2201. 
3          1965 JP        Japan        Atlantic Longline      15.3
4          1965 JP        Japan        Indian   Longline   28095. 
5          1965 JP        Japan        Pacific  Longline   12579. 
6          1965 ZA        South Africa Indian   Longline       2 
final data sample:
fishing_fleet gear_type time_start   time_end geographic_identifier fishing_mode species measurement_type
1            JP  Longline 1965-01-01 1965-01-12              Atlantic          ALL     SBF              ALL
2            JP  Longline 1968-01-01 1968-01-12              Atlantic          ALL     SBF              ALL
3            JP  Longline 1969-01-01 1969-01-12              Atlantic          ALL     SBF              ALL
4            JP  Longline 1970-01-01 1970-01-12              Atlantic          ALL     SBF              ALL
5            JP  Longline 1971-01-01 1971-01-12              Atlantic          ALL     SBF              ALL
6            JP  Longline 1972-01-01 1972-01-12              Atlantic          ALL     SBF              ALL
measurement_unit measurement_value source_authority
1                t          15.33201            CCSBT
2                t         411.48727            CCSBT
3                t        1869.37842            CCSBT
4                t        7574.64216            CCSBT
5                t        2125.58909            CCSBT
6                t        3928.10401            CCSBT

```{r}
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/format_time_db_format.R")
```

packages

```{r}
  if(!require(reshape)){
    install.packages("reshape")
    require(reshape)
  }
  if(!require(readxl)){
    install.packages("readxl")
    require(readxl)
  }
  if(!require(dplyr)){
    install.packages("dplyr")
    require(dplyr)
  }
```


```{r}
  opts <- options()
  options(encoding = "UTF-8")
```


```{r}
  CCSBT_NC <- readxl::read_excel(path_to_raw_dataset, sheet = "Sheet1")
  
  CCSBT_NC <- CCSBT_NC %>% dplyr::select(Year = Calendar_Year, fishing_fleet = Flag_Code, 
                                  geographic_identifier = Ocean, gear_type = Gear, 
                                  measurement_value = Catch_mt)
```

Year and period

```{r}
  CCSBT_NC$MonthStart<-1
  CCSBT_NC$Period<-12
```

Format inputDataset time to have the time format of the DB, which is one column time_start and one time_end

```{r}
  CCSBT_NC<-format_time_db_format(CCSBT_NC)
```

School

```{r}
  CCSBT_NC$fishing_mode<-"ALL"
```

Species

```{r}
  CCSBT_NC$species<-"SBF"
```

CatchType

```{r}
  CCSBT_NC$measurement_type<-"NC"
```

Geographic identifier

```{r}
  CCSBT_NC <- CCSBT_NC  %>% dplyr::mutate(geographic_identifier = case_when(geographic_identifier == "Indian"~"IOTC", 
                                        geographic_identifier == "Pacific" ~ "WCPO", TRUE ~ geographic_identifier))
```

measurement_unit

```{r}
  CCSBT_NC$measurement_unit<-"t"
```

remove 0 and NA values 

```{r}
  CCSBT_NC <- CCSBT_NC[CCSBT_NC$measurement_value != 0,]
  CCSBT_NC <- CCSBT_NC[!is.na(CCSBT_NC$measurement_value),] 
  
  NC <- aggregate(CCSBT_NC$measurement_value,
                  FUN = sum,
                  by = list(
                    fishing_fleet = CCSBT_NC$fishing_fleet,
                    gear_type = CCSBT_NC$gear_type,
                    time_start = CCSBT_NC$time_start,
                    time_end = CCSBT_NC$time_end,
                    geographic_identifier = CCSBT_NC$geographic_identifier,
                    fishing_mode = CCSBT_NC$fishing_mode,
                    species = CCSBT_NC$species,
                    measurement_type = CCSBT_NC$measurement_type,
                    measurement_unit = CCSBT_NC$measurement_unit
                  )
  )
  
  colnames(NC)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
  
  NC$source_authority<-"CCSBT"
  NC$measurement <- "catch" 
```


```{r}
  NC$time_start <- as.Date(NC$time_start)
  NC$time_end <- as.Date(NC$time_end)
  dataset_temporal_extent <- paste(
    paste0(format(min(NC$time_start), "%Y"), "-01-01"),
    paste0(format(max(NC$time_end), "%Y"), "-12-31"),
    sep = "/"
  )
  
  output_name_dataset <- "Dataset_harmonized.csv"
  write.csv(NC, output_name_dataset, row.names = FALSE)
georef_dataset <- NC
```

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

