---
title: "Harmonize IOTC Surface Effort Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'iotc', 'effort', 'data', 'IOTC-DATASETS-2023-04-24-CE-Surface_1950-2021.csv')
```

Harmonize IOTC Surface Effort Datasets

This function processes and harmonizes Indian Ocean Tuna Commission (IOTC) surface effort datasets.
It prepares the data for integration into the Tuna Atlas database by ensuring compliance with
data standardization requirements and optionally includes metadata if the dataset is intended for database loading.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to include only essential fields, performs any necessary calculations
         for effort units, and standardizes the format for date fields and geographical identifiers.
         Metadata integration is contingent on the final use of the dataset within the Tuna Atlas database.

@importFrom dplyr filter mutate
@importFrom readr read_csv write_csv
@seealso \code{\link{FUN_efforts_IOTC_CE}} for initial effort data processing,
         \code{\link{IOTC_CE_effort_pivotDSD_to_harmonizedDSD}} for converting effort data to a standardized structure.
@export
@keywords IOTC, tuna, fisheries, data harmonization, effort data
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}

```{r}
if(!require(readr)){
  install.packages("readr")
}

require(rtunaatlas)
require(readr)
  if(!require(dplyr)){
    install.packages("dplyr")
    require(dplyr)
  }
```

Input data sample:
Fleet Gear Year MonthStart MonthEnd      iGrid    Grid Effort EffortUnits QualityCode Source CatchUnits YFT.FS YFT.LS YFT.UNCL BET.FS BET.LS BET.UNCL SKJ.FS SKJ.LS SKJ.UNCL ALB.FS ALB.LS ALB.UNCL SBF.FS SBF.LS
EUESP     PS 1996          9        9 5100043    5100043   36.2      FHOURS           3     LO         MT     NA  13.36       NA     NA   5.58       NA     NA  20.26       NA     NA     NA       NA     NA     NA
EUESP     PS 2004          8        8 5100043    5100043   12.1      FHOURS           3     LO         MT     NA   8.74       NA     NA   2.66       NA     NA  32.97       NA     NA     NA       NA     NA     NA
NEIPS     PS 1992         10       10 5100043    5100043   12.1      FHOURS           3   RFRI         MT     NA   4.53       NA     NA   3.00       NA     NA  37.47       NA     NA     NA       NA     NA     NA
EUESP     PS 1991          8        8 5100044    5100044   12.1      FHOURS           3     LO         MT  76.26     NA       NA   9.95     NA       NA  68.14     NA       NA     NA     NA       NA     NA     NA
EUESP     PS 1995          6        6 5100044    5100044   12.1      FHOURS           3     LO         MT     NA   0.79       NA     NA   0.68       NA     NA   4.93       NA     NA     NA       NA     NA     NA
EUESP     PS 1995          9        9 5100044    5100044   12.1      FHOURS           3     LO                NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA
SBF.UNCL LOT.FS LOT.LS LOT.UNCL FRZ.FS FRZ.LS FRZ.UNCL KAW.FS KAW.LS KAW.UNCL COM.FS COM.LS COM.UNCL TUX.FS TUX.LS TUX.UNCL FAL.FS FAL.LS FAL.UNCL OCS.FS OCS.LS OCS.UNCL SKH.FS SKH.LS SKH.UNCL NTAD.FS NTAD.LS
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA     NA     NA       NA      NA      NA
NTAD.UNCL
NA
NA
NA
NA
NA
NA
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
AUS   BB 1992-01-01 1992-02-01  6230130    IND       HRSRH    403
AUS   BB 1992-02-01 1992-03-01  6230130    IND       HRSRH    366
AUS   BB 1992-02-01 1992-03-01  6230135    IND       HRSRH     30
AUS   BB 1992-02-01 1992-03-01  6235115    IND       HRSRH     23
AUS   BB 1992-03-01 1992-04-01  6230130    IND       HRSRH    221
AUS   BB 1992-03-01 1992-04-01  6235130    IND       HRSRH     68
----------------------------------------------------------------------------------------------------------------------------
Historical name for the dataset at source  IOTC-DATASETS-2023-04-24-CE-Surface_1950-2021.csv

```{r}
  opts <- options()
  options(encoding = "UTF-8")  
```

#Efforts

```{r}
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/FUN_efforts_IOTC_CE.R")
efforts_pivot_IOTC<-FUN_efforts_IOTC_CE(path_to_raw_dataset,12)
efforts_pivot_IOTC$CatchUnits<-NULL
efforts_pivot_IOTC$Source<-NULL

colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/IOTC_CE_effort_pivotDSD_to_harmonizedDSD.R")
efforts<-IOTC_CE_effort_pivotDSD_to_harmonizedDSD(efforts_pivot_IOTC,colToKeep_efforts)


colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"IOTC"
efforts$measurement <- "efforts"
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
```

----------------------------------------------------------------------------------------------------------------------------
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

