---
title: "Harmonize IOTC Coastal and Longline Effort Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'iotc', 'effort', 'data', 'IOTC-DATASETS-2023-04-24-CE-Longline_1950-2021.csv')
```

Harmonize IOTC Coastal and Longline Effort Datasets

This function processes and harmonizes Indian Ocean Tuna Commission (IOTC) effort datasets for
coastal and longline fisheries. It prepares the data for integration into the Tuna Atlas database,
ensuring compliance with data standardization requirements and optionally including metadata
if the dataset is intended for database loading.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to include only essential fields, performs any necessary calculations
         for effort units, and standardizes the format for date fields and geographical identifiers.
         Metadata integration is contingent on the final use of the dataset within the Tuna Atlas database.

@importFrom dplyr filter mutate
@importFrom readr read_csv write_csv
@seealso \code{\link{FUN_efforts_IOTC_CE}} for initial effort data processing,
         \code{\link{IOTC_CE_effort_pivotDSD_to_harmonizedDSD}} for converting effort data to a standardized structure.
@export
@keywords IOTC, tuna, fisheries, data harmonization, effort data
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}

```{r}
  if(!require(dplyr)){
    install.packages("dplyr")
    require(dplyr)
  }
```

Input data sample:
Fleet Gear Year MonthStart MonthEnd      iGrid    Grid Effort EffortUnits QualityCode Source YFT.NO YFT.MT BET.NO BET.MT SKJ.NO SKJ.MT ALB.NO ALB.MT SBF.NO SBF.MT LOT.NO LOT.MT FRZ.NO FRZ.MT KAW.NO KAW.MT
MDV     LLCO 2014         12       12 5100067    5100067   6000       HOOKS           0     LO     NA   0.25     NA   0.55     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
MDV     HAND 2013          2        2 5100069    5100069      2       FDAYS           0     LO     NA  74.60     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
MDV     HAND 2013          3        3 5100069    5100069      1       FDAYS           0     LO     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
MDV     HAND 2013          4        4 5100069    5100069      5       FDAYS           0     LO     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
MDV     HAND 2013          5        5 5100069    5100069      1       FDAYS           0     LO     NA   0.18     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
MDV     HAND 2013          7        7 5100069    5100069      4       FDAYS           0     LO     NA   3.14     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA     NA
COM.NO COM.MT SWO.NO SWO.MT BILL.NO BILL.MT TUX.NO TUX.MT SKH.NO SKH.MT NTAD.NO NTAD.MT
NA     NA     NA   1.07      NA    0.02     NA     NA     NA     NA      NA    0.01
NA     NA     NA     NA      NA      NA     NA     NA     NA     NA      NA      NA
NA     NA     NA     NA      NA      NA     NA     NA     NA     NA      NA      NA
NA     NA     NA     NA      NA      NA     NA     NA     NA     NA      NA      NA
NA     NA     NA     NA      NA      NA     NA     NA     NA     NA      NA      NA
NA     NA     NA     NA      NA      NA     NA     NA     NA     NA      NA      NA
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
AUS HAND 2001-04-01 2001-05-01  5221114    IND        DAYS      4
AUS HAND 2001-04-01 2001-05-01  5229114    IND        DAYS      1
AUS HAND 2001-05-01 2001-06-01  5221114    IND        DAYS      3
AUS HAND 2001-06-01 2001-07-01  5221114    IND        DAYS      9
AUS HAND 2001-07-01 2001-08-01  5221114    IND        DAYS      3
AUS HAND 2001-07-01 2001-08-01  5225113    IND        DAYS      1
Historical name for the dataset at source  IOTC-DATASETS-2023-04-24-CE-Coastal_1950-2021.csv

```{r}
  opts <- options()
  options(encoding = "UTF-8")
```

#Efforts

```{r}
colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/FUN_efforts_IOTC_CE.R")
efforts_pivot_IOTC<-FUN_efforts_IOTC_CE(path_to_raw_dataset,11)

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/IOTC_CE_effort_pivotDSD_to_harmonizedDSD.R")
efforts<-IOTC_CE_effort_pivotDSD_to_harmonizedDSD(efforts_pivot_IOTC,colToKeep_efforts)

colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"IOTC"
efforts$measurement<-"effort"
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
```

----------------------------------------------------------------------------------------------------------------------------
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

