---
title: '`r ifelse(exists("title"), title, "Comparison of two datasets")`'
date: '`r format(Sys.time(), "%d %B, %Y")`'
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
csl: dmk-format.csl
link-citations: TRUE
output:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter"]
    toc: TRUE
    keep_tex: TRUE
    template: template.tex
    number_sections: TRUE
    citation_package: default
  bookdown::html_document2:
  bookdown::word_document2:
    toc: TRUE
lang: en
geometry: "margin=25mm"
papersize: a4
endfloat: FALSE
numberlines: FALSE
authblk: FALSE
footnotehyper: FALSE
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE
tables: TRUE
fancyhdr:
  first:
    headleft: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headright: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headrulewidth: 0pt
  subsequent:
    headleft: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headright: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headrulewidth: 1pt
    footrulewidth: 0pt
---

```{r setup, include=FALSE}
library(knitr)
library(rmarkdown)
library(ggplot2)
# Chargez d'autres packages nécessaires ici
```

```{r parameter-settings, include=FALSE}
# Contenu de Parameters_settings.R
# Définissez vos paramètres ici
```

---
title: '`r ifelse(exists("title"), title, "Comparison of two datasets")`'
date: '`r format(Sys.time(), "%d %B, %Y")`'
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
csl: dmk-format.csl
link-citations: TRUE
output:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter"]
    toc: TRUE
    keep_tex: TRUE
    template: template.tex
    number_sections: TRUE
    citation_package: default
  bookdown::html_document2:
  bookdown::word_document2:
    toc: TRUE
lang: en
geometry: "margin=25mm"
papersize: a4
endfloat: FALSE
numberlines: FALSE
authblk: FALSE
footnotehyper: FALSE
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE
tables: TRUE
fancyhdr:
  first:
    headleft: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headright: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headrulewidth: 0pt
  subsequent:
    headleft: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headright: '`r ifelse(exists("headleft_title"), headleft_title, "Comparison")`'
    headrulewidth: 1pt
    footrulewidth: 0pt
---

```{r setup, include=FALSE}
library(knitr)
library(rmarkdown)
library(ggplot2)
# Chargez d'autres packages nécessaires ici
```

```{r parameter-settings, include=FALSE}
# Contenu de Parameters_settings.R
# Définissez vos paramètres ici
```

---
title: "Setup_markdown"
author: "BastienG"
date: "2023-09-13"
output: html_document
---

```{r include=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	error = 0,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	tab.cap.pre = "Table ",
	tab.cap.sep = ": ", results = 'asis'
)
base::options(knitr.duplicate.label = "allow")

```


```{r options, include=FALSE}

base::options(scipen=9999)
plotting_type <- "plot" 
if (knitr::is_html_output()){plotting_type <- "view" }
tmap_mode(plotting_type)

```


```{r message=FALSE, warning=FALSE, include=FALSE}

set_flextable_defaults(
  font.size = 10,
  font.color = "black",
  digits = 2,
  theme_fun = "theme_box"
)


#set the dataset caption styling
knitr::opts_chunk$set(tab.cap.pre = "Table ", tab.cap.sep = ": ")

#set the dataset caption styling
autonum <- officer::run_autonum(seq_id = "tab", bkm = "TC1", bkm_all = TRUE) # number the table, bkm (bookmark) is important as the cross-referencing is done using the bookmark


```

```{r explenation, results='asis'}
---
title: "Explenation"
author: "BastienG"
date: "2023-09-08"
output: html_document
---

```{r explanationongainandloss, results='asis', eval=(!parameter_short && !unique_analyse)}

if(!parameter_short && !unique_analyse) {
  cat(
    "*Attention ! In the following document:*",
    "- ***All the differences inferior to 0 corresponds to gain in captures.***", 
    "- ***The initial dataset, referred as, dataset 1 is", titre_1, "***",
    "- ***The final dataset, referred as, dataset 2 is", titre_2, "***", 
    sep = "\n"
  )
}

```

```{r results='asis'}

if(!parameter_mapped && !unique_analyse){
  
  cat("In this comparison, the differences due to the mapping, that is to say, the renaming of codes to harmonize data, will not be presented. The recap of the mapping is presented in the annex.", sep = "\n")
  
}


```

```{r checkifnodiff,  eval=!unique_analyse , results='asis'}

if (identical(init, final)) {
  
  cat("There are no differences between the two datasets; this step is not changing the data in any way")
  
  knitr::knit_exit(fully = FALSE)
  
}

```


```{r filterprinting, echo=FALSE, results='asis'}

if (!isNullList(parameter_filtering)){
  cat (paste0("The filter used on this data are:  \n "))
  for (i in 1:length(parameter_filtering)){
    if (parameter_filtering[i] %in% parameter_filtering & !is.null(parameter_filtering[[i]])){
      cat(paste0("***- On ", names((parameter_filtering)[i]), " : ", paste((parameter_filtering)[[i]], collapse = " ; "),"*** \n"))
    } else {""}
  }
}

if (!is.null(parameter_resolution_filter)) {

  cat (paste0("The geographical filter used on this data is:  \n "))

  cat(parameter_resolution_filter)
    
  }


```

```

```{r Summarydifferences, results='asis'}
---
title: "Summarydifferences"
author: "BastienG"
date: "2023-09-12"
output: html_document
---


```{r summaryofdifferences}

summary_of_differences <- compute_summary_of_differences(init, final, titre_1, titre_2)

```


```{r summarydiffplotting,echo=FALSE, results='asis', out.width="100%"}

qflextable2(summary_of_differences, captionn = "Summary of the difference between the two datasets", columns_to_color = c("Difference"), save_folder = "Summary", interactive_plot = FALSE)

```

```

```{r Compnumberstratas, results='asis'}
---
title: "Compnumberstratas"
author: "BastienG"
date: "2023-09-12"
output: html_document
---

```{r stratasloss, eval=!unique_analyse, message=FALSE, warning=FALSE, include=FALSE}


nb_ligne_init_millions <- nrow(init)
nb_ligne_final_millions<- nrow(final)
# 
sum_valeur_init <- sum(init$measurement_value, na.rm = TRUE)
sum_valeur_final <- sum(final$measurement_value, na.rm = TRUE)

Groupped_all$Dimension <-as.character(Groupped_all$Dimension)
Groupped_all$Precision <-as.character(Groupped_all$Precision)

strates_perdues <- Groupped_all %>%dplyr::filter(value_sum_2 == 0 & `Loss / Gain` == "Loss")
strates_gagnees <- Groupped_all %>%dplyr::filter(value_sum_1 == 0 & `Loss / Gain` == "Gain")
nombre_strates_perdues <- nrow(strates_perdues)
nombre_strates_gagnees <- nrow(strates_gagnees)
nombres_de_strates_totales <- nrow(Groupped_all)
pourcentage_strates_perdues <- 100-(100*((nombres_de_strates_totales-nombre_strates_perdues)/nombres_de_strates_totales))
rm(init_no, init_t, final_no, final_t)
gc()

strates_perdues_first_10 <- rbind(strates_perdues,strates_gagnees) %>%dplyr::filter(`Loss / Gain` != 'Egal')%>%ungroup %>%dplyr::filter(Dimension!="geographic_identifier") %>%dplyr::group_by(`Loss / Gain`,Dimension,measurement_unit)%>%# slice(1:10) %>%
  dplyr::select(Dimension, everything())

if (nrow(strates_perdues_first_10)== 0){strates_perdues_first_10 <- rbind(strates_perdues,strates_gagnees) %>%dplyr::filter(`Loss / Gain` != 'Egal')%>%ungroup %>%dplyr::filter(Dimension!="geographic_identifier") %>%dplyr::group_by(`Loss / Gain`,Dimension,measurement_unit)%>% #slice(1:10) %>%
  dplyr::select(Dimension, everything()) }

strates_perdues_first_10[strates_perdues_first_10==""] <- "NA"
strates_perdues_first_10[is.na(strates_perdues_first_10)] <- "NA"

```


```{r compnumberstratas, eval=!unique_analyse}

number_init_column <- as.data.frame(t(init %>% dplyr::select(-measurement_value) %>% summarise_all(list(~n_distinct(.))))) %>%  dplyr::rename(!!as.symbol(eval(parse(text = "titre_1"))) := V1)
number_final_column <- as.data.frame(t(final %>% dplyr::select(-measurement_value) %>% summarise_all(list(~n_distinct(.)))))%>%  dplyr::rename(!!as.symbol(eval(parse(text = "titre_2"))) := V1)

number_final_column<-number_final_column[order(row.names(number_final_column)), ,drop=F]
number_init_column<-number_init_column[order(row.names(number_init_column)), ,drop=F]


number_init_column_final_column <- cbind(number_init_column , number_final_column ) %>% mutate_all(as.numeric)
number_init_column_final_column$Difference <- number_init_column_final_column[,2]-number_init_column_final_column[,1]

rownames(number_init_column_final_column) <- paste0("Number of ", rownames(number_init_column_final_column))

number_init_column_final_column <- as.data.frame(number_init_column_final_column %>% tibble::rownames_to_column(" "))

```


```{r diffstratas,echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
if (nrow(strates_perdues_first_10)!=0){

  disapandap <- strates_perdues_first_10 %>%
  dplyr::mutate(`Loss / Gain` = dplyr::case_when(
    loss > 1 ~ "Loss",
    abs(loss) <= 1 ~ "No differences",
    loss < 1 ~ "Gain",
    TRUE ~ NA_character_
  )) %>%
  dplyr::mutate(`Loss / Gain` = dplyr::case_when(
    is.na(`Loss / Gain`) ~ "Gain",
    value_sum_1 == value_sum_2 ~ "Egal",
    TRUE ~ `Loss / Gain`
  )) %>%
  dplyr::ungroup() %>%
  dplyr::rename(
    `Values dataset 1` = "value_sum_1",
    `Values dataset 2` = "value_sum_2"
  ) %>%
  dplyr::select(parameter_columns_to_keep) %>%
  dplyr::group_by(Dimension, measurement_unit, `Loss / Gain`) %>%
  dplyr::arrange(Dimension, measurement_unit, `Loss / Gain`, desc(`Difference in value`)) %>%
  dplyr::ungroup() %>%  dplyr::select(Dimension, measurement_unit, `Loss / Gain`, Precision, `Difference in millions` = `Difference in value`)
}

```


```{r diffstratas,echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}

if (nrow(strates_perdues_first_10)!=0){
cat("The strata differences (completely lost or appearing) between the first one and the second one (representing ", round(pourcentage_strates_perdues), " % of the total number of strata) are :")

} else {
  cat("No stratum is gained nor lost")
}

```


```{r disapandapflextable, eval = (nrow(strates_perdues_first_10)!=0),results="asis"}

qflextable2(disapandap, captionn=paste0("Disappearing or appearing strata between ",`titre_1`," and ",`titre_2`),columns_to_color =c("Difference in millions"), grouped_data = c("Dimension" ,"measurement_unit",  "Loss / Gain" ),save_folder="Diffstratas", fig.pathinside = fig.path, interactive_plot = FALSE, find_and_print = FALSE)


```

```{r compnumberstratascolumns,  results='asis', out.width="100%"}

qflextable2(number_init_column_final_column, captionn = "Comparison of number of stratas between the two datasets", columns_to_color = c("Difference"), save_folder = "Diffstratas", fig.pathinside = fig.path, interactive_plot = FALSE)

```

```

```{r Main-differences, results='asis'}
---
title: "Main_differences"
author: "BastienG"
date: "2023-11-03"
output: html_document
---

`r cat_title("# Differences between the two datasets", child_headerinside = child_header)`


```{r timediffchild, results='asis', eval = (length(parameter_time_dimension) != 0)}

res <- knitr::knit_child("Timediff.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

\clearpage

```{r Geodiff, results='asis', eval = (length(parameter_geographical_dimension) != 0)}

res <- knitr::knit_child("Geographicdiff.Rmd", quiet = TRUE, envir = environment())
cat(res, sep =  "\\clearpage")

```

\clearpage

```{r Diff-for-other-dimensions, results='asis', eval = (length(Other_dimensions) != 0)}

res <- knitr::knit_child("Differences_for_each_dimensions.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

\clearpage
```

```{r Main-caracteristics-of-dataset, results='asis'}
---
title: "Main_caracteristics_of_dataset"
author: "BastienG"
date: "2023-11-03"
output: html_document
---

`r if(!unique_analyse){cat_title("# Main characteristics of the two datasets", child_headerinside = child_header)}`


```{r timecoverage, echo=FALSE, results='asis', eval = (length(parameter_time_dimension) != 0)}

res <- knitr::knit_child("Timecoverage.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

\clearpage

```{r eval= print_map, results='asis', eval = (length(parameter_geographical_dimension) != 0)}

res <- knitr::knit_child("Spatialcoverage.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```


\clearpage


```{r otherdim, results='asis', eval = length(Other_dimensions) != 0}

res <- knitr::knit_child("Otherdimensions.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

\clearpage
```

```{r Recap-without-mapping, results='asis'}
```

```{r Annexe, results='asis'}
---
title: "Annexe"
author: "BastienG"
date: "2023-09-08"
output:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter"]
    toc: FALSE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
  bookdown::html_document2:
  bookdown::word_document2: 
    # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
---

`r if(!parameter_mapped){cat_title("# Annexe", child_headerinside = child_header)}`

```{r eval=!parameter_mapped}

df_mapping_final_this_dimension <- read_csv("data/mapping_codelist_summary.csv")

df_mapping_final_this_dimension_tidy <- df_mapping_final_this_dimension %>% dplyr::select(-trg_codingsystem)%>%dplyr::group_by(across(everything())) %>%  dplyr::mutate(src_codingsystem = unlist(str_split(src_codingsystem, "_"))[1]) %>% dplyr::ungroup() 

df_mapping_final_this_dimension_tidy_pivoted<-pivot_wider(df_mapping_final_this_dimension_tidy %>% dplyr::mutate(src_code = as.character(src_code)), names_from = source_authority,names_prefix = "Ancient name for ", values_from = "src_code", values_fn = list)  %>% dplyr::rename("New code" = trg_code) %>% dplyr::rename(Dimension = src_codingsystem)%>%
    relocate(Dimension)
mapping <- as_grouped_data(df_mapping_final_this_dimension_tidy_pivoted, groups = c("Dimension")) %>% as.data.frame() %>% dplyr::ungroup()

mapping_final <- mapping %>% rowwise() %>% dplyr::mutate(Dimension = ifelse(is.na(Dimension), " ", Dimension)) %>% dplyr::ungroup() %>% as.data.frame() %>% dplyr::mutate(`New code` = ifelse(is.na(`New code`) &  (Dimension == " "), "UNK", `New code`))
mapping_final[is.na(mapping_final)] <- ""


```

```{r eval=!parameter_mapped}

qflextable2(mapping_final, captionn = "Replacement code during the mapping", folder = "replacement_code_mapping")

```


```
