---
title: "Summary of the treatment for a specific entity of the global tuna atlas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Emmanuel Blondel
    affiliation: FAO
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
# bibliography: ["IOTC-2022-WPDCS18.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter", "makecell"]
    toc: TRUE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
    always_allow_html: TRUE
    keep_md: TRUE
  bookdown::html_document2: 
      # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
    keep_md: true
  bookdown::word_document2: 
      # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
lang: en # Main document language in BCP47 format
abstract: "This document is aimed to recap the impact on the data of the different treatment of the Global Tuna Atlas. \nFirst we present the data created by the GlobalTuna Atlas, as well as the comparison between this data and the rawdata provided by the RFMO’s before the global Tuna Atlas treatment. \nIn a second time we will present the process of the GlobalTuna Atlas in order to reference exactly the choices made."
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: FALSE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
fancyhdr:
  first:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 0pt
    #footleft: A left foot
    footrulewidth: 0pt
  subsequent:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 1pt
    footrulewidth: 0pt
---


```{r setuprmd, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

res <- knitr::knit_child("Setup_markdown.Rmd", quiet = TRUE, envir = parent.frame())

```

# Summary of the final provided data : `r gsub("_", "", entity$identifiers[["id"]])`

```{r analysislast, message=TRUE, warning=TRUE, include=FALSE, results='hide'}

child_env_last <- list2env(as.list(child_env_last_result), parent = new.env())

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/Analysis_markdown/functions/Functions_markdown.R", local = child_env_last)

res_last <- knitr::knit_child("comparison.Rmd", quiet = TRUE,envir = child_env_last)


# rm(child_env_last_result)
rm(child_env_last, envir = environment())
gc()

```

```{r catanalyslast, echo=FALSE, results='asis'}

cat(res_last, sep =  "\\clearpage")

```


```{r cleaninganalysislast, include=FALSE}

rm(res_last, envir = environment())
gc()

```

\clearpage

# Summary of the treatment done on the data (comparison between the Initial data and the data resulting from treatments)


```{r complastfirst, include=FALSE, results='hide'}


child_env_first_to_last <- list2env(as.list(child_env_first_to_last_result), parent =  new.env())

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/Analysis_markdown/functions/Functions_markdown.R", local = child_env_first_to_last)

res_comparison_first_to_last <- knitr::knit_child("comparison.Rmd", quiet = TRUE, envir = child_env_first_to_last)

rm(child_env_first_to_last, envir = environment())
# rm(child_env_first_to_last_result)
gc()

```


```{r complastfirstcat, echo=FALSE, results='asis'}

cat(res_comparison_first_to_last, sep =  "\\clearpage")

```


```{r include=FALSE}

rm(res_comparison_first_to_last, envir = environment())
gc()

```


\clearpage


# Analyse of the processing of the data

The following table recap all the treatment done on the data provided by tRFMOs.

```{r results='asis'}

qflextable2(process_fisheries_data_list$reduced, captionn =process_fisheries_data_list$fig.capp, columns_to_color = process_fisheries_data_list$columns_to_color,  pgwidth = 6, save_folder = "Summary", fig.pathinside = fig.path)

```

```{r message=FALSE, warning=FALSE}

save_image(title= process_fisheries_data_list$fig.capp, plott = process_fisheries_data_list$cowplot, folder = "Summary", fig.pathinside = fig.path)
```


```{r evolamountduringprocess,  echo=FALSE, results='asis', fig.cap = process_fisheries_data_list$fig.capp, out.width='100%'}

process_fisheries_data_list$cowplot

```


# Detail of all the steps of treatment of the data

For each treatment on the data, we provide a summary of what has been added, removed or transformed. This allows to see where major changes occurs for specific data needs.


```{r processallandHIDe, echo=FALSE, results='hide'}

# the comprehensive_cwp_dataframe_analysis function is ran inside the rmd because for a reason I cannot understand, runnning it outside and then passing the result end up in the crashing of pandoc
# Function to process each list item and knit the child document
process_list_item <- function(list_item, template) {
  child_env_result <- list2env(list_item, new.env())
  
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/Analysis_markdown/functions/Functions_markdown.R", local = child_env_result)
  
  res <- knitr::knit_child(template, envir = child_env_result, quiet = TRUE)
  rm(child_env_result)
  gc()
  return(res)
}

# Iterate over all items in all_list and process each one
results <- lapply(all_list, process_list_item, template = "comparison.Rmd")

# Perform garbage collection
gc()

```


```{r unlistallcomp,echo=FALSE, results='asis'}

cat(unlist(results), sep =  "\\clearpage")

```

```{r include=FALSE}

rm(results)
gc()

```


# Annexe

```{r reviewimpact}

df2 <- process_fisheries_data_list$df2

df3 <- flextable(df2) %>% width("Explanation",width = 4)
df3 <- set_caption(df3, caption ="Review of all the impact and purpose of every the treatment done flextable")

```


```{r reviewimpactknitting, results='asis'}

knit_print(df3)

```

