---
title: "Harmonize WCPFC Pole-and-line Effort Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'wcpfc', 'effort', 'data', 'WCPFC_P_PUBLIC_BY_YR_MON.csv')
```

Harmonize WCPFC Pole-and-line Effort Datasets

This function harmonizes the WCPFC Pole-and-line effort datasets, preparing them
for integration into the Tuna Atlas database according to specified format requirements.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@import dplyr
@import tidyr
@import reshape
@importFrom stringr str_replace
@seealso \code{\link{WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD}} for converting WCPFC Driftnet data structure,
@export
@keywords data harmonization, fisheries, WCPFC, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
Input data sample:
YY MM LAT5 LON5 DAYS SKJ_C YFT_C OTH_C
1950  1  30N 135E    0     0     0     0
1950  1  30N 140E    0     0     0     0
1950  1  35N 140E    0     0     0     0
1950  1  40N 140E    0     0     0     0
1950  1  40N 145E    0     0     0     0
1950  2  30N 135E    0     0     0     0
Effort: pivot data sample:
YY MM LAT5 LON5 Effort EffortUnits School Gear
1950  1  30N 135E      0        DAYS    ALL    P
1950  1  30N 140E      0        DAYS    ALL    P
1950  1  35N 140E      0        DAYS    ALL    P
1950  1  40N 140E      0        DAYS    ALL    P
1950  1  40N 145E      0        DAYS    ALL    P
1950  2  30N 135E      0        DAYS    ALL    P
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
ALL    P 1970-03-01 1970-04-01  6200150    ALL        DAYS     82
ALL    P 1970-04-01 1970-05-01  6200150    ALL        DAYS     74
ALL    P 1970-05-01 1970-06-01  6200150    ALL        DAYS     82
ALL    P 1970-06-01 1970-07-01  6200150    ALL        DAYS     81
ALL    P 1970-07-01 1970-08-01  6200150    ALL        DAYS     75
ALL    P 1970-12-01 1971-01-01  6200150    ALL        DAYS     56
packages

```{r}
  if(!require(foreign)){
    install.packages("foreign")
    require(foreign)
  }
  
  if(!require(reshape)){
    install.packages("reshape")
    require(reshape)
  }
  
  
  if(!require(tidyr)){
    install.packages("tidyr")
    require(tidyr)
  }
  
  if(!require(dplyr)){
    install.packages("dplyr")
    require(dplyr)
  }
```

----------------------------------------------------------------------------------------------------------------------------
Historical name for the dataset at source  WCPFC_P_PUBLIC_BY_YR_MON.csv

```{r}
opts <- options()
options(encoding = "UTF-8")
```

#Efforts

```{r}
DF <- read.csv(path_to_raw_dataset)
colnames(DF) <- toupper(colnames(DF))
DF$CWP_GRID <- NULL 
DF <- DF %>% tidyr::gather(variable, value, -c(colnames(DF[1:5])))

DF <- DF %>% dplyr::filter(!value %in% 0) %>% dplyr::filter(!is.na(value))
DF$variable <- as.character(DF$variable)
colnames(DF)[which(colnames(DF) == "variable")] <- "Species"
DF$CatchUnits <- substr(DF$Species, nchar(DF$Species), nchar(DF$Species))
DF$Species <- sub("_C", "", DF$Species)
DF$Species <- sub("_N", "", DF$Species)
DF$School <- "OTH"
DF$EffortUnits <- colnames(DF[5])
colnames(DF)[5] <- "Effort"
efforts_pivot_WCPFC <- DF; rm(DF)
```

-------------
Gear

```{r}
efforts_pivot_WCPFC$Gear<-"P"
```

Catchunits

```{r}
index.kg <- which( efforts_pivot_WCPFC[,"CatchUnits"] == "C" )
efforts_pivot_WCPFC[index.kg,"CatchUnits"]<- "t"

index.nr <- which( efforts_pivot_WCPFC[,"CatchUnits"] == "N" )
efforts_pivot_WCPFC[index.nr,"CatchUnits"]<- "no" 
```

School

```{r}
efforts_pivot_WCPFC$School<-"ALL"

colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD.R")

efforts<-WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD(efforts_pivot_WCPFC,colToKeep_efforts)

colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"WCPFC"
efforts$measurement <- "effort" 
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
```

----------------------------------------------------------------------------------------------------------------------------
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

