---
title: "Harmonize WCPFC Pole-and-Line Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'wcpfc', 'catch', 'data', 'WCPFC_P_PUBLIC_BY_YR_MON.csv')
```

Harmonize WCPFC Pole-and-Line Catch Datasets

This function processes and harmonizes the structure of Western and Central Pacific Fisheries Commission (WCPFC)
pole-and-line catch datasets. The output is tailored for integration into the Tuna Atlas database, aligning with 
the requirements for data standardization and optionally including metadata if the dataset is intended for database loading.


@return None; the function outputs files directly, including harmonized datasets, optional metadata, and code lists for database integration.

@details The function restructures datasets to include only essential fields, performs any necessary calculations for catch units, 
         and standardizes the format for date fields and geographical identifiers. Metadata incorporation is contingent on the final 
         data usage within the Tuna Atlas database.

@importFrom dplyr mutate filter select
@importFrom readr read_csv write_csv
@importFrom tidyr gather
@importFrom reshape2 melt
@seealso \code{\link{WCPFC_CE_catches_pivotDSD_to_harmonizedDSD}} for detailed data structuring operations.
@export
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
@keywords WCPFC, tuna, fisheries, data harmonization, pole-and-line catch
Input data sample:
YY MM LAT5 LON5 DAYS SKJ_C YFT_C OTH_C
1950  1  30N 135E    0     0     0     0
1950  1  30N 140E    0     0     0     0
1950  1  35N 140E    0     0     0     0
1950  1  40N 140E    0     0     0     0
1950  1  40N 145E    0     0     0     0
1950  2  30N 135E    0     0     0     0
Catch: pivot data sample:
YY MM LAT5 LON5 Effort Species  value CatchUnits School EffortUnits Gear
1970  3  05S 150E     82     SKJ 279.16         MT    ALL        DAYS    P
1970  4  05S 150E     74     SKJ 336.61         MT    ALL        DAYS    P
1970  5  05S 150E     82     SKJ 361.80         MT    ALL        DAYS    P
1970  6  05S 150E     81     SKJ 438.48         MT    ALL        DAYS    P
1970  7  05S 150E     75     SKJ 472.75         MT    ALL        DAYS    P
1970 12  05S 150E     56     SKJ 215.05         MT    ALL        DAYS    P
Catch: final data sample:
FishingFleet Gear time_start   time_end AreaName School Species CatchType CatchUnits  Catch
ALL    P 1970-03-01 1970-04-01  6200150    ALL     OTH       ALL         MT   0.01
ALL    P 1970-03-01 1970-04-01  6200150    ALL     SKJ       ALL         MT 279.16
ALL    P 1970-03-01 1970-04-01  6200150    ALL     YFT       ALL         MT  27.81
ALL    P 1970-04-01 1970-05-01  6200150    ALL     OTH       ALL         MT   0.14
ALL    P 1970-04-01 1970-05-01  6200150    ALL     SKJ       ALL         MT 336.61
ALL    P 1970-04-01 1970-05-01  6200150    ALL     YFT       ALL         MT  11.34

```{r}
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/sardara_functions/WCPFC_CE_catches_pivotDSD_to_harmonizedDSD.R")
```

packages 

```{r}
if(!require(reshape)){
  install.packages("reshape")
  require(reshape)
}

if(!require(tidyr)){
  install.packages("tidyr")
  require(tidyr)
}

if(!require(dplyr)){
  install.packages("dplyr")
  require(dplyr)
}
```

Historical name for the dataset at source  WCPFC_P_PUBLIC_BY_YR_MON.csv

```{r}
opts <- options()
options(encoding = "UTF-8")
```

# Catches
# # Reach the catches pivot DSD using a function stored in WCPFC_functions.R
Changes:
- switch to csv
- switch to upper colnames
- CWP grid (removed for the timebeing to apply rtunaatlas codes)
- toupper applied to Species/CatchUnits

```{r}
DF <- read.csv(path_to_raw_dataset)
colnames(DF) <- toupper(colnames(DF))
DF$CWP_GRID <- NULL 

DF <- DF %>% tidyr::gather(variable, value, -c(colnames(DF[1:5])))

DF <- DF %>% dplyr::filter(!value %in% 0) %>% dplyr::filter(!is.na(value))
DF$variable <- as.character(DF$variable)
colnames(DF)[which(colnames(DF) == "variable")] <- "Species"
DF$CatchUnits <- substr(DF$Species, nchar(DF$Species), nchar(DF$Species))
DF$CatchUnits <- toupper(DF$CatchUnits) 
DF$Species <- toupper(DF$Species) 
DF$Species <- sub("_C", "", DF$Species)
DF$Species <- sub("_N", "", DF$Species)
DF$School <- "OTH"
DF$EffortUnits <- colnames(DF[5])
colnames(DF)[5] <- "Effort"
catches_pivot_WCPFC <- DF; rm(DF)
```

Gear

```{r}
catches_pivot_WCPFC$Gear<-"P"
```

Catchunits

```{r}
index.kg <- which( catches_pivot_WCPFC[,"CatchUnits"] == "C" )
catches_pivot_WCPFC[index.kg,"CatchUnits"]<- "t"

index.nr <- which( catches_pivot_WCPFC[,"CatchUnits"] == "N" )
catches_pivot_WCPFC[index.nr,"CatchUnits"]<- "no" 
```

School

```{r}
catches_pivot_WCPFC$School<-"ALL"
```

# # Reach the catches harmonized DSD using a function in WCPFC_functions.R

```{r}
colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
catches<-WCPFC_CE_catches_pivotDSD_to_harmonizedDSD(catches_pivot_WCPFC,colToKeep_captures)

colnames(catches)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
catches$source_authority<-"WCPFC"
catches$measurement_type <- "RC" # Retained catches
catches$measurement <- "catch" 
```


```{r}
catches$time_start <- as.Date(catches$time_start)
catches$time_end <- as.Date(catches$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(catches$time_start), "%Y"), "-01-01"),
	paste0(format(max(catches$time_end), "%Y"), "-12-31"),
	sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
```

@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

