---
title: "Harmonize IATTC Longline Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset_catch <- here::here('tunaatlas_scripts/pre-harmonization', 'iattc', 'catch', 'data', 'PublicCatchOrigFormatTunaBillfish.csv')
path_to_raw_dataset_effort <- here::here('tunaatlas_scripts/pre-harmonization', 'iattc', 'catch', 'data', 'PublicEffortOrigFormatTunaBillfish.csv')
```

Harmonize IATTC Longline Catch Datasets

This function processes and harmonizes Inter-American Tropical Tuna Commission (IATTC) longline
catch datasets for shark and tuna_billfish species. It prepares the data for integration into the
Tuna Atlas database, ensuring compliance with data standardization requirements and optionally
including metadata if the dataset is intended for database loading.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details The function restructures the dataset to include only essential fields, performs any necessary calculations
         for catch units, and standardizes the format for date fields and geographical identifiers.
         Metadata integration is contingent on the final use of the dataset within the Tuna Atlas database.

@importFrom readr read_csv write_csv
@importFrom dplyr filter mutate left_join
@importFrom tidyr gather
@importFrom reshape2 melt
@seealso \code{\link{IATTC_CE_catches_pivotDSD_to_harmonizedDSD}} for specific data structuring operations.
@export
@keywords IATTC, tuna, fisheries, data harmonization, longline catch
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
Catch input data sample:
Record Spp DTypeID Number Weight
11407 ALB       2     17     NA
11407 BUM       2      4     NA
11407 BLM       2      2     NA
11407 SWO       2     10     NA
11407 BET       2    403     NA
11407 BIL       2      1     NA
Catch: final data sample:
FishingFleet Gear time_start   time_end AreaName School Species CatchType CatchUnits Catch
USA   LL 1992-07-01 1992-08-01  6425135    ALL     BSH       ALL         NO     4
USA   LL 1993-04-01 1993-05-01  6425135    ALL     BSH       ALL         NO    75
USA   LL 1993-04-01 1993-05-01  6430135    ALL     BSH       ALL         NO    15
USA   LL 1993-05-01 1993-06-01  6425135    ALL     BSH       ALL         NO    24
USA   LL 1994-03-01 1994-04-01  6425135    ALL     BSH       ALL         NO    14
USA   LL 1994-03-01 1994-04-01  6430135    ALL     BSH       ALL         NO     4
packages

```{r}
if(!require(reshape)){
  install.packages("reshape")
  require(reshape)
}
if(!require(dplyr)){
  install.packages("dplyr")
  require(dplyr)
}
if(!require(tidyr)){
  install.packages("tidyr")
  require(tidyr)
}
```

Historical name for the dataset at source  PublicCatchOrigFormatTunaBillfish.csv or PublicCatchOrigFormatShark.csv
Historical name for the dataset at source  PublicEffortOrigFormatTunaBillfish.csv or PublicEffortOrigFormatShark.csv (keep the same order as for catch)

```{r}
opts <- options()
options(encoding = "UTF-8")
```

# Catches

```{r}
catches<-read.csv(path_to_raw_dataset_catch, stringsAsFactors = F)
efforts<-read.csv(path_to_raw_dataset_effort, stringsAsFactors = F)
catches <- catches %>% tidyr::gather(variable, value, -c("Record","Spp","DTypeID"))
```

remove values=0

```{r}
catches <- catches  %>% 
  dplyr::filter( ! value %in% 0 ) %>%
  dplyr::filter( ! is.na(value)) 
```

Set catchunit values
DType code 1 means that the data was submitted in both weight and number for the same catch
DType code 2 means that the data was submitted as number only
DType code 3 means that the data was submitted as weight only
Initialisation

```{r}
catches$CatchUnits<-'NA'

index.CatchunitMT<-which(catches[,"DTypeID"]==3 & catches[,"variable"]=="Weight")
index.CatchunitNO<-which(catches[,"DTypeID"]==2 & catches[,"variable"]=="Number")
index.CatchunitMTNO<-which(catches[,"DTypeID"]==1 & catches[,"variable"]=="Weight")
index.CatchunitNOMT<-which(catches[,"DTypeID"]==1 & catches[,"variable"]=="Number")

catches$CatchUnits[index.CatchunitMT]<-"t"
catches$CatchUnits[index.CatchunitNO]<-"no"
catches$CatchUnits[index.CatchunitMTNO]<-"MTNO"
catches$CatchUnits[index.CatchunitNOMT]<-"NOMT"
```

Merge catches and efforts to have the strata in the catch file

```{r}
catches<-left_join(catches,efforts,by=c("Record"))
catches <- catches[c("Spp","value","CatchUnits","Year","Month","FlagAbv","Lat","Lon")]

colnames(catches)<-c("variable","value","CatchUnits","Year","Month","Flag","Lat","Lon")

catches$SquareSize<-5
catches$CodeSquareSize<-6
catches$Gear<-"LL"
catches$SetType<-"ALL"

catches$variable[which(catches[,"variable"]=="BuM")]<-"BUM"

colnames(catches)[colnames(catches)=="Flag"] <- "FishingFleet"
```

Reach the catches harmonized DSD using a function in IATTC_functions.R

```{r}
colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/IATTC_CE_catches_pivotDSD_to_harmonizedDSD.R")
catches<-IATTC_CE_catches_pivotDSD_to_harmonizedDSD(catches,colToKeep_captures)

colnames(catches)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
catches$source_authority<-"IATTC"
catches$measurement_type <- "RC" # Retained catches
catches$measurement <- "catch"
```


```{r}
catches$time_start <- as.Date(catches$time_start)
catches$time_end <- as.Date(catches$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(catches$time_start), "%Y"), "-01-01"),
	paste0(format(max(catches$time_end), "%Y"), "-12-31"),
	sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
```

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

