---
title: "Format pre-harmonized data to cwp format"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.
Format pre-harmonized data to cwp format

This function preprocesses catch data for harmonization, including renaming columns, converting data types,
and filtering records. It then exports the processed data along with associated code lists as CSV files.


@details
The function performs several steps to prepare the catch data for further processing or integration:
- Reads the dataset from a specified path.
- Optionally renames the 'fishingfleet' column to 'fishing_fleet' if it exists.
- Converts time fields to character format, then to Date, and numerically transforms the 'measurement_value'.
- Filters out records with a 'measurement_value' of zero.
- Selects specific columns for the final dataset.
- Calculates and sets the temporal extent of the dataset based on the time fields.

It relies on the 'dplyr' package for data manipulation and 'readr' for reading CSV files. The function assumes


@examples
\dontrun{
}

@importFrom dplyr filter mutate rename select
@importFrom readr read_csv
@export

```{r}
  require(dplyr)
```

Historical name for the dataset at source  iccat_catch_all_1m_firms_level0_20230405.csv or iotc_catch_all_1m_firms_level0.csv
Historical name for the dataset at source  iccat_catch_all_1m_firms_level0_20230405.csv or iotc_catch_all_1m_firms_level0.csv
Historical name for the dataset at source  iccat_catch_all_1m_firms_level0_20230405.csv or iotc_catch_all_1m_firms_level0.csv

```{r}
  opts <- options()
  options(encoding = "UTF-8")
```

# Catches

```{r}
  if("fishingfleet" %in%colnames(catches)){
  catches <- catches %>% dplyr::rename(fishing_fleet = fishingfleet) 
  }
  
  catches <- catches %>% dplyr::mutate( time_start = as.character(time_start), time_end = as.character(time_end),  geographic_identifier= as.character(geographic_identifier), measurement_value = as.numeric(catches$measurement_value))
  
  catches <- catches %>% filter(measurement_value!= 0)
  
  catches<-catches %>% dplyr::select("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value", "source_authority")
```


```{r}
  catches$time_start <- as.Date(catches$time_start)
  catches$time_end <- as.Date(catches$time_end)
  dataset_temporal_extent <- paste(
    paste0(format(min(catches$time_start), "%Y"), "-01-01"),
    paste0(format(max(catches$time_end), "%Y"), "-12-31"),
    sep = "/"
  )
```


```{r}
  dataset_temporal_extent <- paste(min(catches$time_start),max(catches$time_end),sep = "/")
  
  output_name_dataset <- "Dataset_harmonized.csv"
  write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
```

@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

