---
title: "Harmonize CCSBT Longline Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is automatically created from the function: https://raw.githubusercontent.com/eblondel/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/southern_hemisphere_oceans_catch_5deg_1m_ll_tunaatlasccsbt_level0.R, the documentation keeps the format of roxygen2 skeleton.

A summary of the mapping process is provided. The path to the dataset is specified. You will find on this same repository on GitHub the first line of each dataset.
The datasets are named after the historical name provided by tRFMOs while exporting and may change.
The information provided in the Rmd allows understanding correctly which dataset should be used in this markdown.

Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.

If you are interested in further details, the results and codes are available for review.

*Each `.Rmd` script requires the user to knit the dataset at the beginning of the script in order to execute the harmonization process correctly.
It is also possible to run the code chunk by chunk but be sure to be in the correct working directory (i.e., the one of the `.Rmd`).*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


````{r}
path_to_raw_dataset <- here::here('R/tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Longline.xlsx')
````

Harmonize CCSBT Longline Catch Datasets

This function harmonizes CCSBT Longline catch datasets for integration into the Tuna Atlas database, ensuring data compliance with specified format requirements.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@importFrom readxl read_excel
@importFrom dplyr %>% filter select mutate
@seealso \code{\link{harmo_time_2}} 
         \code{\link{harmo_spatial_5}} 
         \code{\link{format_time_db_format}} 
@export
@keywords data harmonization, fisheries, CCSBT, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}

````{r}
  # Input data sample (after importing as data.frame in R):
  # YEAR MONTH COUNTRY_CODE TARGET_SPECIES CCSBT_STATISTICAL_AREA LATITUDE LONGITUDE NUMBER_OF_HOOKS NUMBER_OF_SBT_RETAINED
  # 1965     1           JP             NA                      1      -15       100            2083                      4
  # 1965     1           JP             NA                      1      -15       110            9647                      0
  # 1965     1           JP             NA                      1      -15       115           91431                    525
  # 1965     1           JP             NA                      1      -10       100           23560                     56
  # 1965     1           JP             NA                      1      -10       105           31232                     35
  # 1965     1           JP             NA                      1      -10       110            4960                     10
  
  
  # Catch: final data sample:
  # FishingFleet Gear time_start   time_end AreaName School Species CatchType CatchUnits Catch
  #   AU   LL 1987-03-01 1987-04-01  6330130    ALL     SBF       ALL         NO     6
  #   AU   LL 1987-05-01 1987-06-01  6335150    ALL     SBF       ALL         NO     1
  #   AU   LL 1987-06-01 1987-07-01  6330150    ALL     SBF       ALL         NO     2
  #   AU   LL 1987-06-01 1987-07-01  6335150    ALL     SBF       ALL         NO    47
  #   AU   LL 1987-07-01 1987-08-01  6325150    ALL     SBF       ALL         NO     1
  #   AU   LL 1987-09-01 1987-10-01  6330150    ALL     SBF       ALL         NO    14

  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_time_2.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_spatial_5.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/format_time_db_format.R")
  #packages


if(!require(readxl)){
	install.packages("readxl")
	require(readxl)
}
````

Historical name for the dataset at source  CEData_Longline.xlsx

````{r}
opts <- options()
options(encoding = "UTF-8")

 
RFMO_CE<-readxl::read_excel(path_to_raw_dataset, sheet = "CEData_Longline", col_names = TRUE, col_types = NULL,na = "")
colnames(RFMO_CE)<-gsub("\r\n", "_", colnames(RFMO_CE))
colnames(RFMO_CE)<-gsub(" ", "_", colnames(RFMO_CE))
RFMO_CE<-as.data.frame(RFMO_CE)
````

Remove lines that are read in the Excel but that are not real

````{r}
RFMO_CE<- RFMO_CE[!is.na(RFMO_CE$YEAR),]
RFMO_CE$NUMBER_OF_SBT_RETAINED<-as.numeric(RFMO_CE$NUMBER_OF_SBT_RETAINED)
````

FishingFleet

````{r}
RFMO_CE$FishingFleet<-RFMO_CE$COUNTRY_CODE
````

Gear

````{r}
RFMO_CE$Gear<-"Longline"
````

Year and period

````{r}
RFMO_CE<-harmo_time_2(RFMO_CE, "YEAR", "MONTH")
````

Format inputDataset time to have the time format of the DB, which is one column time_start and one time_end

````{r}
RFMO_CE<-format_time_db_format(RFMO_CE)
````

Area 

````{r}
RFMO_CE<-harmo_spatial_5(RFMO_CE,"LATITUDE","LONGITUDE",5,6)
````

School

````{r}
RFMO_CE$School<-"UNK"
````

Species

````{r}
RFMO_CE$Species<-"SBF"
````

CatchType

````{r}
RFMO_CE$CatchType<- "RC" # Retained catches
````

Catch

````{r}
RFMO_CE$Catch<-RFMO_CE$NUMBER_OF_SBT_RETAINED

RFMO_CE$CatchUnits<-"no"

colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
catches <-RFMO_CE[colToKeep_captures]
````

remove whitespaces on columns that should not have withespace

````{r}
catches[,c("AreaName","FishingFleet")]<-as.data.frame(apply(catches[,c("AreaName","FishingFleet")],2,function(x){gsub(" *$","",x)}),stringsAsFactors=FALSE)
````

remove 0 and NA values 

````{r}
catches <- catches[!is.na(catches$Catch),]
catches <- catches[catches$Catch != 0,]

catches <- aggregate(catches$Catch, FUN = sum,
                     by = list(
                       FishingFleet = catches$FishingFleet,
                       Gear = catches$Gear,
                       time_start = catches$time_start,
                       time_end = catches$time_end,
                       AreaName = catches$AreaName,
                       School = catches$School,
                       Species = catches$Species,
                       CatchType = catches$CatchType,
                       CatchUnits = catches$CatchUnits
                     )
)

colnames(catches)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
catches$source_authority<-"CCSBT"
catches$measurement_type <- "RC" # Retained catches
catches$measurement <- "catch"
catches$measurement_processing_level <- "unknown"
catches$time_start <- as.Date(catches$time_start)
catches$time_end <- as.Date(catches$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(catches$time_start), "%Y"), "-01-01"),
  paste0(format(max(catches$time_end), "%Y"), "-12-31"),
  sep = "/"
)

````

output in same folder as path_to_raw_dataset 

````{r}
output_name_dataset <- here::here('R/tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Longline_harmonized.csv')

write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
````

@ Load pre-harmonization scripts and apply mappings

````{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
````

@ Handle unmapped values and save the results

````{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
data.table::fwrite(mapping_codelist$recap_mapping, here::here('R/tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Longline_recap_mapping.csv'))
data.table::fwrite(mapping_codelist$not_mapped_total, here::here('R/tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Longline_not_mapped_total.csv'))
data.table::fwrite(georef_dataset, here::here('R/tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Longline_CWP_dataset.csv'))
````

Display the first few rows of the mapping summaries

````{r}
print(head(mapping_codelist$recap_mapping))
````

