---
title: "Harmonize CCSBT Surface Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'ccsbt', 'catch', 'data', 'CEData_Surface.xlsx')
```

Harmonize CCSBT Surface Catch Datasets

This function harmonizes the CCSBT Surface catch datasets,
preparing them for integration into the Tuna Atlas database, according to specified format requirements.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@import dplyr
@import readxl
@importFrom stringr str_replace
@seealso \code{\link{harmo_time_2}} 
         \code{\link{harmo_spatial_5}} 
         \code{\link{format_time_db_format}} 
@export
@keywords data harmonization, fisheries, CCSBT, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
Input data sample (after importing as data.frame in R):
YEAR MONTH COUNTRY_CODE GEAR_CODE CCSBT_STATISTICAL_AREA LATITUDE LONGITUDE NUMBER_OF_HOURS_SEARCHED WEIGHT_(Kg)_OF_SBT_RETAINED
1975    12           AU        PS                      4      -37       150                       17                       59800
1976     8           AU        BB                      4      -36       150                        8                           0
1976     9           AU        BB                      4      -37       150                       14                       36200
1976     9           AU        BB                      4      -35       150                        4                           0
1976     9           AU        BB                      4      -34       151                       21                           0
1976    10           AU        BB                      4      -37       150                       51                       41000
Catch: final data sample:
FishingFleet Gear time_start   time_end AreaName School Species CatchType CatchUnits Catch
AU   BB 1976-09-01 1976-10-01  6337150    ALL     SBT       ALL         MT  36.2
AU   BB 1976-10-01 1976-11-01  6337150    ALL     SBT       ALL         MT  41.0
AU   BB 1976-12-01 1977-01-01  6332132    ALL     SBT       ALL         MT 167.5
AU   BB 1976-12-01 1977-01-01  6333134    ALL     SBT       ALL         MT  35.6
AU   BB 1976-12-01 1977-01-01  6334134    ALL     SBT       ALL         MT  56.5
AU   BB 1976-12-01 1977-01-01  6334135    ALL     SBT       ALL         MT  37.0

```{r}
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_time_2.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/harmo_spatial_5.R")
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/format_time_db_format.R")
```

packages

```{r}
if(!require(readxl)){
	install.packages("readxl")
	require(readxl)
}
```

Historical name for the dataset at source  CEData_Surface.xlsx

```{r}
opts <- options()
options(encoding = "UTF-8")
```

----------------------------------------------------------------------------------------------------------------------------  

```{r}
RFMO_CE<-readxl::read_excel(path_to_raw_dataset, sheet = "CEData_Surface", col_names = TRUE, col_types = NULL,na = "")
colnames(RFMO_CE)<-gsub("\r\n", "_", colnames(RFMO_CE))
colnames(RFMO_CE)<-gsub(" ", "_", colnames(RFMO_CE))
colnames(RFMO_CE)<-gsub("\\(", "", colnames(RFMO_CE))
colnames(RFMO_CE)<-gsub("\\)", "", colnames(RFMO_CE))
RFMO_CE<-as.data.frame(RFMO_CE)
```

Remove lines that are read in the Excel but that are not real

```{r}
RFMO_CE<- RFMO_CE[!is.na(RFMO_CE$YEAR),] 
RFMO_CE$WEIGHT_Kg_OF_SBT_RETAINED<-as.numeric(RFMO_CE$WEIGHT_Kg_OF_SBT_RETAINED)
RFMO_CE$NUMBER_OF_HOURS_SEARCHED<-as.numeric(RFMO_CE$NUMBER_OF_HOURS_SEARCHED)
```

FishingFleet

```{r}
RFMO_CE$FishingFleet<-RFMO_CE$COUNTRY_CODE
```

Gear

```{r}
RFMO_CE$Gear<-RFMO_CE$GEAR_CODE
```

replace PS by Purse Seine and BB by Pole and Line

```{r}
RFMO_CE$Gear[RFMO_CE$Gear == "PS"] <- "Purse Seine"
RFMO_CE$Gear[RFMO_CE$Gear == "BB"] <- "Pole and Line"
```

Year and period

```{r}
RFMO_CE<-harmo_time_2(RFMO_CE, "YEAR", "MONTH")
```

Format inputDataset time to have the time format of the DB, which is one column time_start and one time_end

```{r}
RFMO_CE<-format_time_db_format(RFMO_CE)
```

Area 

```{r}
RFMO_CE<-harmo_spatial_5(RFMO_CE,"LATITUDE","LONGITUDE",1,5)
```

School

```{r}
RFMO_CE$School<-"ALL"
```

Species

```{r}
RFMO_CE$Species<-"SBF"
```

CatchType

```{r}
RFMO_CE$CatchType<-"ALL"
```

Catch

```{r}
RFMO_CE$Catch<-RFMO_CE$WEIGHT_Kg_OF_SBT_RETAINED/1000

RFMO_CE$CatchUnits<-"t"

colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
catches <-RFMO_CE[,colToKeep_captures]
```

remove whitespaces on columns that should not have withespace

```{r}
catches[,c("AreaName","FishingFleet")]<-as.data.frame(apply(catches[,c("AreaName","FishingFleet")],2,function(x){gsub(" *$","",x)}),stringsAsFactors=FALSE)
```

remove 0 and NA values 

```{r}
catches <- catches[!is.na(catches$Catch),]
catches <- catches[catches$Catch != 0,]


catches <- aggregate(catches$Catch, FUN = sum,
	by = list(
		FishingFleet = catches$FishingFleet,
		Gear = catches$Gear,
		time_start = catches$time_start,
		time_end = catches$time_end,
		AreaName = catches$AreaName,
		School = catches$School,
		Species = catches$Species,
		CatchType = catches$CatchType,
		CatchUnits = catches$CatchUnits
	)
)

colnames(catches)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
catches$source_authority<-"CCSBT"
catches$measurement_type <- "RC" # Retained catches
catches$measurement <- "catch" 
```


```{r}
catches$time_start <- as.Date(catches$time_start)
catches$time_end <- as.Date(catches$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(catches$time_start), "%Y"), "-01-01"),
	paste0(format(max(catches$time_end), "%Y"), "-12-31"),
	sep = "/"
)
output_name_dataset <- "Dataset_harmonized.csv"
write.csv(catches, output_name_dataset, row.names = FALSE)
georef_dataset <- catches
```

@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

