---
title: "Harmonize IATTC Nominal Catch Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('tunaatlas_scripts/pre-harmonization', 'iattc', 'nominal', 'data', 'CatchFlagGear.csv')
```

Harmonize IATTC Nominal Catch Datasets

This function harmonizes the IATTC nominal catch datasets,
preparing them for integration into the Tuna Atlas database, according to specified format requirements.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@import dplyr
@import readr
@importFrom stringr str_replace
@seealso \code{\link{format_time_db_format}} for converting time format,
@export
@keywords data harmonization, fisheries, IATTC, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
Input data sample:
AnoYear BanderaFlag ArteGear EspeciesSpecies    t
1918         OTR       LP             SKJ 1361
1918         OTR       LP             YFT    0
1919         OTR       LP             SKJ 3130
1919         OTR       LP             YFT  136
1920         OTR       LP             SKJ 3583
1920         OTR       LP             YFT  907
Catch: final data sample:
FishingFleet Gear time_start   time_end AreaName School Species CatchType CatchUnits Catch
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     ALB       ALL         MT  4854
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     BET       ALL         MT  1987
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     BIL       ALL         MT   122
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     PBF       ALL         MT   131
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     SFA       ALL         MT    93
BLZ   LL 2001-01-01 2002-01-01    IATTC    ALL     SKH       ALL         MT  1326

```{r}
  source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/format_time_db_format.R")
```

packages

```{r}
if(!require(dplyr)){
  install.packages("dplyr")
  require(dplyr)
}
```


```{r}
opts <- options()
options(encoding = "UTF-8")
```

# # Nominal NC

```{r}
NC <- read.csv(path_to_raw_dataset, header=TRUE, stringsAsFactors=FALSE, strip.white=TRUE)

colToKeep_NC<-c("AnoYear","BanderaFlag","ArteGear","EspeciesSpecies","t")
NC_harm_IATTC<-NC[,colToKeep_NC]
colnames(NC_harm_IATTC)<-c("Year", "Flag","Gear","Species","Catch")

NC_harm_IATTC$AreaName<-"EPO"
NC_harm_IATTC$AreaCWPgrid<-NA
NC_harm_IATTC$School<-"ALL"
NC_harm_IATTC$CatchType<-"ALL"
NC_harm_IATTC$CatchUnits<-"t"
NC_harm_IATTC$RFMO<-"IATTC"
NC_harm_IATTC$Ocean<-"PAC_E"

NC_harm_IATTC$MonthStart<-1
NC_harm_IATTC$Period<-12
```

Format inputDataset time to have the time format of the DB, which is one column time_start and one time_end

```{r}
NC_harm_IATTC<-format_time_db_format(NC_harm_IATTC)

NC <- NC_harm_IATTC[NC_harm_IATTC$Catch != 0,]

rm(NC_harm_IATTC)

colnames(NC)[colnames(NC)=="Flag"] <- "FishingFleet"

colToKeep_captures <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","Species","CatchType","CatchUnits","Catch")
NC <-NC[,colToKeep_captures]
```

remove 0 and NA values 

```{r}
NC <- NC[NC$Catch != 0,]
NC <- NC[!is.na(NC$Catch),] 

NC <- aggregate(NC$Catch,
		FUN = sum,
		by = list(
			FishingFleet = NC$FishingFleet,
			Gear = NC$Gear,
			time_start = NC$time_start,
			time_end = NC$time_end,
			AreaName = NC$AreaName,
			School = NC$School,
			Species = NC$Species,
			CatchType = NC$CatchType,
			CatchUnits = NC$CatchUnits
		)
	)

colnames(NC)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","species","measurement_type","measurement_unit","measurement_value")
NC$source_authority<-"IATTC"
NC$measurement_type<-"NC"
NC$measurement<-"catch"
```


```{r}
NC$time_start <- as.Date(NC$time_start)
NC$time_end <- as.Date(NC$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(NC$time_start), "%Y"), "-01-01"),
	paste0(format(max(NC$time_end), "%Y"), "-12-31"),
	sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(NC, output_name_dataset, row.names = FALSE)
georef_dataset <- NC
```

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "catch"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

