---
title: "Harmonize WCPFC Driftnet Effort Datasets"
output: 
  html_document: 
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is created from a function so the documentation keep the format of roxygen2 skeleton
A summary of the mapping process is provided. The path to the dataset is specified, you will find on this same repository on github the first line of each dataset. The datasets are named after the historical name provided by tRFMOs while exporting and may change. The information provided in the Rmd allows to understand correctly which dataset should be used in this markdown.
Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.
If you are interested in further details, the results and codes are available for review.

```{r}
path_to_raw_dataset <- here::here('R/tunaatlas_scripts/pre-harmonization', 'wcpfc', 'effort', 'data', 'WCPFC_G_PUBLIC_BY_YR_MON.csv')
```

Harmonize WCPFC Driftnet Effort Datasets

This function harmonizes the WCPFC Driftnet effort datasets,
preparing them for integration into the Tuna Atlas database according to specified format requirements.


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@import dplyr
@import tidyr
@import readr
@importFrom stringr str_replace
@seealso \code{\link{WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD}} for converting WCPFC Driftnet data structure,
@export
@keywords data harmonization, fisheries, WCPFC, driftnet, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
This script works with any dataset that has the first 5 columns named and ordered as follow: {YY|MM|LAT5|LON5|DAYS} followed by a list of columns specifing the species codes with "_N" for catches expressed in number and "_T" for catches expressed in tons
packages 

```{r}
  if(!require(reshape)){
    install.packages("reshape")
    require(reshape)
  }
  
  if(!require(tidyr)){
    install.packages("tidyr")
    require(tidyr)
  }
  
  if(!require(dplyr)){
    install.packages("dplyr")
    require(dplyr)
  }
```

Input data sample:
YY MM LAT5 LON5 DAYS ALB_N  ALB_C
1983 11  30S 170W    0     0  0.000
1983 11  35S 170W  133   886  4.960
1983 12  35S 165W    0     0  0.000
1983 12  35S 170W  133   870  4.872
1983 12  40S 165W    0     0  0.000
1983 12  40S 170W  248  3822 21.402
Effort: final data sample:
Flag Gear time_start   time_end AreaName School EffortUnits Effort
ALL    D 1983-11-01 1983-12-01  6330165    ALL        DAYS    133
ALL    D 1983-12-01 1984-01-01  6330165    ALL        DAYS    133
ALL    D 1983-12-01 1984-01-01  6335165    ALL        DAYS    248
ALL    D 1984-01-01 1984-02-01  6230165    ALL        DAYS     85
ALL    D 1984-01-01 1984-02-01  6240160    ALL        DAYS     59
ALL    D 1984-01-01 1984-02-01  6335165    ALL        DAYS    704
----------------------------------------------------------------------------------------------------------------------------
Historical name for the dataset at source  WCPFC_G_PUBLIC_BY_YR_MON.csv

```{r}
opts <- options()
options(encoding = "UTF-8")
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
```

#Efforts

```{r}
DF <- read.csv(path_to_raw_dataset)
colnames(DF) <- toupper(colnames(DF))
```

DF <- melt(DF, id = c(colnames(DF[1:5]))) #@juldebar error with melt function from reshape package
DF <- melt(as.data.table(DF), id=c(colnames(DF[1:5])))

```{r}
DF <- DF %>% tidyr::gather(variable, value, -c(colnames(DF[1:5])))

DF <- DF %>% dplyr::filter(!value %in% 0) %>% dplyr::filter(!is.na(value))
DF$variable <- as.character(DF$variable)
colnames(DF)[which(colnames(DF) == "variable")] <- "Species"
DF$CatchUnits <- substr(DF$Species, nchar(DF$Species), nchar(DF$Species))
DF$Species <- toupper(DF$Species) 
DF$Species <- sub("_C", "", DF$Species)
DF$Species <- sub("_N", "", DF$Species)
DF$School <- "OTH"
DF$EffortUnits <- colnames(DF[5])
colnames(DF)[5] <- "Effort"
```

--------------------------------------------------

```{r}
efforts_pivot_WCPFC <- DF; rm(DF)
```

Gear

```{r}
efforts_pivot_WCPFC$Gear<-"D"
```

Catchunits

```{r}
index.kg <- which( efforts_pivot_WCPFC[,"CatchUnits"] == "C" )
efforts_pivot_WCPFC[index.kg,"CatchUnits"]<- "t"

index.nr <- which( efforts_pivot_WCPFC[,"CatchUnits"] == "N" )
efforts_pivot_WCPFC[index.nr,"CatchUnits"]<- "no" 
```

School

```{r}
efforts_pivot_WCPFC$School<-"ALL"

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD.R")
efforts<-WCPFC_CE_efforts_pivotDSD_to_harmonizedDSD(efforts_pivot_WCPFC,colToKeep_efforts)

colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"WCPFC"
efforts$measurement <- "effort" 
```

----------------------------------------------------------------------------------------------------------------------------

```{r}
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
  paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
  paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
  sep = "/"
)

output_name_dataset <- "Dataset_harmonized.csv"
write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
```

----------------------------------------------------------------------------------------------------------------------------
@ Load pre-harmonization scripts and apply mappings

```{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
```

@ Handle unmapped values and save the results

```{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(mapping_codelist$not_mapped_total, 'not_mapped_total.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
```

Display the first few rows of the mapping summaries

```{r}
print(head(mapping_codelist$recap_mapping))
print(head(mapping_codelist$not_mapped_total))
```

