# Introduction

This R Markdown document is designed to transform data that is not in CWP format into CWP format.
Initially, it changes the format of the data; subsequently, it maps the data to adhere to CWP standards.
This markdown is automatically created from the function: https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/refs/heads/master/R/tunaatlas_scripts/pre-harmonization/east_pacific_ocean_effort_1deg_1m_ps_tunaatlasiattc_level0__byflag.R, the documentation keeps the format of roxygen2 skeleton.

A summary of the mapping process is provided. The path to the dataset is specified. You will find on this same repository on GitHub the first line of each dataset.
The datasets are named after the historical name provided by tRFMOs while exporting and may change.
The information provided in the Rmd allows understanding correctly which dataset should be used in this markdown.

Additional operations are performed next to verify other aspects of the data, such as the consistency of the geolocation, the values, and the reported catches in numbers and tons.

If you are interested in further details, the results and codes are available for review.

*Each `.Rmd` script requires the user to knit the dataset at the beginning of the script in order to execute the harmonization process correctly.
It is also possible to run the code chunk by chunk but be sure to be in the correct working directory (i.e., the one of the `.Rmd`).*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


````{r}
path_to_raw_dataset <- 
````

Harmonize IATTC PS ByFlag Effort Datasets

This function harmonizes the effort data for the IATTC PS ByFlag datasets,
preparing them for integration into the Tuna Atlas database, following the required format specifications.
This script works with any data that has the first 5 columns named and ordered as follow: {Year|Month|Flag|LatC1|LonC1|NumSets}


@return None; the function outputs files directly, including harmonized datasets,
        optional metadata, and code lists for integration within the Tuna Atlas database.

@details This function modifies the dataset to ensure compliance with the standardized
         format, including renaming, reordering, and recalculating specific fields as necessary.
         Metadata integration is contingent on the intended use within the Tuna Atlas database.

@import dplyr
@import readr
@importFrom stringr str_replace
@seealso \code{\link{FUN_efforts_IATTC_CE_allbutLLTunaBillfish}} for initial effort data preprocessing,
         \code{\link{IATTC_CE_efforts_pivotDSD_to_harmonizedDSD}} for converting effort data to a standardized structure.
@export
@keywords data harmonization, fisheries, IATTC, tuna
@author Paul Taconet, IRD \email{paul.taconet@ird.fr}
@author Bastien Grasset, IRD \email{bastien.grasset@ird.fr}
packages
Historical name for the dataset at source  PublicPSSharkFlag.csv or PublicPSBillfishFlag.csv or PublicPSTunaFlag.csv

````{r}
opts <- options()
options(encoding = "UTF-8")
````

#Efforts
Reach the efforts pivot DSD using a function in IATTC_functions.R

````{r}
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/FUN_efforts_IATTC_CE_allbutLLTunaBillfish.R")
efforts_pivot_IATTC <-FUN_efforts_IATTC_CE_allbutLLTunaBillfish(path_to_raw_dataset,"NumSets","Flag","PS")

colToKeep_efforts <- c("FishingFleet","Gear","time_start","time_end","AreaName","School","EffortUnits","Effort")
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/sardara_functions/IATTC_CE_efforts_pivotDSD_to_harmonizedDSD.R")
efforts<-IATTC_CE_efforts_pivotDSD_to_harmonizedDSD(efforts_pivot_IATTC,colToKeep_efforts)

colnames(efforts)<-c("fishing_fleet","gear_type","time_start","time_end","geographic_identifier","fishing_mode","measurement_unit","measurement_value")
efforts$source_authority<-"IATTC"
efforts$measurement <- "effort"
efforts$measurement_processing_level <- "unknown" 
efforts$time_start <- as.Date(efforts$time_start)
efforts$time_end <- as.Date(efforts$time_end)
dataset_temporal_extent <- paste(
	paste0(format(min(efforts$time_start), "%Y"), "-01-01"),
	paste0(format(max(efforts$time_end), "%Y"), "-12-31"),
	sep = "/"
)
base1 <- tools::file_path_sans_ext(basename(filename1))
````

sorties same folder as path_to_raw_dataset 

````{r}
output_name_dataset <- "Dataset_harmonized.csv"

write.csv(efforts, output_name_dataset, row.names = FALSE)
georef_dataset <- efforts
````

----------------------------------------------------------------------------------------------------------------------------  
@ Load pre-harmonization scripts and apply mappings

````{r}
download.file('https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/R/tunaatlas_scripts/pre-harmonization/map_codelists_no_DB.R', destfile = 'local_map_codelists_no_DB.R')
source('local_map_codelists_no_DB.R')
fact <- "effort"
mapping_codelist <- map_codelists_no_DB(fact, mapping_dataset = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/global/firms/gta/codelist_mapping_rfmos_to_global.csv", dataset_to_map = georef_dataset, mapping_keep_src_code = FALSE, summary_mapping = TRUE, source_authority_to_map = c("IATTC", "CCSBT", "WCPFC", "ICCAT", "IOTC"))
````

@ Handle unmapped values and save the results

````{r}
georef_dataset <- mapping_codelist$dataset_mapped %>% dplyr::mutate(fishing_fleet = ifelse(fishing_fleet == 'UNK', 'NEI', fishing_fleet), gear_type = ifelse(gear_type == 'UNK', '99.9', gear_type))
fwrite(mapping_codelist$recap_mapping, 'recap_mapping.csv')
fwrite(georef_dataset, 'CWP_dataset.csv')
````

Display the first few rows of the mapping summaries

````{r}
print(head(mapping_codelist$recap_mapping))
````

