---
title: "`r ifelse(exists('title_markdown'), title_markdown, 'Comparison of two datasets')`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "BastienG"
output: html_document
---

```{r setuprmd,include = FALSE, echo=FALSE}

res <- knitr::knit_child("Setup_markdown.Rmd", quiet = TRUE, envir = environment())
# envir = parent.frame() ne marche pas bien car il faut appeler dans le nouveau markdown (child) envir$obj
```

```{r eval=exists("Add_lines") && file.exists(Add_lines), results='asis'}

res <- knitr::knit_child(Add_lines, quiet = TRUE, envir = environment())
cat(res)

```

`r if(step_title_t_f){cat_title(step_title, child_headerinside = child_header)}`

```{r explenation, results='asis'}

res <- knitr::knit_child("Explenation.Rmd", quiet = TRUE, envir = environment())

cat(res)

```

```{r include=FALSE}

if(!unique_analyse && parameter_short == FALSE && coverage){
  main_title <- TRUE
} else {
  main_title <- FALSE
  child_header <- sub("#", "", child_header)
}

```

`r if(main_title){cat_title("# Main differences", child_headerinside = child_header)}`

```{r Compnumberstratas, eval=!unique_analyse, echo=FALSE, results='asis'}

res <- knitr::knit_child("Compnumberstratas.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

```{r parametershort,  eval=parameter_short, results='asis'}

exit <- "This document only presents the summary of the differences between the two datasets. If you need a more detailed summary please provide FALSE as argument for the 'short' parameter"

if(child){exit <- ""}

if (parameter_short){knitr::knit_exit(exit, fully = FALSE)}


```


`r if(!unique_analyse && (exists("plot_titles_list") && !is.null(plot_titles_list))){cat_title("## Differences in temporal data", child_headerinside = child_header)}`


```{r timediffchild, results='asis', eval = !unique_analyse && (exists("plot_titles_list") && !is.null(plot_titles_list))}

res <- knitr::knit_child("Differenceintimedata.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

\clearpage

`r if(!unique_analyse && (exists("Geographicdiff") && !is.null(Geographicdiff))){cat_title("## Differences in geographical data", child_headerinside = child_header)}`


```{r Geodiff, results='asis', eval = !unique_analyse && (exists("Geographicdiff") && !is.null(Geographicdiff))}

res <- knitr::knit_child("Geographicdiff.Rmd", quiet = TRUE, envir = environment())
cat(res, sep =  "\\clearpage")

```

\clearpage

`r if(!unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))){cat_title("## The differences for each other dimension", child_headerinside = child_header)}`


```{r Diff-for-other-dimensions, results='asis', eval = !unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))}

res <- knitr::knit_child("Differencesforeachdimensions.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```

```{r eval = !coverage}

knitr::knit_exit(fully = FALSE)

```



```{r Maincharactofdataset, results='asis', eval = coverage}

res <- knitr::knit_child("Main_caracteristics_of_dataset.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```
