---
title: "SetupMarkdown"
date: "2023-07-19"
---

```{r libraries,message=FALSE, warning=FALSE, include=FALSE}



# Prevent package installation from upgrading
formals(install.packages)$upgrade <- "never"

# List of required packages
required_packages <- c("dplyr", "knitr", "stringr", "purrr", "readxl", "base", "flextable", "remotes", 
                        "utils", "DBI", "odbc", "rlang", "sf", "kableExtra", "readr", "tidyr", "ggplot2", 
                        "stats", "RColorBrewer", "cowplot", "tmap", "RPostgreSQL", "officer", "gdata","tibble")

# Check if each package is already installed, and if not, install and load it
for (package in required_packages) {
  # if (!require(package, character.only = TRUE)) {
  #   install.packages(package)
    require(package)
  # }
}

# Install and load sutdycatchesird package from GitHub if not already installed
# if(!require(sutdycatchesird)){
  # remotes::install_github("BastienIRD/sutdycatchesird", upgrade = "never")
  require(sutdycatchesird)
# }

```



```{r utilityfunctions, include=FALSE}


`%notin%` <- Negate(`%in%`)

last_path = function(x) {
  substr(x, max(gregexpr("/", x)[[1]]) + 1, nchar(x))
}

last_path_reduced = function(x) {
  gsub("georef_dataset", "", last_path(x))
}

is_null_or_not_exist = function(x) {
  var_name <- deparse(substitute(x))
  if (!exists(var_name, envir = parent.frame()) || 
      is.null(eval(parse(text = var_name), envir = parent.frame()))) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
cat_title = function(x, child_headerinside = ifelse(exists("child_header"), child_header, "")) {
  output = paste0(child_headerinside, x, " \n")
    return(output)
}

```



```{r GTAMARKDOWNFUNCTIONS}

fSpatPlan_Convert2PacificRobinson = function(df, buff = 0){
  
  if(!exists("df_robinson")){
    df_robinson <- df}
  return(df_robinson)
  
}

knit_child_map =function(list_map_unit, fig.pathinside = fig.path, folder = "Geographicdiff"){
  assign("unit_name_map", list_map_unit[2][[1]], envir = environment())
  assign("map_for_knit", list_map_unit[1][[1]], envir = environment())
  
  unit_title <- gsub("_","..",unit_name_map)
  save_image(title = paste0("Map of the differences between two datasets for the unit ", unit_title),folder = folder,  plott = map_for_knit,fig.pathinside = fig.pathinside)
  
  knit_child(text = c(
    '```{r results = "asis"}',
    '',
    'unit_title <- gsub("_","..",unit_name_map)',
    '```',
    '',
#     '    ```{r setup, include=FALSE}',
# 'knitr::opts_chunk$set(label = paste0("mapdiff", unit_title))', 
# '```',
'', 
'', 

    '',
    '```{r results="asis", fig.cap = paste0("Map of the differences between two datasets for the unit ", unit_title)}',
    '',
    'map_for_knit',
    '',
    '```',
    '',
    '',''),
    envir = environment(), quiet= TRUE)
}


qflextable2 <- function(x, captionn = NULL, autonumm = autonum, pgwidth = 6, columns_to_color = NULL, save_folder = NULL, fig.pathinside = fig.path, grouped_data = NULL, interactive_plot = FALSE, find_and_print = ifelse(exists("outputonly"), outputonly, FALSE)) {
  
  if(find_and_print){
    filepath <- file.path(fig.pathinside, save_folder, paste0(make.names(captionn), ".png"))
    knitr::knit_child(text =c(
'', 
    '```{r evolvaluedim, fig.cap=`captionn`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'knitr::include_graphics(filepath)',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
  } else {
  if (all(class(x) == "flextable")) {
    flextabley <- x
  } else {
    if (!(all(class(x) == "data.frame"))) {
      x <- as.data.frame(x %>% dplyr::ungroup())
    }
    if (!is.null(save_folder)) {
      if (!dir.exists(file.path(fig.pathinside, save_folder))) {
        dir.create(file.path(fig.pathinside, save_folder))
      }
      save_path_data <- file.path(fig.pathinside, save_folder, paste0(make.names(captionn), ".csv" ))  # Adjust the file name as needed
      fwrite(x, file = save_path_data)
    }

    y <- x %>% 
      dplyr::ungroup() %>% 
      dplyr::mutate_if(is.factor, as.character) %>% 
      dplyr::mutate_if(is.character, ~ str_replace_all(., ",", ", \n" )) %>% 
      dplyr::mutate_if(is.character, ~ str_replace_all(., "_", "-" )) %>% 
      dplyr::mutate_if(is.numeric, function(.) { round(., 2) }) 

    if(!is.null(grouped_data)){
    y <- as_grouped_data(y, groups = grouped_data)
  }

if (!is.null(columns_to_color)) {
    colormatrix <- ifelse(y %>% dplyr::select(all_of(columns_to_color)) < 0, "red", 
                           ifelse(y %>% dplyr::select(all_of(columns_to_color)) > 0, "green", "white"))
    flextabley <- flextable::flextable(y)
    
     flextabley <- flextabley %>% highlight(j = all_of(columns_to_color), color = colormatrix)
    

    } else {    flextabley <- flextable::flextable(y)}
    
}

if (!is.null(captionn)) {
    flextable_captionned <- flextable::set_caption(flextabley, caption = captionn, style = "Table Caption")
  } else {
    flextable_captionned <- flextabley
  }
  

  
  ft_out <- flextable_captionned %>% autofit()
  ft_out <- width(ft_out, width = dim(ft_out)$widths * pgwidth / flextable_dim(ft_out)$widths)
  
  if (!is.null(save_folder)) {
    save_path_flextable <- file.path(fig.pathinside, save_folder, paste0(make.names(captionn), ".png"))  # Adjust the file name as needed
    save_as_image(ft_out, path = save_path_flextable)
  }
  if (interactive_plot) {
    # Create an interactive plot using DT
    return(DT::datatable(y))
  }
  ft_out
  }
}

isNullList = function(x) all(!lengths(x))

filtering_function= function(dataframe_to_filter, parameter_filtering){
  
  matchingList <- parameter_filtering %>% purrr::keep( ~ !is.null(.) )
  if(length(matchingList)!= 0){  
    
        colnames_to_filter <- colnames(dataframe_to_filter %>% dplyr::select(names(matchingList)))

  names(matchingList) <- colnames_to_filter

  matchingList <- lapply(matchingList, function(x){ #handling single filter
  if(length(x) == 1){
  x <- c(x, x) }else {
    x
  }

}
  )
    dataframe_to_filter <- dataframe_to_filter%>% dplyr::filter(!! rlang::parse_expr(str_c(colnames_to_filter, matchingList, sep = '%in%', collapse="&")))}
  return(dataframe_to_filter)
}

tidying_data_for_GTA = function(dataframe, time_dim, shape, species_group_dataframe = species_group, cl_cwp_gear_level2_dataframe = cl_cwp_gear_level2, parameter_colnames_to_keep_dataframe = parameter_colnames_to_keep){
  
  dataframe <- dataframe %>% dplyr::select(any_of(parameter_colnames_to_keep_dataframe))
  
  if("geographic_identifier"%in%colnames(dataframe)){
    dataframe <- dataframe%>%  dplyr::left_join(shape, by = c("geographic_identifier"="CWP_CODE"))
    print_map <- TRUE
  } else {print_map <- FALSE}
  
  
  dataframe <- dataframe %>% mutate_at(all_of(time_dim), as.character)
  
  if("GRIDTYPE"%in%colnames(dataframe)){
    dataframe <- dataframe%>%dplyr::mutate(GRIDTYPE = as.character(GRIDTYPE))
  }
  
  dataframe <- dataframe %>% dplyr::left_join(species_group_dataframe%>% dplyr::distinct(), by = c("species"))
  if("gear_type" %in%colnames(dataframe) ){
    dataframe <- dataframe %>% dplyr::left_join(cl_cwp_gear_level2_dataframe, by = c("gear_type" = "Code"))
  }
  
  
  dataframe <- dataframe%>%dplyr::mutate(measurement_unit = dplyr::case_when(measurement_unit %in% c("MT","t","MTNO", "Tons")~ "Tons", measurement_unit %in% c("NO", "NOMT","no", "Number of fish")~"Number of fish", TRUE ~ measurement_unit)) 
  
  return(list(dataframe = dataframe, print_map = print_map))
}


function_geographic_identifier_renaming_and_not_standards_unit= function(dataframe_to_filter, geo_dim , parameter_fact, parameter_UNK_for_not_standards_unit = TRUE){
  if(  parameter_UNK_for_not_standards_unit & parameter_fact == "effort"){
    dataframe_to_filter <- dataframe_to_filter %>% dplyr::mutate(measurement_unit = ifelse(measurement_unit%in%c("HOOKS","FDAYS"), measurement_unit, "UNK" ))}
  
  if(geo_dim != "geographic_identifier" && "geographic_identifier"%in%colnames(dataframe_to_filter)){
    dataframe_to_filter <- dataframe_to_filter %>% dplyr::rename("geographic_identifier" := {{geo_dim}})
  }
  dataframe_to_filter
}

fonction_groupement = function(these_col, init, final){
  
  # Compute sum of values for each combination of the columns in "these_col" and "measurement_unit" in the "init" dataframe
  groupement_1 <- init %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!!rlang::syms(these_col), measurement_unit) %>%
    dplyr::summarise(value_sum_1 = ((sum(measurement_value, na.rm=TRUE)))) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(value_sum_1 = dplyr::coalesce(value_sum_1, 0)) %>%
    dplyr::mutate(measurement_unit = as.character(measurement_unit)) %>%
    dplyr::mutate(number_lines1 = n())
  
  # Compute sum of values for each combination of the columns in "these_col" and "measurement_unit" in the "final" dataframe
  groupement_2 <- final %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!!rlang::syms(these_col), measurement_unit) %>%
    dplyr::summarise(value_sum_2 = ((sum(measurement_value, na.rm=TRUE)))) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    dplyr::mutate(measurement_unit = as.character(measurement_unit)) %>%
    dplyr::mutate(number_lines2 = n())
  
  # Join the two dataframes based on the columns in "these_col" and "measurement_unit"
  fulljoin <- dplyr::full_join(groupement_1, groupement_2) %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    dplyr::mutate(value_sum_1 = dplyr::coalesce(value_sum_1, 0)) %>%
    dplyr::mutate(loss = value_sum_1 - value_sum_2) %>%
    # Compute whether the difference in value is a gain or loss
    dplyr::mutate(`Loss / Gain` = dplyr::case_when(abs(loss) <= 1 ~ "Egal", loss > 1 ~"Loss",  loss < 1 ~"Gain")) %>%
    dplyr::mutate(Loss_pourcent = - (100*((value_sum_1 - value_sum_2)/value_sum_1))) %>%
    dplyr::mutate(Dimension = colnames(groupement_1[1])) %>%
    dplyr::rename("Precision" = 1) %>%
    dplyr::mutate(Precision = as.character(Precision)) %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    # Set Loss_pourcent to 100 if it is NA or -Inf or 0 
    dplyr::mutate(Loss_pourcent = base::ifelse(is.na(Loss_pourcent)|Loss_pourcent==-Inf, 100, Loss_pourcent)) %>%
    dplyr::ungroup() %>%
    # Compute the difference in the number of lines between the two dataframes
    dplyr::mutate(loss_nb_ligne = - (number_lines1 - number_lines2)) %>%
    dplyr::mutate(`Difference in value`= - (value_sum_1 - value_sum_2)) %>% #final - init
    dplyr::rename(`Difference (in %)` = Loss_pourcent,`Difference in number of lines` = loss_nb_ligne)%>%
    dplyr::mutate_if(is.numeric, list(~replace_na(., 0)))
  
  return(fulljoin)
}

save_image = function(title, plott = last_plot(), folder = NULL, fig.pathinside = fig.path, find_and_print = FALSE){
  if(find_and_print){
    filepath <- file.path(fig.pathinside, folder, paste0(make.names(title), ".png"))
    knitr::knit_child(text =c(
'', 
    '```{r image, fig.cap=`title`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'knitr::include_graphics(filepath)',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
  } else {
  current <- tmap_mode()
  if(!is.null(folder) && !dir.exists(folder)){
    dir.create(file.path(fig.pathinside, folder))
  }
  
    if(all(class(plott) == "flextable")){
      save_as_image(plott, path = file.path(fig.pathinside, folder))
      

    } else if(all(class(plott) == "tmap")){
    tmap_mode("plot")
    filenametmap <- file.path(fig.pathinside, file.path(folder, paste0( make.names(title), ".png")))
    tmap_save(tm = plott, filename = filenametmap)
    tmap_mode(current)
  } else {

ggsave(paste0( make.names(title), ".png"),plot = plott,   device = "png", path = file.path(fig.pathinside, folder))
# ggsave(paste0( make.names(title), ".pdf"), plot = plott,  device = "pdf", path = file.path(fig.pathinside, folder))
  }
  plott

}
}

resume_knit_child = function(x,find_and_print = FALSE, folder = "Piechartsdimensions", fig.pathinside = fig.path){
    if(find_and_print){
    for (i in files(folder)){
  titles <- gsub(".", "", i)
    knitr::knit_child(text =c(
'', 
    '```{r otherdim, fig.cap=`titles`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'knitr::include_graphics(file.path(fig.path, file.path(folder, i)))',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}

  } else {

  dimension_title <- gsub("_", "-", unique(x$dimension))

  title_knit <- paste0("Distribution in value for the dimension : ",(dimension_title))
  
  assign("dimension_title", gsub("_", "-", unique(x$dimension)), envir = environment())
save_image(title = title_knit, plott = x$pie_chart, folder = folder, fig.pathinside = fig.pathinside)
  assign("x", x, envir = environment())

  knitr::knit_child(text = c(

    '',
    '',
'', 

    '',
    '```{r distribinvaluedim, results = "asis" ,fig.cap =`title_knit`}',
    '',
    'x$pie_chart',
'',
'```',
    '',
    '',
    '',
    '```{r results = "asis", eval= !is.null(x$df)}',
    'x$df',
    '',
    '',
    '```',
    '',
    '', 
    ''
    
    
    
    
  ), envir = environment(), quiet= TRUE)
  }
}

function_pie_chart_df = function(dimension, ...){
  formals(pie_chart_2_default)$dimension <- dimension
  dimension_returned <- dimension
  pie_chart = pie_chart_2_default(...)
  if(exists("pie_chart$df")){
    df = pie_chart$df
    pie_chart = pie_chart$plot} else{
      pie_chart = pie_chart
      df = NULL}
  
  if(!is.null(df)){
    df <- qflextable2(df, captionn = paste0("Strata movement for the dimension : ", dimension_returned))}
  return(list(pie_chart=pie_chart,df=df, dimension=dimension_returned))
}


function_knitting = function(x, titre_init = titre_1, titre_final = titre_2, find_and_print = FALSE,folder, difference = FALSE, fig.pathinside = fig.path, unique_analyse = FALSE){
  if(find_and_print){
    for (i in files(folder)){
  titles <- gsub(".", "", i)
    knitr::knit_child(text =c(
'', 
    '```{r evolvaluedim, fig.cap=`titles`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'knitr::include_graphics(file.path(fig.pathinside, file.path(folder, i)))',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}

  } else {
  assign("Dimension2", gsub("_", "-", unique(x$data$Dimension)))
  assign("y", x, envir = environment())
  if(difference){
    title_knit <- paste0("Differences in percent of value for temporal dimension :  ", Dimension2, " between ", titre_init, " and ",titre_final, " dataset ")
    folder = "Temporaldiff"

  } else {
  title_knit <- paste0("Evolutions of values for the dimension ", Dimension2, " for ", titre_init, " and ",titre_final, " dataset ")
  if(unique_analyse){title_knit = paste0("Evolutions of values for the dimension ", Dimension2, " for ", titre_init, " dataset ")}
  }
  save_image(title = title_knit, plott = y, folder = folder, fig.pathinside = fig.pathinside)


  knitr::knit_child(text =c(
'', 
    '```{r evolvaluedimdiff, fig.cap=`title_knit`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'y',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}
}


```




```{r}
pie_chart_2_default = function (dimension, first, second = NULL, topn = 10, titre_1 = "first", 
  titre_2 = "second", title_yes_no = TRUE, dataframe = FALSE) 
{
  first[is.na(first)] <- "NA"
  if (deparse(substitute(dimension)) == "X[[i]]") {
    r <- dimension
  }
  else {
    r <- deparse(substitute(dimension))
  }
  dimension <- gsub("\"", "", r)
  if (dimension == "source_authority") {
    topn = 6
  }
  name1 <- dplyr::enquo(titre_1)
  name2 <- dplyr::enquo(titre_2)
  all_class_i <- first %>% dplyr::group_by(across(c(dimension, 
    "measurement_unit"))) %>% dplyr::summarise(measurement_value = sum(measurement_value, 
    na.rm = TRUE)) %>% filter(measurement_value != 0) %>% 
    dplyr::select(-measurement_value)
  colnames(all_class_i) <- c("class", "measurement_unit")
  all_class_i <- all_class_i %>% mutate(class = paste(class, 
    measurement_unit, sep = " / "))
  provisoire_i <- first %>% dplyr::group_by(dplyr::across(c(dimension, 
    "measurement_unit"))) %>% dplyr::summarise(measurement_value = sum(measurement_value, 
    na.rm = TRUE)) %>% dplyr::group_by(measurement_unit) %>% 
    dplyr::arrange(desc(measurement_value)) %>% dplyr::mutate(id = row_number()) %>% 
    dplyr::mutate(class = as.factor(ifelse(id < topn, !!rlang::sym(dimension), 
      "Others"))) %>% dplyr::group_by(class, measurement_unit) %>% 
    dplyr::summarise(measurement_value = sum(measurement_value, 
      na.rm = TRUE)) %>% dplyr::ungroup() %>% dplyr::select(measurement_value, 
    class, measurement_unit) %>% dplyr::group_by(measurement_unit) %>% 
    dplyr::mutate(pourcentage = prop.table(measurement_value) * 
      100) %>% dplyr::mutate(labels = paste0(pourcentage, 
    " ", " % ")) %>% dplyr::arrange(desc(class)) %>% dplyr::mutate(ypos_ligne = cumsum(pourcentage) - 
    0.5 * pourcentage) %>% dplyr::distinct() %>% dplyr::filter(!is.na(class))
  if (!is.null(second)) {
    all_class_t <- first %>% dplyr::group_by(across(c(dimension, 
      "measurement_unit"))) %>% dplyr::summarise(measurement_value = sum(measurement_value, 
      na.rm = TRUE)) %>% filter(measurement_value != 0) %>% 
      dplyr::select(-measurement_value)
    colnames(all_class_t) <- c("class", "measurement_unit")
    all_class_t <- all_class_t %>% mutate(class = paste(class, 
      measurement_unit, sep = " / "))
    provisoire_t <- second %>% dplyr::group_by(across(c(dimension, 
      "measurement_unit"))) %>% dplyr::summarise(measurement_value = sum(measurement_value, 
      na.rm = TRUE)) %>% dplyr::group_by(measurement_unit) %>% 
      dplyr::arrange(desc(measurement_value)) %>% dplyr::mutate(id = row_number()) %>% 
      dplyr::mutate(class = as.factor(ifelse(id < topn, 
        !!rlang::sym(dimension), "Others"))) %>% dplyr::group_by(class, 
      measurement_unit) %>% dplyr::summarise(measurement_value = sum(measurement_value, 
      na.rm = TRUE)) %>% dplyr::ungroup() %>% dplyr::select(measurement_value, 
      class, measurement_unit) %>% dplyr::group_by(measurement_unit) %>% 
      dplyr::mutate(pourcentage = prop.table(measurement_value) * 
        100) %>% dplyr::mutate(labels = paste0(pourcentage, 
      " ", " % ")) %>% dplyr::arrange(desc(class)) %>% 
      dplyr::mutate(ypos_ligne = cumsum(pourcentage) - 
        0.5 * pourcentage) %>% dplyr::distinct() %>% 
      dplyr::filter(!is.na(class))
  }
  if (!is.null(second)) {
    disappearing_stratas <- anti_join(all_class_i %>% dplyr::select(class), 
      all_class_t %>% dplyr::select(class)) %>% distinct()
    appearing_stratas <- anti_join(all_class_t %>% dplyr::select(class), 
      all_class_i %>% dplyr::select(class)) %>% distinct()
    number_disappearing_stratas <- nrow(disappearing_stratas)
    number_appearing_stratas <- nrow(appearing_stratas)
    summary_apparition <- ggdraw() + draw_label(paste0("Number of appearing stratas : ", 
      number_appearing_stratas), size = 10)
    if (number_appearing_stratas != 0) 
      summary_apparition <- summary_apparition + draw_label(paste0(" \nThey are ", 
        paste((appearing_stratas %>% dplyr::select(class) %>% 
          distinct())$class, sep = ";")), size = 10)
    summary_apparition <- summary_apparition + draw_label(paste0(" \nNumber of disappearing stratas : ", 
      number_disappearing_stratas), size = 10)
    if (number_disappearing_stratas != 0) 
      summary_apparition <- summary_apparition + draw_label(paste0(" \nThey are ", 
        paste((disappearing_stratas %>% dplyr::select(class) %>% 
          distinct())$class, sep = ";")), size = 10)
  }
  set.seed(2)
  if (!(is.null(second))) {
    number <- length(unique(unlist(as.character(c(provisoire_i$class, 
      provisoire_t$class)))))
  }
  else {
    number <- length(unique(unlist(as.character(c(provisoire_i$class)))))
  }
  pal <- brewer.pal(number, "Paired")
  if (!(is.null(second))) {
    pal = setNames(pal, unique(unlist(as.character(c(provisoire_i$class, 
      provisoire_t$class)))))
  }
  else {
    pal = setNames(pal, unique(unlist(as.character(c(provisoire_i$class)))))
  }
  ggplot_i <- ggplot(provisoire_i %>% dplyr::filter(!is.na(class))) + 
    aes(x = "", fill = class, group = class, weight = pourcentage) + 
    geom_bar(position = "fill") + scale_fill_hue(direction = 1) + 
    scale_color_hue(direction = 1) + theme_minimal() + coord_polar("y", 
    start = 0) + geom_text(first = (provisoire_i %>% dplyr::filter(!is.na(class)) %>% 
    dplyr::mutate_if(is.numeric, round)), size = 3, aes(x = 1, 
    y = ypos_ligne/100, label = paste0(round(pourcentage), 
      "%")), color = "black") + theme(axis.ticks.x = element_blank(), 
    axis.text.x = element_blank()) + labs(x = "", y = "") + 
    scale_fill_manual(values = pal) + guides(fill = guide_legend(title = toupper(r))) + 
    facet_wrap("measurement_unit")
  if (!is.null(second)) {
    to_get_legend <- ggplot(rbind(provisoire_i %>% dplyr::filter(!is.na(class)), 
      provisoire_t %>% dplyr::filter(!is.na(class)))) + 
      aes(x = "", fill = class, group = class, weight = pourcentage) + 
      geom_bar(position = "fill") + guides(fill = guide_legend(title = toupper(r)))+ 
      scale_fill_manual(values = pal)
    legend <- cowplot::get_legend(to_get_legend)
    ggplot_t <- ggplot(provisoire_t %>% dplyr::filter(!is.na(class))) + 
      aes(x = "", fill = class, group = class, weight = pourcentage) + 
      geom_bar(position = "fill") + scale_fill_hue(direction = 1) + 
      scale_color_hue(direction = 1) + theme_minimal() + 
      coord_polar("y", start = 0) + geom_text(first = (provisoire_t %>% 
      dplyr::filter(!is.na(class)) %>% dplyr::mutate_if(is.numeric, 
      round)), size = 3, aes(x = 1, y = ypos_ligne/100, 
      label = paste0(round(pourcentage), "%")), color = "black") + 
      theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())  +
      labs(x = "", y = "") +  scale_fill_manual(values = pal) +
      guides(fill = guide_legend(title = toupper(r))) + 
    facet_wrap("measurement_unit") + 
      theme(legend.position = "none") 
  }
  else {
    legend <- cowplot::get_legend(ggplot_i + 
      scale_fill_manual(values = pal))
  }
  if (title_yes_no) {
    title <- ggdraw() + draw_label(paste0("Distribution in measurement_value for the dimension : ", 
      r), fontface = "bold", x = 0, hjust = 0) + theme(plot.margin = margin(0, 
      0, 0, 7))
  }
  else {
    title <- ggdraw() + draw_label(" \n ")
  }
  if (!is.null(second)) {
    graph <- plot_grid(ggplot_i + theme(legend.position = "none"), 
      ggplot_t, nrow = 2, labels = c(gsub("\"", "", gsub("~\"", 
        "", deparse(substitute(name1)))), gsub("\"", 
        "", gsub("~\"", "", deparse(substitute(name2))))), 
      label_size = 10, vjust = 1.3, label_x = c(0, 0), 
      label_y = 1.025, axis = "l", align = "v")
    ploting_map <- plot_grid(title, nrow = 2, plot_grid(graph, 
      legend, ncol = 2), rel_heights = c(0.1, 1)) + theme(plot.background = element_rect(color = "black"))
    if (sum(!(round(provisoire_i$pourcentage) == round(provisoire_t$pourcentage))) == 
      0) {
      title <- ggdraw() + draw_label(paste0("Distribution in measurement_value for the dimension : ", 
        r, "\n(same distribution to the nearest rounding for both datasets : \n", 
        gsub("\"", "", gsub("~\"", "", deparse(substitute(name1)))), 
        " and \n", gsub("\"", "", gsub("~\"", "", deparse(substitute(name2)))), 
        ")"), fontface = "bold", x = 0, hjust = 0, vjust = 0.5, 
        size = 13) + theme(plot.margin = margin(0, 0, 
        0, 7))
      graph <- ggplot_i + theme(legend.position = "none")
    }
  }
  else {
    graph <- plot_grid(ggplot_i + theme(legend.position = "none"), 
      nrow = 1, labels = c(gsub("\"", "", gsub("~\"", 
        "", deparse(substitute(name1))))), label_size = 10, 
      vjust = 1.3, label_x = c(0, 0), label_y = 0.8, axis = "l", 
      align = "v")
  }
  if (title_yes_no) {
    if (exists("provisoire_t")) 
      if (sum(!(round(provisoire_i$pourcentage) == round(provisoire_t$pourcentage))) == 
        0) {
        title <- ggdraw() + draw_label(paste0("(same distribution to the nearest rounding for both datasets :\n", 
          gsub("\"", "", gsub("~\"", "", deparse(substitute(name1)))), 
          " and ", gsub("\"", "", gsub("~\"", "", deparse(substitute(name2)))), 
          ")"), fontface = "bold", x = 0, hjust = 0, 
          vjust = 0.5, size = 13) + theme(plot.margin = margin(0, 
          0, 0, 7))
      }
      else {
        title <- ggdraw() + draw_label(" \n ")
      }
  }
  ploting_map <- plot_grid(title, nrow = 2, plot_grid(graph, 
    legend, ncol = 2), rel_heights = c(0.1, 1)) + theme(plot.background = element_rect(color = "black"))
  if (exists("summary_apparition") & dataframe) {
    df <- data.frame(` ` = c("Stratas appearing", "Stratas disappearing"), 
      Number = c(number_appearing_stratas, number_disappearing_stratas), 
      Detail = c(toString(paste((appearing_stratas %>% 
        dplyr::select(class) %>% mutate(class = gsub(" ", 
        "", class)) %>% distinct())$class, sep = ";")), 
        toString(paste((disappearing_stratas %>% dplyr::select(class) %>% 
          mutate(class = gsub(" ", "", class)) %>% distinct())$class, 
          sep = ";"))), check.names = FALSE, fix.empty.names = FALSE)
    if (number_disappearing_stratas == 0 & number_appearing_stratas == 
      0) {
      df <- df %>% dplyr::select(-Detail)
    }
    list_df_plot <- list(plot = ploting_map, df = df)
    return(list_df_plot)
  }
  else {
    return(ploting_map)
  }
}
```

```{r functionprintingmaps}

fonction_empreinte_spatiale = function(variable_affichee, initial_dataset = init, final_dataset = final, titre_1 = "Dataset 1", titre_2 = "Dataset 2"){
  selection = function(x){x %>% dplyr::ungroup() %>% dplyr::select(geographic_identifier, measurement_value, GRIDTYPE, measurement_unit)}
  Initial_dataframe <-selection(initial_dataset)
  Final_dataframe <-selection(final_dataset)
  geo_data <- gdata::combine(`Initial_dataframe`, `Final_dataframe`)
  rm(`Initial_dataframe`, `Final_dataframe`)
  gc()
  geo_data <- geo_data %>% dplyr::mutate(source = dplyr::case_when(source == "Initial_dataframe"~  eval(parse(text = "titre_1")),source == "Final_dataframe"~ eval(parse(text = "titre_2")), TRUE ~ "Error"))
  inner_join <- st_as_sf(geo_data%>% dplyr::group_by(geographic_identifier, measurement_unit,source)  %>%  dplyr::summarise(measurement_value = sum(measurement_value, na.rm = TRUE))  %>% dplyr::filter(measurement_value != 0) %>% dplyr::inner_join(shapefile.fix, by = c("geographic_identifier"="CWP_CODE")))
  
  if(nrow(inner_join %>% dplyr::filter(measurement_unit == variable_affichee)) != 0){
     if (plotting_type == "view"){image <- tm_shape(inner_join %>% dplyr::filter(measurement_unit == variable_affichee))+
      tm_fill("measurement_value", palette="RdYlGn", style="cont", n=8,
              id="name", midpoint = 0) +
      tm_layout(legend.outside = FALSE) + tm_facets(by=c("GRIDTYPE","source"), free.scales = TRUE)} else {image <- tm_shape(inner_join %>% dplyr::filter(measurement_unit == variable_affichee))+
        tm_fill("measurement_value", palette="RdYlGn", style="cont", n=8,
                id="name", midpoint = 0) +
        tm_layout(legend.outside = FALSE) + tm_facets(by=c("GRIDTYPE","source"), free.scales = TRUE)+tm_shape(continent)+tm_borders()}
    
    return(list(image = image, variable_affichee = variable_affichee))
  }
}

map_unit_knit = function(map, find_and_print = FALSE,folder = "Geodistrib", fig.pathinside = fig.path){
  if(find_and_print){
    for (i in files(folder)){
  titles <- gsub(".", "", i)
    knitr::knit_child(text =c(
'', 
    '```{r distributioninvalueforunit, fig.cap=`titles`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            'knitr::include_graphics(file.path(fig.pathinside, file.path(folder, i)))',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}

  } else {
  # if (deparse(substitute(measurement_unit)) == "X[[i]]"){ #for sapply function bug
  # assign("r", measurement_unit, envir = environment())
  # }else {  assign("r", substitute(measurement_unit), envir = environment())}
  assign("unit_name_map", map[2][[1]], envir = environment())
  assign("map_for_knit", map[1][[1]], envir = environment())
  dimension_title <- gsub("_","..",unit_name_map)
save_image(title = paste0("Distribution in value for the unit : ",(unit_name_map)),plott = map_for_knit,  folder = folder, fig.pathinside = fig.pathinside)
  
  knitr::knit_child(text = c(
    '',
    '',
    '```{r}', 
    'dimension_title <- gsub("_","..",unit_name_map)',
    '```',
    '',
    '```{r distributioninvalueforunit, fig.cap = paste0("Distribution in value for the unit : ",(unit_name_map)), out.width = "100%"}',
    '',
    'map_for_knit',
    '```',
    '',
    ''
 ), envir = environment(), quiet= TRUE)
  }
    }
```





