---
title: "`r ifelse(exists('title_markdown'), title_markdown, 'Comparison of two CWP format datasets')`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Emmanuel Blondel
    affiliation: FAO
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
csl: dmk-format.csl
link-citations: TRUE
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
        scroll_highlight: true
        before: null
        after: null
  bookdown::pdf_document2:
    latex_engine: lualatex 
    extra_dependencies: ["adjustbox", "subfig", "flafter", "makecell"]
    toc: TRUE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
    always_allow_html: TRUE
    keep_md: TRUE
abstract: "This document provides an overview of the impact of various processing steps on a specific dataset from the Global Tuna Atlas. We begin by presenting the final dataset along with its key characteristics, followed by a global comparison with the FIRMS level 0 dataset. Next, we outline each step involved in creating the final dataset, detailing the decisions made and their effects on specific data points.\n If you would like to explore the impact of individual processing steps with more specific filters, feel free to contact us or rerun the process and automated reporting using the project available on GitHub at firms-gta/geoflow-tunaatlas."
geometry: "margin=20mm"
papersize: a4
endfloat: FALSE
numberlines: FALSE
authblk: FALSE
footnotehyper: FALSE
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE
tables: TRUE
fancyhdr:
  first:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 0pt
  subsequent:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 1pt
    footrulewidth: 0pt
---


```{r setuprmd,include = FALSE, echo=FALSE}

res <- knitr::knit_child("Setup_markdown.Rmd", quiet = TRUE, envir = environment())
# envir = parent.frame() ne marche pas bien car il faut appeler dans le nouveau markdown (child) envir$obj
```

```{r}

if(!exists("parameter_short")){parameter_short = FALSE}
if(!exists("child_header")){child_header = ""}

```


```{r eval=exists("Add_lines") && file.exists(Add_lines), results='asis'}

res <- knitr::knit_child(Add_lines, quiet = TRUE, envir = environment())
cat(res)

```

`r if(exists("step_title_t_f") && step_title_t_f){cat_title(step_title, child_headerinside = child_header)}`

```{r explanationongainandloss, results='asis', eval=(!parameter_short && !unique_analyse)}

  cat(
    "*In the following document:* \n",
    "- ***All the differences inferior to 0 corresponds to gain in captures.*** \n", 
    "- ***The initial dataset, sometimes referred as, dataset 1 is: ", parameter_titre_dataset_1, "*** \n",
    "- ***The final dataset, sometimes referred as, dataset 2 is: ", parameter_titre_dataset_2, "***", 
    sep = "\n"
  )

```

```{r filterprinting, echo=FALSE, results='asis'}

if (!isNullList(parameter_filtering)){
  cat (paste0("The filter used on this data are:  \n "))
  for (i in 1:length(parameter_filtering)){
    if (parameter_filtering[i] %in% parameter_filtering & !is.null(parameter_filtering[[i]])){
      cat(paste0("***- On ", names((parameter_filtering)[i]), " : ", paste((parameter_filtering)[[i]], collapse = " ; "),"*** \n"))
    } else {""}
  }
}

if (!is.null(parameter_resolution_filter)) {

  cat (paste0("The geographical filter used on this data is:  \n "))

  cat(parameter_resolution_filter)
    
  }


```

```{r include=FALSE}

if(!unique_analyse && parameter_short == FALSE && coverage){
  main_title <- TRUE
} else {
  main_title <- FALSE
  child_header <- sub("#", "", child_header)
}

```

`r if(main_title){cat_title("# Main differences", child_headerinside = child_header)}`

```{r summarydiffplotting, eval=!unique_analyse,echo=FALSE, results='asis', out.width="100%"}

qflextable2(summary_of_differences, captionn = "Summary of the difference between the two datasets", columns_to_color = c("Difference"), save_folder = "Summary", interactive_plot = FALSE)

```

```{r include=FALSE, eval=!unique_analyse}

title <- gsub("_", "\\_", combined_summary_histogram$labels$title, fixed = TRUE)

combined_summary_histogram$labels$title <- ""

```


```{r, echo=FALSE, out.width="100%", warning=FALSE, message=FALSE, fig.cap=title, eval=!unique_analyse}

combined_summary_histogram

```

```{r diffstratas,echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis', eval=!unique_analyse}

if (nrow(compare_strata_differences_list$strates_perdues_first_10) != 0) {
  cat("The strata differences (completely lost or appearing) between the first one and the second one (representing ", round(compare_strata_differences_list$pourcentage_strates_perdues), " % of the total number of strata) are :")
} else {
  cat("No stratum is gained nor lost")
}
```

```{r disapandapflextable, eval = (nrow(compare_strata_differences_list$strates_perdues_first_10) != 0) & !unique_analyse, results="asis"}

qflextable2(compare_strata_differences_list$disapandap, captionn = compare_strata_differences_list$title, columns_to_color = c("Difference in millions"), grouped_data = c("Dimension", "measurement_unit", "Loss / Gain"), save_folder = "Diffstratas", fig.pathinside = fig.path, interactive_plot = FALSE, find_and_print = FALSE)

```

```{r compnumberstratascolumns, results='asis', out.width="100%", eval=!unique_analyse}

qflextable2(compare_strata_differences_list$number_init_column_final_column, captionn = "Comparison of number of stratas between the two datasets", columns_to_color = c("Difference"), save_folder = "Diffstratas", fig.pathinside = fig.path, interactive_plot = FALSE)

```

```{r parametershort,  eval=parameter_short && !unique_analyse, results='asis'}

exit <- "This document only presents the summary of the differences between the two datasets. If you need a more detailed summary please provide FALSE as argument for the 'short' parameter"

if(child){exit <- ""}

if (parameter_short){knitr::knit_exit(exit, fully = FALSE)}


```


`r if(!unique_analyse && (exists("plot_titles_list") && !is.null(plot_titles_list)) && !=unique_analyse){cat_title("## Differences in temporal data", child_headerinside = child_header)}`

```{r include = FALSE,eval = !unique_analyse && (exists("plot_titles_list") && !is.null(plot_titles_list))}

timediffplot_knit_list <- mapply(FUN = knitting_plots_subfigures, plot = plot_titles_list$plots, title = plot_titles_list$titles, MoreArgs = list(folder = "Temporal", fig.pathinside = fig.path))

```


```{r timedimdiffmultipleplotsprinting, results='asis',eval = !unique_analyse && (exists("plot_titles_list") && !is.null(plot_titles_list))}

cat(unlist(timediffplot_knit_list), sep = "\n")

```

\clearpage

`r if(!unique_analyse && (exists("Geographicdiff") && !is.null(Geographicdiff))){cat_title("## Differences in geographical data", child_headerinside = child_header)}`

```{r tmapsaving, message=FALSE, warning=FALSE, eval = !unique_analyse && (exists("Geographicdiff") && !is.null(Geographicdiff))}

save_image(title = Geographicdiff$title, plott = Geographicdiff$plott, folder = "Spatialdiffmap", find_and_print = outputonly, fig.pathinside = fig.path)

```


```{r spatialdifferencesplotting, out.width="100%", eval = !unique_analyse && (exists("Geographicdiff") && !is.null(Geographicdiff))}

Geographicdiff$plott

```

\clearpage

`r if(!unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))){cat_title("## The differences for each other dimension", child_headerinside = child_header)}`

```{r topn, include=FALSE, eval = !unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))}
if(exists("compare_dimension_differences_list$topn")){topn <- compare_dimension_differences_list$topn} else{topn <- 6}
```


```{r topnsentence, eval = !unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))}

cat(paste0("We present for each dimension the ", topn, " most important differences without presenting the stratas completely appearing or disappearing."))

```



```{r printing-table-diff-dimensions, paged.print=TRUE, results='asis', eval = !unique_analyse && (exists("compare_dimension_differences_list") && !is.null(compare_dimension_differences_list))}
if(nrow(compare_dimension_differences_list$Groupped_all_not_disap_or_app_to_dysplay) != 0) {
  qflextable2(compare_dimension_differences_list$Groupped_all_not_disap_or_app_to_dysplay, captionn = compare_dimension_differences_list$title, grouped_data = c("Dimension", "measurement_unit", "Loss / Gain"), save_folder = "Diffstratas", columns_to_color = c("Difference (in %)"))
} else {
  cat("There are no differences between stratas aside the appearing and disappearing ones")
}
```


```{r eval = !coverage}

knitr::knit_exit(fully = FALSE)

```

```{r Maincharactofdataset, results='asis', eval = coverage}

res <- knitr::knit_child("Main_caracteristics_of_dataset.Rmd", quiet = TRUE, envir = environment())

cat(res, sep =  "\\clearpage")

```
