---
title: "SetupMarkdown"
date: "2023-07-19"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  error = TRUE
)

```



```{r utility_functions, include=FALSE}
`%notin%` <- Negate(`%in%`)

last_path = function(x) {
  substr(x, max(gregexpr("/", x)[[1]]) + 1, nchar(x))
}

last_path_reduced = function(x) {
  gsub("georef_dataset", "", last_path(x))
}

is_null_or_not_exist = function(x) {
  if (!exists(deparse(substitute(x))) || is.null(x)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

```{r GTAMARKDOWNFUNCTIONS}

fSpatPlan_Convert2PacificRobinson = function(df, buff = 0){
  
  if(!exists("df_robinson")){
    df_robinson <- df}
  return(df_robinson)
  
}

knit_child_map =function(list_map_unit){
  assign("unit_name_map", list_map_unit[2][[1]], envir = environment())
  assign("map_for_knit", list_map_unit[1][[1]], envir = environment())
  
  knit_child(text = c(
    '```{r results = "asis"}',
    '',
    'unit_title <- gsub("_","..",unit_name_map)',
    '```',
    '',
    '    ```{r setup, include=FALSE}',
'knitr::opts_chunk$set(label = paste0("mapdiff", unit_title))', 
'```',
'', 
'', 

    '',
    '```{r results="asis", fig.cap = paste0("Map of the differences between two datasets for the unit ", unit_title)}',
    '',
    'map_for_knit',
    '',
    '```',
    '',
    '',''),
    envir = environment(), quiet= TRUE)
}


qflextable2 <- function(x, captionn = NULL, autonumm = autonum, pgwidth = 6, columns_to_color = NULL, save_folder = NULL, fig.pathinside = fig.path, grouped_data = NULL) {
  if (all(class(x) == "flextable")) {
    flextabley <- x
  } else {
    if (!(all(class(x) == "data.frame"))) {
      x <- as.data.frame(x %>% dplyr::ungroup())
    }
    if (!is.null(save_folder)) {
      if (!dir.exists(paste0(fig.pathinside, save_folder))) {
        dir.create(paste0(fig.pathinside, save_folder))
      }
      save_path_data <- file.path(fig.pathinside, save_folder, paste0(make.names(captionn), ".csv" ))  # Adjust the file name as needed
      fwrite(x, file = save_path_data)
    }

    y <- x %>% 
      dplyr::ungroup() %>% 
      dplyr::mutate_if(is.factor, as.character) %>% 
      dplyr::mutate_if(is.character, ~ str_replace_all(., ",", ", \n" )) %>% 
      dplyr::mutate_if(is.character, ~ str_replace_all(., "_", "-" )) %>% 
      dplyr::mutate_if(is.numeric, function(.) { round(., 2) }) 

    if(!is.null(grouped_data)){
    y <- as_grouped_data(y, groups = grouped_data)
  }

if (!is.null(columns_to_color)) {
    colormatrix <- ifelse(y %>% dplyr::select(all_of(columns_to_color)) < 0, "red", 
                           ifelse(y %>% dplyr::select(all_of(columns_to_color)) > 0, "green", "white"))
    flextabley <- flextable::flextable(y)
    
     flextabley <- flextabley %>% highlight(j = all_of(columns_to_color), color = colormatrix)
    

    } else {    flextabley <- flextable::flextable(y)}
    
}

if (!is.null(captionn)) {
    flextable_captionned <- flextable::set_caption(flextabley, caption = captionn, style = "Table Caption")
  } else {
    flextable_captionned <- flextabley
  }
  

  
  ft_out <- flextable_captionned %>% autofit()
  ft_out <- width(ft_out, width = dim(ft_out)$widths * pgwidth / flextable_dim(ft_out)$widths)
  
  if (!is.null(save_folder)) {
    save_path_flextable <- file.path(fig.pathinside, save_folder, paste0(make.names(captionn), ".png"))  # Adjust the file name as needed
    save_as_image(ft_out, path = save_path_flextable)
  }
  
  return(ft_out)
}

filtering_function= function(dataframe_to_filter, parameter_filtering, geo_dim = "geographic_identifier", parameter_fact = "catch", parameter_UNK_for_not_standards_unit = TRUE){
  
  matchingList <- parameter_filtering %>% purrr::keep( ~ !is.null(.) )


  
  if(length(matchingList)!= 0){  
    
        colnames_to_filter <- colnames(dataframe_to_filter %>% select(names(matchingList)))

  names(matchingList) <- colnames_to_filter

  matchingList <- lapply(matchingList, function(x){ #handling single filter
  if(length(x) == 1){
  x <- c(x, x) }else {
    x
  }

}
  )

    
    dataframe_to_filter <- dataframe_to_filter%>% dplyr::filter(!! rlang::parse_expr(str_c(colnames_to_filter, matchingList, sep = '%in%', collapse="&")))}
  
  if(  parameter_UNK_for_not_standards_unit & parameter_fact == "effort"){
    dataframe_to_filter <- dataframe_to_filter %>% dplyr::mutate(measurement_unit = ifelse(measurement_unit%in%c("HOOKS","FDAYS"), measurement_unit, "UNK" ))}
  
  if(geo_dim != "geographic_identifier" && "geographic_identifier"%in%colnames(dataframe_to_filter)){
    dataframe_to_filter <- dataframe_to_filter %>% dplyr::rename("geographic_identifier" := {{geo_dim}})
  }
  return(dataframe_to_filter)
}

tidying_data_for_GTA = function(dataframe, time_dim, shape, species_group_dataframe = species_group, cl_cwp_gear_level2_dataframe = cl_cwp_gear_level2){
  if("geographic_identifier"%in%colnames(dataframe)){
    dataframe <- dataframe%>%  dplyr::left_join(shape, by = c("geographic_identifier"="CWP_CODE"))
    print_map <- TRUE
  } else {print_map <- FALSE}
  
  
  dataframe <- dataframe %>% mutate_at(all_of(time_dim), as.character)
  
  if("GRIDTYPE"%in%colnames(dataframe)){
    dataframe <- dataframe%>%dplyr::mutate(GRIDTYPE = as.character(GRIDTYPE))
  }
  
  dataframe <- dataframe %>% dplyr::left_join(species_group_dataframe%>% dplyr::distinct(), by = c("species"))
  if("gear_type" %in%colnames(dataframe) ){
    dataframe <- dataframe %>% dplyr::left_join(cl_cwp_gear_level2_dataframe, by = c("gear_type" = "Code"))
  }
  
  
  dataframe <- dataframe%>%dplyr::mutate(measurement_unit = dplyr::case_when(measurement_unit %in% c("MT","t","MTNO", "Tons")~ "Tons", measurement_unit %in% c("NO", "NOMT","no", "Number of fish")~"Number of fish", TRUE ~ measurement_unit)) 
  
  return(list(dataframe = dataframe, print_map = print_map))
}


fonction_groupement = function(these_col, init, final){
  
  # Compute sum of values for each combination of the columns in "these_col" and "measurement_unit" in the "init" dataframe
  groupement_1 <- init %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!!rlang::syms(these_col), measurement_unit) %>%
    dplyr::summarise(value_sum_1 = ((sum(measurement_value, na.rm=TRUE)))) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(value_sum_1 = dplyr::coalesce(value_sum_1, 0)) %>%
    dplyr::mutate(measurement_unit = as.character(measurement_unit)) %>%
    dplyr::mutate(number_lines1 = n())
  
  # Compute sum of values for each combination of the columns in "these_col" and "measurement_unit" in the "final" dataframe
  groupement_2 <- final %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!!rlang::syms(these_col), measurement_unit) %>%
    dplyr::summarise(value_sum_2 = ((sum(measurement_value, na.rm=TRUE)))) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    dplyr::mutate(measurement_unit = as.character(measurement_unit)) %>%
    dplyr::mutate(number_lines2 = n())
  
  # Join the two dataframes based on the columns in "these_col" and "measurement_unit"
  fulljoin <- dplyr::full_join(groupement_1, groupement_2) %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    dplyr::mutate(value_sum_1 = dplyr::coalesce(value_sum_1, 0)) %>%
    dplyr::mutate(loss = value_sum_1 - value_sum_2) %>%
    # Compute whether the difference in value is a gain or loss
    dplyr::mutate(`Loss / Gain` = dplyr::case_when(abs(loss) <= 1 ~ "Egal", loss > 1 ~"Loss",  loss < 1 ~"Gain")) %>%
    dplyr::mutate(Loss_pourcent = - (100*((value_sum_1 - value_sum_2)/value_sum_1))) %>%
    dplyr::mutate(Dimension = colnames(groupement_1[1])) %>%
    dplyr::rename("Precision" = 1) %>%
    dplyr::mutate(Precision = as.character(Precision)) %>%
    dplyr::mutate(value_sum_2 = dplyr::coalesce(value_sum_2, 0)) %>%
    # Set Loss_pourcent to 100 if it is NA or -Inf or 0 
    dplyr::mutate(Loss_pourcent = base::ifelse(is.na(Loss_pourcent)|Loss_pourcent==-Inf, 100, Loss_pourcent)) %>%
    dplyr::ungroup() %>%
    # Compute the difference in the number of lines between the two dataframes
    dplyr::mutate(loss_nb_ligne = - (number_lines1 - number_lines2)) %>%
    dplyr::mutate(`Difference in value`= - (value_sum_1 - value_sum_2)) %>% #final - init
    dplyr::rename(`Difference (in %)` = Loss_pourcent,`Difference in number of lines` = loss_nb_ligne)%>%
    dplyr::mutate_if(is.numeric, list(~replace_na(., 0)))
  
  return(fulljoin)
}

save_image = function(title, plott = last_plot(), folder = NULL, fig.pathinside = fig.path){
  if(!is.null(folder) && !dir.exists(folder)){
    dir.create(paste0(fig.path, folder))
  }
  
    if(all(class(plott) == "flextable")){
      save_as_image(plott, title, path = paste0(fig.pathinside, folder))
      

    }

ggsave(paste0( make.names(title), ".png"),plot = plott,   device = "png", path = paste0(fig.pathinside, folder))
ggsave(paste0( make.names(title), ".pdf"), plot = plott,  device = "pdf", path = paste0(fig.pathinside, folder))

}


resume_knit_child = function(x){
  dimension_title <- gsub("_", "-", unique(x$dimension))

  title_knit <- paste0("Distribution in value for the dimension : ",(dimension_title))
  assign("dimension_title", gsub("_", "-", unique(x$dimension)), envir = environment())
save_image(title_knit, x$pie_chart, folder = "Piechartsdimensions")
  assign("x", x, envir = environment())

  knitr::knit_child(text = c(

    '',
    '',
'', 

    '',
    '```{r distribinvaluedim, results = "asis" ,fig.cap =`title_knit` }',
    '',
    'x$pie_chart',
'',
'```',
    '',
    '',
    '',
    '```{r results = "asis", eval= !is.null(x$df)}',
    'x$df',
    '',
    '',
    '```',
    '',
    '', 
    ''
    
    
    
    
  ), envir = environment(), quiet= TRUE)
}
function_pie_chart_df = function(dimension, ...){
  formals(pie_chart_2_default)$dimension <- dimension
  dimension_returned <- dimension
  pie_chart = pie_chart_2_default(...)
  if(exists("pie_chart$df")){
    df = pie_chart$df
    pie_chart = pie_chart$plot} else{
      pie_chart = pie_chart
      df = NULL}
  
  if(!is.null(df)){
    df <- qflextable2(df, captionn = paste0("Strata movement for the dimension : ", dimension_returned))}
  return(list(pie_chart=pie_chart,df=df, dimension=dimension_returned))
}


function_knitting = function(x){
  assign("Dimension2", gsub("_", "-", unique(x$data$Dimension)))
  assign("y", x, envir = environment())
  title_knit <- paste0("Evolutions of values for the dimension ", Dimension2, " for ", `titre_1`, " and ",`titre_2`, " dataset ")
  if(nrow(final)==0){title_knit = paste0("Evolutions of values for the dimension ", Dimension2, " for ", `titre_1`, " dataset ")}
  save_image(title_knit,x,  folder = "Temporal")


  knitr::knit_child(text =c(
'', 
    '```{r evolvaluedim, fig.cap=`title_knit`, fig.align = "center", out.widht = "100%"}',
                            '',

                            '',
                            '(y)',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}
function_knitting_difference = function(x){

  assign("Dimension2", gsub("_", "-", unique(x$data$Dimension)))
  title_knit <- paste0("Differences in percent of value for temporal dimension :  ", Dimension2, " between ", `titre_1`, " and ",`titre_2`, " dataset ")
save_image(title_knit,x,  folder = "Temporaldiff")


  assign("y", x, envir = environment())
  knitr::knit_child(text =c(

    '```{r difftempdim, fig.cap=`title_knit`}',
                            '',

                            
                            '',
                            '(y)',
                            '',
                            '```'), envir = environment(), quiet= TRUE)
}

```


```{r libraries,message=FALSE, warning=FALSE, include=FALSE}



# Prevent package installation from upgrading
formals(install.packages)$upgrade <- "never"

# List of required packages
required_packages <- c("dplyr", "knitr", "stringr", "purrr", "readxl", "base", "flextable", "remotes", 
                        "utils", "DBI", "odbc", "rlang", "sf", "kableExtra", "readr", "tidyr", "ggplot2", 
                        "stats", "RColorBrewer", "cowplot", "tmap", "RPostgreSQL", "officer", "gdata")

# Check if each package is already installed, and if not, install and load it
for (package in required_packages) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package)
    require(package)
  }
}

# Install and load sutdycatchesird package from GitHub if not already installed
if(!require(sutdycatchesird)){
  remotes::install_github("BastienIRD/sutdycatchesird", upgrade = "never")
  require(sutdycatchesird)
}

```


```{r message=FALSE, warning=FALSE, include=FALSE}
set_flextable_defaults(
  font.size = 10,
  font.color = "black",
  digits = 2,
  theme_fun = "theme_box"
)


#set the dataset caption styling
knitr::opts_chunk$set(tab.cap.pre = "Table ", tab.cap.sep = ": ")

#set the dataset caption styling
autonum <- officer::run_autonum(seq_id = "tab", bkm = "TC1", bkm_all = TRUE) # number the table, bkm (bookmark) is important as the cross-referencing is done using the bookmark


```



```{r options, include=FALSE}
base::options(scipen=9999)
plotting_type <- "plot" 
if (knitr::is_html_output()){plotting_type <- "view" }
```
