---
title: 'Analysis of the data georeferenced not declared in nominal'
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Emmanuel Blondel
    affiliation: FAO
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
  - name: Paul Taconet 
    affiliation: IRD
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
# bibliography: ["IOTC-2022-WPDCS18.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::html_document2:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter"]
    toc: FALSE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
  bookdown::word_document2: 
    # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: FALSE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
fancyhdr:
  first:
    headleft: ""
    headright: ""
    headrulewidth: 0pt
    #footleft: A left foot
    footrulewidth: 0pt
  subsequent:
    headleft: ""
    headright: ""
    headrulewidth: 1pt
    footrulewidth: 0pt
---

```{r setuprmd,include = FALSE}

res <- knitr::knit_child("Setup_markdown.Rmd", quiet = TRUE, envir = environment())

```

```{r sourcingfunctionsmarkdown,include = FALSE}

source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/Developement/Analysis_markdown/functions/Functions_markdown.R", local = TRUE)

```

For the remaining of the analysis we only keep Dimension present in both dataset. 

# General overview of this inconsistent data 

The measurement_value of catches in the georeferenced dataset but not in nominal is different depending on what we take as a strata. The strata kept in this report are : **`r strata`**, the corresponding measurement_values are : `r round(sum(georef_no_nominal_groupped$measurement_value)/1000000,2)` millions, which corresponds of a total of  `r round(100*sum(georef_no_nominal_groupped$measurement_value)/sum(nominal_groupped$measurement_value),2)` % of the nominal catch.


```{r}
if(nrow(georef_no_nominal_groupped)==0){parameter_georef_no_nom = FALSE} else {parameter_georef_no_nom= TRUE}
```


```{r georefnotnominalgrouppedbyyear , eval=parameter_georef_no_nom}

ggplot(georef_no_nominal_groupped %>% 
         mutate(year = lubridate::year(year))) +
 aes(x = year, y = measurement_value, fill = species, colour = species) +
 geom_col() +
 scale_fill_hue(direction = 1) +
 scale_color_hue(direction = 1) +
  labs(x = "Year", y = "Value (in tons)") +

 theme_minimal() +
 facet_wrap(vars(species_group), scales = "free")

```


```{r eval = parameter_georef_no_nom}

georef_no_nominal_groupped$GRIDTYPE <- "GRIDTYPE"
child_env <- new.env()

child_env$parameter_titre_dataset_1 <- "Georeferenced but not nominal data"
child_env$parameter_init <- georef_no_nominal_groupped
child_env$parameter_final <- NULL
child_env$parameter_colnames_to_keep <- c( "species",  "source_authority", "year", "measurement_unit", "measurement_value", "Gear", "species_group")
child_env$parameter_fact <- "catch"
child_env$parameter_geographical_dimension <- "geographic_identifier"
child_env$parameter_geographical_dimension_groupping <- "GRIDTYPE"
child_env$parameter_time_dimension <- "year"
child_env$shapefile.fix <- shapefile.fix
child_env$child_header <- ""

# rmarkdown::render("comparison.Rmd", envir = child_env, quiet = TRUE)
res2 <- knitr::knit_child("comparison.Rmd", envir = child_env, quiet = TRUE)

```

```{r eval=parameter_georef_no_nom , echo=FALSE, results='asis'}

cat(res2, sep =  "\\clearpage")

```

## Stratas with declaration superior in georeferenced than in nominal 

**For this analysis, we will only compare data with superior value in georeferenced than in nominal by `r strata`**

The species concerned by declaration (on a year) of georeferenced captures superior to nominal are : `r paste0(unique(georef_sup_to_nom$species), sep = " ; ")`, for a total (species of) `r length(unique(georef_sup_to_nom$species))`.

```{r}
table2 <- georef_sup_to_nom %>% mutate_if(is.numeric, round)%>% arrange(desc(Difference) ) %>% filter(Difference != 0)
```

```{r strataswithpostivediff, eval=TRUE}

qflextable2(table2, captionn = "Stratas of georeferenced dataset being superior to nominal dataset for a specific year and species (only data superior to 1 ton are presented)")

```

```{r sumpositivesdiffeveryyear}

qflextable2(georef_sup_to_nom %>% group_by(species, source_authority) %>% summarise(Difference = sum(Difference))%>% mutate_if(is.numeric, round)%>% arrange(desc(Difference) ) %>% filter(Difference != 0), captionn = "Sum of positive differences in every year between georeferenced and nominal")

```

```{r}

ggplotgeoref_sup_to_nom <- ggplot(georef_sup_to_nom) +
 aes(x = year, y = Difference, fill = species, colour = species) +
 geom_col() +
 scale_fill_hue(direction = 1) +
 scale_color_hue(direction = 1) +
 theme_minimal() +
 facet_wrap(vars(species_group), scales = "free")

# save_image(title = "Plot of georeferenced dataset being superior to nominal dataset for a specific year and species", ggplotgeoref_sup_to_nom)

```


```{r plotgeorefsuptonom, fig.cap = "Plot of georeferenced dataset being superior to nominal dataset for a specific year and species"}

ggplotgeoref_sup_to_nom


```

