---
title: 'Stratas in georeferenced data but not in nominal'
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Emmanuel Blondel
    affiliation: FAO
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
  - name: Paul Taconet 
    affiliation: IRD
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
# bibliography: ["IOTC-2022-WPDCS18.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::html_document2:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter"]
    toc: FALSE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
  bookdown::word_document2: 
    # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: FALSE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
fancyhdr:
  first:
    headleft: "Global Tuna Atlas comparison between two datasets"
    headright: "GTA"
    headrulewidth: 0pt
    #footleft: A left foot
    footrulewidth: 0pt
  subsequent:
    headleft: "Global Tuna Atlas comparison between two datasets"
    headright: "GTA"
    headrulewidth: 1pt
    footrulewidth: 0pt
---

```{r child = 'setup_markdown.Rmd'}

```

```{r}

url_mapping_asfis_rfmo = "https://raw.githubusercontent.com/fdiwg/fdi-mappings/main/cross-term/codelist_mapping_source_authority_species.csv"
species_to_be_kept_by_rfmo_in_level0 <- readr::read_csv(url_mapping_asfis_rfmo)

is_null_or_not_exist = function(x) {
  if (!exists(deparse(substitute(x))) || is.null(x)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}


```

```{r eval = TRUE, include=FALSE}

opts <- entity$data$actions[[1]]$options
con <- config$software$output$dbi
parameter_fact <-  opts$fact
if(!is.null(opts$filtering)){parameter_filtering <-opts$filtering} else{ parameter_filtering <-list(species = NULL, fishing_fleet = NULL)}
if(is.character(parameter_filtering)){
  parameter_filtering <- eval(parse(text=toString(parameter_filtering)))
}

parameters_child <- list(config = config, entity = entity,opts = opts ,action = action, parameter_filtering = parameter_filtering, parameter_fact = parameter_fact, parameter_con = con,parameter_host =config$software$input$dbi_config$parameters$host, parameter_port = config$software$input$dbi_config$parameters$port,  parameter_user = config$software$input$dbi_config$parameters$user,parameter_dbname=config$software$input$dbi_config$parameters$dbname, parameter_password = config$software$input$dbi_config$parameters$password,  child = TRUE, no_recap_options = TRUE)
child_env = new.env()
list2env(parameters_child, env = child_env)

```



```{r echo=FALSE, results='asis'}

isNullList = function(x) all(!lengths(x))
if (!exists("parameter_filtering") || isNullList(parameter_filtering)){cat("There are no filter on this data in the following report.")
  parameter_filtering <- list(species = NULL, source_authority = NULL)} else {
  
  cat (paste0("The filter used on the data in the following report are:  \n "))
  for (i in 1:length(parameter_filtering)){
    if (!is.null(parameter_filtering[[i]])){
      cat(paste0("***- On ", names((parameter_filtering)[i]), " : ", paste((parameter_filtering)[[i]], collapse = " ; "),"*** \n"))
    } else {""}
  }
}



```



# Analysis of the data georeferenced not declared in nominal

```{r}

nominal <- read_csv("data/global_nominal_catch_firms_level0.csv") %>% mutate(measurement_unit ="t")
georef_mapped <- readRDS("Markdown/mapping_codelist/rds.rds")

georef_mapped <- georef_mapped%>% semi_join(species_to_be_kept_by_rfmo_in_level0, by = c("species", "source_authority") )
nominal <- nominal%>% semi_join(species_to_be_kept_by_rfmo_in_level0, by = c("species", "source_authority") )

formals(filtering_function)$parameter_filtering = parameter_filtering

georef_mapped <- filtering_function(georef_mapped) %>% mutate(year = lubridate::year(time_start)) %>% select("fishing_fleet"  ,   "species"    ,      "measurement_unit"          ,  "gear_type", "source_authority",
"year", "measurement_value") %>% mutate(measurement_unit = ifelse(measurement_unit %in% c("MTNO", "MT"),"t",measurement_unit)) %>% filter(measurement_unit == "t") #%>% group_by(year) %>% summarise(measurement_value = sum(measurement_value))
row.names(georef_mapped) <- NULL
nominal <- filtering_function(nominal)%>% mutate(year = lubridate::year(time_start)) %>% select("fishing_fleet"  ,   "species"    ,      "measurement_unit"          ,  "gear_type",  "source_authority",
"year","measurement_value")
```

```{r}
georef_no_nominal <- dplyr::anti_join(georef_mapped,nominal, by = strata)
```


```{r}
if (!exists("georef_no_nominal")) {
  knitr::opts_knit$set(stop_on_error = FALSE)
  stop("Table 'georef_no_nominal' does not exist.")
}
```


```{r}
georef_no_nominal <- georef_no_nominal %>% mutate(year = as.character(year))%>% mutate(year = paste0(year, "-01-01"))
dir.create("data/georef_not_nominal")
saveRDS(georef_no_nominal, "data/georef_not_nominal/rds.rds")
```

## General overview of this inconsistent data 

The measurement_value of catches in the georeferenced dataset but not in nominal is different depending on what we take as a strata. For now the strata kept are : 

**`r strata`**, the corresponding measurement_values are : `r round(sum(georef_no_nominal$measurement_value)/1000000,2)` millions, which corresponds of a total of  `r round(100*sum(georef_no_nominal$measurement_value)/sum(nominal$measurement_value),2)` % of the nominal catch and `r round(100*sum(georef_no_nominal$measurement_value)/sum(georef_mapped$measurement_value),2)` % of the georeferenced catch.


```{r}
if(nrow(georef_no_nominal)==0){knitr::knit_exit()}
```





```{r}
georef_no_nominal <- georef_no_nominal%>% mutate(year = lubridate::year(year))

georef_no_nominal_groupped <- georef_no_nominal  %>% group_by(year) %>% summarise(measurement_value = sum(measurement_value/1000000))


# ggplot(georef_no_nominal) +
#   aes(x = year, y = measurement_value) +
#   geom_point(shape = "circle", size = 1.5, colour = "#112446") +
#   labs(
#     title = paste0("Tonnages in georef but not in nominal \n depending on the year (in millions of tons): \n Joined by", str_remove(toString(strata), "year,")) 
#   ) +
#   theme_minimal()

georef_no_nominal_groupped <- georef_no_nominal %>% group_by(year, species, source_authority) %>% summarise(measurement_value = sum(measurement_value))
```


```{r georefnotnominalgrouppedbyyear}
ggplot(georef_no_nominal_groupped) +
  aes(
    x = year,
    y = measurement_value,
    fill = species,
    colour = species
  ) +
  geom_area() +
  scale_fill_hue(direction = 1) +
  scale_color_hue(direction = 1) +
  labs(x = "Year", y = "Value (in tons)") +
  theme_minimal() +
  facet_wrap(vars(source_authority), scales = "free")
```



<!-- The number of row of the table of georef with no equivalent in nominal is `r nrow(georef_no_nominal)`.  -->

<!-- ## Joining without gear_type -->


The number of stratas georeferenced with no equivalent in nominal is `r nrow(georef_no_nominal)`. (Around `r round(100-(((nrow(georef_mapped)-nrow(georef_no_nominal))/nrow(georef_mapped))*100),2)` % of the georeferenced stratas total.


## Analyzis of the data having a declaration in georeferenced with `r strata` but no equivalent in nominal

```{r eval = TRUE}

child_env$parameter_titre_dataset_1 <- "Georeferenced but not nominal data"
child_env$parameter_init <- "data/georef_not_nominal"
child_env$parameter_final <- NULL
child_env$parameter_colnames_to_keep <- c( "species",  "source_authority", "year", "measurement_unit", "measurement_value")
child_env$child <- FALSE
child_env$no_recap_options <- TRUE
child_env$parameter_fact <- "catch"
child_env$time_dimension <- c("year")

res2 <- knitr::knit_child("comparison.Rmd", envir = child_env, quiet = TRUE)

```

```{r eval = TRUE , echo=FALSE, results='asis'}

cat(res2, sep =  "\\clearpage")

```



```{r}

nominal_year_species_source_auth <- nominal %>% group_by(species, year, source_authority) %>% summarise(measurement_value_nominal = sum(measurement_value, na.rm = TRUE))

georef_year_species_source_auth <- georef_mapped %>% group_by(species, year, source_authority) %>% summarise(measurement_value_georef = sum(measurement_value, na.rm = TRUE))

join <- full_join(georef_year_species_source_auth, nominal_year_species_source_auth, by = c("species", "year", "source_authority")) %>% mutate(measurement_value_georef = round(measurement_value_georef, 3), measurement_value_nominal = round(measurement_value_nominal))

georef_sup_to_nom <- join %>% filter(measurement_value_georef > measurement_value_nominal) %>% mutate(Difference = measurement_value_georef - measurement_value_nominal)

```

## Stratas with declaration superior in georeferenced than in nominal

The species concerned by declaration (on a year) of georeferenced captures superior to nominal are : `r paste0(unique(georef_sup_to_nom$species), sep = " ; ")`, for a total (by species of) `r length(unique(georef_sup_to_nom$species))`.


```{r strataswithpostivediff, fig.cap = "Stratas of georeferenced dataset being superior to nominal dataset for a specific year and species", eval=TRUE, include=FALSE}

qflextable2(georef_sup_to_nom %>% mutate_if(is.numeric, round)%>% arrange(desc(Difference) ), save_folder = "georef_sup_nominal")

```


<!-- Only the stratas for which year that there is more declaration in georef than in nominal which are groupped. -->


```{r sumpositivesdiffeveryyear, fig.cap ="Sum of positive differences in every year between georeferenced and nominal"}

qflextable2(georef_sup_to_nom %>% group_by(species, source_authority) %>% summarise(Difference = sum(Difference))%>% mutate_if(is.numeric, round)%>% arrange(desc(Difference) ), captionn = "Sum of positive differences in every year between georeferenced and nominal", save_folder = "georef_sup_nominal")

```

```{r}
georef_sup_to_nom <-ggplot(georef_sup_to_nom) +
  aes(
    x = year,
    y = Difference,
    fill = source_authority,
    colour = source_authority,
    group = source_authority
  ) +
  geom_jitter(size = 1.5) +
  scale_fill_hue(direction = 1) +
  scale_color_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(species), scales = "free")

save_image(georef_sup_to_nom, folder = "georef_sup_nominal")

```


```{r plotgeorefsuptonom, fig.cap = "Plot of georeferenced dataset being superior to nominal dataset for a specific year and species"}

georef_sup_to_nom


```

