---
title: "Conversion factor analysis"
output:
  bookdown::pdf_document2:
    toc: !expr if(!exists('child')) {TRUE} else {child}
    number_sections: TRUE
    keep_tex: true
    extra_dependencies: ["adjustbox", "subfig"]
  bookdown::html_document2:
    toc: !expr if(!exists('child')) {TRUE} else {child}
    number_sections: TRUE
date: '2022-03-23'
params:
  final: "~/Documents/Tunaatlas_level1/jobs/20220927110301_global_catch/entities/global_catch_1deg_level1_final/Markdown/Modifying_IOTC_cwp_errors"
  plotting: "plot"
  world_coastline: "gshhs_world_coastlines"
  con: con
  action: action
  entity: entity
  config: config
  filtering: NULL
  debugging: FALSE

---

\clearpage

This report analyse the conversion factor provided by the RFMOS in the GTA as well as the ones provided by Alain Fonteneau. Both are compared and a conclusion on the quality of each is given as well as leads to improve the treatment operated during level 1 of GTA (converting data provided in number of fish to tons)

# Study of the data provided in MTNO and NOMT

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


```{r}
# RMFOs
RFMO_LEVELS = c("IATTC", "ICCAT", "WCPFC", "IOTC", "CCSBT")
RFMO_COL    = data.frame(RFMO = RFMO_LEVELS, 
                         FILL = ggsci::pal_jco("default", alpha = 0.6)(length(RFMO_LEVELS)), 
                         OUTLINE = colorspace::darken(ggsci::pal_jco("default", alpha = 0.6)(length(RFMO_LEVELS)), 0.2))

# SPECIES
SPECIES_LEVELS = c("ALB", "BET", "BFT", "SKJ", "YFT")
SPECIES_COL    = data.frame(SPECIES_CODE = SPECIES_LEVELS, 
                         FILL = ggsci::pal_simpsons("springfield", alpha = 0.6)(length(SPECIES_LEVELS)), 
                         OUTLINE = colorspace::darken(ggsci::pal_simpsons("springfield", alpha = 0.6)(length(SPECIES_LEVELS)), 0.2))

```

```{r maps}

if(exists("params")){for (i in 1:length(params)){assign(paste0("parameter_",names(params)[i]), params[i][[1]])}
}
shapefile.fix <- st_read("data/world_sf.csv") %>% select(-WKT) %>% rename(geom = geometry)


shape_without_geom  <- shapefile.fix %>% as_tibble() %>%dplyr::select(-geom)

continent <- st_read("data/continent.csv")%>% select(-WKT) %>% rename(geom = geometry)



```

```{r}
is_null_or_not_exist = function(x) {
  if (!exists(deparse(substitute(x))) || is.null(x)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

```


```{r}
rdslevel0 <- readRDS("~/Documents/Tunaatlas_level1/jobs/20220927110301_global_catch/entities/global_catch_1deg_level1_final/Markdown/mapping_codelist/rds.rds")

parameter_filtering <- if(!is_null_or_not_exist(params$action$options$filtering)){params$action$options$filtering} else{ list(species = NULL, fishingfleet = NULL)}


matchingList <- parameter_filtering %>% purrr::keep( ~ !is.null(.) )

colnames_to_filter <- colnames(rdslevel0 %>% select(names(matchingList)))

names(matchingList) <- colnames_to_filter

matchingList <- lapply(matchingList, function(x){ #handling single filter
  if(length(x) == 1){
  x <- c(x, x) }else {
    x
  }

}
  )

is_null_or_not_exist = function(x) {
  if (!exists(deparse(substitute(x))) || is.null(x)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}


if(is_null_or_not_exist(geographical_dimension)){geographical_dimension = "geographic_identifier"}

filtering_function= function(dataframe_to_filter, colnames_to_filter, matchingList, geo_dim){
  if(length(colnames_to_filter)!= 0){  dataframe_to_filter <- dataframe_to_filter%>% filter(!! rlang::parse_expr(str_c(colnames_to_filter, matchingList, sep = '%in%', collapse="&")))}
if("geographic_identifier" %in% colnames(dataframe_to_filter)) {dataframe_to_filter <- dataframe_to_filter%>% dplyr::rename("geographic_identifier" := {{geo_dim}})}
return(dataframe_to_filter)
}

formals(filtering_function)$geo_dim = geographical_dimension

rdslevel0 <- filtering_function(rdslevel0, colnames_to_filter = colnames_to_filter, matchingList = matchingList)

rds_level_0_strata <-rdslevel0 %>% select(-c(value, unit
             # , schooltype, catchtype, fishingfleet
                     ))%>% distinct()

strata_no <- rdslevel0 %>% filter(unit %in% c("no","NO")) %>% select(-c(value, unit
             # , schooltype, catchtype, fishingfleet
                     ))%>% distinct()
strata_t <- rdslevel0 %>% filter(unit %in% c("t","MT")) %>% select(-c(value, unit
             # , schooltype, catchtype, fishingfleet
                     ))%>% distinct()
```


```{r}
if(!exists("colnames_to_keep")){colnames_to_keep <- colnames(rdslevel0)}
time_dimension <- c("time_start", "time_end")
geographical_dimension <- c("geographic_identifier")
list_pie_chart <- list(setdiff(colnames_to_keep, unlist(list(time_dimension, geographical_dimension, "unit", "value"))))[[1]]
```



<!-- There are `r nrow(strata_no)` strata at the end of level0 which are declared in number. -->
<!-- There are `r nrow(strata_t)` strata at the end of level0 which are declared in tons. -->

```{r}
strata_nomt <- rdslevel0 %>% filter(unit %in% c("NOMT"))
strata_mtno <- rdslevel0 %>% filter(unit %in% c("MTNO"))

```

```{r}
conversion_factor_level0 <- rbind(inner_join(strata_nomt , strata_mtno, by = setdiff(colnames(strata_mtno), c("value", "unit") )) %>% dplyr::rename(NO =value.x, MT = value.y) %>%
                                    group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet) %>% 
  summarise(NO = sum(NO), MT = sum(MT)) %>% ungroup()#%>% select(-c(unit.y, unit.x))
  , inner_join(strata_mtno , strata_nomt, by = setdiff(colnames(strata_mtno), c("value", "unit") )) %>% dplyr::rename(MT =value.x, NO = value.y)%>%
  group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet)%>%
  summarise(NO = sum(NO), MT = sum(MT)) %>%ungroup() #%>%  select(-c(unit.y, unit.x))
  )%>% distinct()%>%
  group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet) %>%
  summarise(NO = sum(NO), MT = sum(MT)) %>% mutate(conversion_factor = MT/NO) %>% distinct()
```


```{r}
qflextable2(as.data.frame((rdslevel0 %>% group_by(unit) %>% summarise(`Number of row` = n()))), captionn = "Recap of number of row for each unit for level 0 data.")
```

\clearpage



## Data provided in NOMT without equivalent in MTNO

Some of the data provided in NOMT has no equivalent in MTNO and so is loss in the level 1 treatment. In the new workflow, this data is not loss but considered as NO, thus some of it can be converted.

This data has the following characteristics.

```{r}
if(nrow(strata_nomt) != nrow(strata_mtno)){diff_nomt_mtno <- TRUE}else {diff_nomt_mtno <- FALSE}

anti_join_nomt <- anti_join(strata_nomt %>% select(-c(value,unit)), strata_mtno %>% select(-c(value,unit))) %>% mutate(unit = "NOMT", value = 1) #these are the stratas provided in nomt without equivalent in mtno

anti_join_mtno <- anti_join(strata_mtno %>% select(-c(value,unit)), strata_nomt %>% select(-c(value,unit)))%>% mutate(unit = "MTNO", value = 1) #these are the stratas provided in nomt without equivalent in mtno


```


```{r}
if(nrow(anti_join_nomt)!=0){
pie_charts_multiple <- lapply(list_pie_chart,FUN = function_pie_chart_df, first= anti_join_nomt %>% mutate_if(is.character, as.factor), topn = 3, title_yes_no = FALSE, titre_1 = "Strata available in NOMT but not in MTNO" )

figures<- lapply(pie_charts_multiple, function(x){x$pie_chart})
dimension_title_subfigures <- lapply(pie_charts_multiple, function(x){paste0("Distribution in value for the dimension : ", x$dimension)})
}

```


```{r echo=FALSE, eval=(nrow(anti_join_nomt)!=0), fig.cap='Other dimensions', fig.subcap=c(unlist(gsub("_","..",dimension_title_subfigures))), fig.ncol = 2, out.width = "50%", fig.align = "center"}
for (i in figures){plot(i)}
```


\clearpage

However, some of the data in MTNO has no equivalent in NOMT. This data is kept in the treatment but could be resulting from errors.


## Data provided in MTNO without equivalent in NOMT

This data has the following characteristics.

```{r}
if(nrow(anti_join_mtno)!=0){

pie_charts_multiple <- lapply(list_pie_chart,FUN = function_pie_chart_df, first= anti_join_mtno, topn = 6, title_yes_no = FALSE , titre_1 = "Strata available in MTNO but not in NOMT" )

figures<- lapply(pie_charts_multiple, function(x){x$pie_chart})
dimension_title_subfigures <- lapply(pie_charts_multiple, function(x){paste0("Distribution in value for the dimension : ", x$dimension)})
}

```


```{r echo=FALSE, eval=(nrow(anti_join_mtno)!=0), fig.cap='Other dimensions', fig.subcap=c(unlist(gsub("_","..",dimension_title_subfigures))), fig.ncol = 2, out.width = "50%", fig.align = "center"}
for (i in figures){plot(i)}
```

\clearpage

# Analyse of the conversion factors

```{r}
conversion_factor_level0 <- rbind(inner_join(strata_nomt , strata_mtno, by = setdiff(colnames(strata_mtno), c("value", "unit") )) %>% dplyr::rename(NO =value.x, MT = value.y) %>%
                                    group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet) %>% 
  summarise(NO = sum(NO), MT = sum(MT)) %>% ungroup()#%>% select(-c(unit.y, unit.x))
  , inner_join(strata_mtno , strata_nomt, by = setdiff(colnames(strata_mtno), c("value", "unit") )) %>% dplyr::rename(MT =value.x, NO = value.y)%>%
  group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet)%>%
  summarise(NO = sum(NO), MT = sum(MT)) %>%ungroup() #%>%  select(-c(unit.y, unit.x))
  )%>% distinct()%>%
  group_by(gear, time_start, time_end, geographic_identifier, species, source_authority, schooltype, catchtype ,fishingfleet) %>%
  summarise(NO = sum(NO), MT = sum(MT)) %>% mutate(conversion_factor = MT/NO) %>% distinct()
```

```{r eval= FALSE}
qflextable2(as.data.frame(summary(conversion_factor_level0 %>%ungroup() %>%  dplyr::mutate_if(is.character, as.factor))))
```


```{r}
max_conversion_factor <- read_csv("data/max_conversion_factor.csv", 
    col_types = cols(max_weight = col_double(), 
        min_weight = col_double()))


outside_limit_conv_factor <- full_join(conversion_factor_level0 %>% ungroup(), max_conversion_factor) %>% mutate(conform = case_when(conversion_factor < min_weight ~ "Too low",  conversion_factor > max_weight  ~ "Too high", TRUE ~ "INSIDE"))

a <- outside_limit_conv_factor %>% group_by(conform) %>% summarise(sum = n())

percent_outside = 100-(100*(((a %>% filter(conform == "INSIDE"))$sum)-((a %>% filter(conform != "INSIDE"))$sum))/((a %>% filter(conform == "INSIDE"))$sum))

```

Some of the conversion factors resulting from the declaration are impossible. This means there are higher the maximum catch for this specie or lower than 2 kilos. 

The table presenting the maximum catch used for the filtering is provided in annex.

\clearpage

There are `r nrow(outside_limit_conv_factor %>% filter(conform %in% c("Too low", "Too high")))` stratas that have not conform conversions factors.

## The non conform conversion factors

### Main characteristics of the data


```{r}
floor_decade    = function(value){ return(value# - value %% 10
) }
```


```{r}

ggplot_outside = function(x, data){
  plot <- data %>% filter(conform == x) %>% mutate(decade= floor_decade(lubridate::year(time_start))) %>% 
 # filter(conversion_factor >= 0.2 & conversion_factor <= 4.4) %>%
 ggplot() +
 aes(x = decade, y = conversion_factor, fill = fishingfleet, colour = fishingfleet) +
 geom_point(shape = "circle", 
 size = 1.5) +
 scale_fill_hue(direction = 1) +
 scale_color_hue(direction = 1) +
 theme_minimal() +
 facet_wrap(vars(species), scales = "free_y", ncol = 4)
return(list(pie_chart =plot, dimension = x))
}


ggplot_multiple <- lapply(c("Too high"), ggplot_outside, data = outside_limit_conv_factor)

```

```{r}
figures<- lapply(ggplot_multiple, function(x){x$pie_chart})
dimension_title_subfigures <- lapply(ggplot_multiple, function(x){paste0("Distribution in value for the conversion factors : ", x$dimension)})
```



```{r echo=FALSE, eval=TRUE, fig.cap='Repartition of non conform conversion factors by time, species and fishingfleet', fig.subcap=c(unlist(gsub("_","..",dimension_title_subfigures))), fig.ncol = 1, out.width = "100%", fig.align = "center"}
for (i in figures){plot(i)}
```


```{r}
pie_charts_multiple <- lapply(list_pie_chart,FUN = function_pie_chart_df, first= outside_limit_conv_factor %>% filter(conform %in% c( "Too high")) %>% mutate(unit = "Number of lines", value = 1), topn = 6, title_yes_no = FALSE, titre_1 = "" )
```


```{r eval=TRUE}
figures<- lapply(pie_charts_multiple, function(x){x$pie_chart})
dimension_title_subfigures <- lapply(pie_charts_multiple, function(x){paste0("Distribution in value for the dimension : ", x$dimension)})


```


```{r echo=FALSE, eval=TRUE, fig.cap='Other dimensions', fig.subcap=c(unlist(gsub("_","..",dimension_title_subfigures))), fig.ncol = 2, out.width = "50%", fig.align = "center"}
for (i in figures){plot(i)}
```

```{r}
mean_conv_by_specie <- conversion_factor_level0 %>% group_by(species) %>% summarise(conversion_factor_mean = mean(conversion_factor), conversion_factor_max = max(conversion_factor))
```


### Spatial coverage of the non conform conversion factors

```{r}
shapefile.fix <- shapefile.fix %>% dplyr::rename(geographic_identifier = code)
mapping_non_conform_conv_factor <- outside_limit_conv_factor %>% filter(conform %in% c("Too low", "Too high")) %>% mutate(unit = "Number of lines", value = 1) %>% group_by(geographic_identifier, species, conform) %>% summarise(conversion_factor = mean(conversion_factor))%>% inner_join(shapefile.fix)

# query <- "SELECT  code,st_area(geom), geom from area.gshhs_world_coastlines"
# continent <- (st_read(con, query = query)%>%dplyr::filter(!st_is_empty(.)))

tm_shape(st_as_sf(mapping_non_conform_conv_factor))+
      tm_fill("conversion_factor", palette="RdYlGn", style="cont", n=8,
              id="name", midpoint = 0)+tm_facets(by = c("species", "conform"), free.scales = TRUE)+tm_shape(continent)+tm_borders()

```


## The conform conversion factors

### Mean conversion factors by species

```{r}
qflextable2(mean_conv_by_specie, captionn = "Mean and Max conversion factors provided by tRFMOs for each species")
```


### Characteristics of treated data (level1)

```{r}
rdslevel1 <- readRDS("~/Documents/Tunaatlas_level1/jobs/20220927110301_global_catch/entities/global_catch_1deg_level1_final/Markdown/raising/rds.rds")

rds_level1 <- filtering_function(rdslevel1, colnames_to_filter = colnames_to_filter, matchingList = matchingList)
rds_level_1_strata <- rdslevel1 %>% select(-c(value, unit
                   # ,  schooltype, catchtype, fishingfleet
         ))%>% distinct()

percent_strata_loss_level0_level1 <- round(100*(nrow(rds_level_0_strata)-nrow(rds_level_1_strata))/nrow(rdslevel0))
number_strata_loss_level0_level1 <- (nrow(rds_level_0_strata)-nrow(rds_level_1_strata))

conversion_factors_num_to_weight_ird_2021 <- read_csv("data/conversion_factors_num_to_weight_ird_2021.csv")%>% select(-c(unit, unit_target))%>% mutate(geographic_identifier = as.character(geographic_identifier)) %>% mutate(Year = lubridate::year(time_start)) 

# allcombs = expand.grid(species = unique(conversion_factors_num_to_weight_ird_2021$species), 
#                        geographic_identifier = unique(conversion_factors_num_to_weight_ird_2021$geographic_identifier), gear = unique(conversion_factors_num_to_weight_ird_2021$gear), source_authority = unique(conversion_factors_num_to_weight_ird_2021$source_authority), 
#   Year = min(lubridate::year(conversion_factors_num_to_weight_ird_2021$time_start)):max(lubridate::year(conversion_factors_num_to_weight_ird_2021$time_start)),
#   Month = 1:12,
#   stringsAsFactors = F
# )

allcombs2 = expand.grid(
  Year = min(lubridate::year(conversion_factors_num_to_weight_ird_2021$time_start)):max(lubridate::year(conversion_factors_num_to_weight_ird_2021$time_start)),
  Month = 1:12,
  stringsAsFactors = F
)


conversion_factors_num_to_weight_ird_2021 <- full_join(allcombs2,conversion_factors_num_to_weight_ird_2021 , by = c("Year")) %>% distinct()


# test <- full_join(allcombs,conversion_factors_num_to_weight_ird_2021 %>% mutate(Year = lubridate::year(time_start)), by = c("species", "gear", "source_authority",  "geographic_identifier", "Year")) %>% distinct() %>%rowwise() %>%  mutate(new_date = lubridate::as_date(paste0(Year,"-",Month, "-01 ")))

# df2 <- rbind(allcombs,conversion_factors_num_to_weight_ird_2021 %>% mutate(year)) %>%
#   group_by(Cust,Year,Month) %>%
#   summarize(Value = sum(Value)) %>%
#   ungroup()
# 
# test <- conversion_factors_num_to_weight_ird_2021 %>%
#   mutate(year2 = lubridate::year(time_start)) %>% 
#   group_by(year2) %>% 
#   mutate(month = 1) %>%
#   complete(year2, month = 1:12) %>%
#   fill(species,gear, conversion_factor)

# out <- try((con <- config$software$output$dbi), silent = TRUE)
# if("try-error" %in% class(out)){con <- params$con}
# 
# query <- "SELECT code, st_area(geom), geom from area.cwp_grid"
# out <- try((world_sf <- st_read(con, query = query)), silent = TRUE)
# if("try-error" %in% class(out)){con <- DBI::dbConnect(RPostgres::Postgres() , dbname = params$dbname, host = params$host,   port = params$port,
#                                                        user = params$user,
#                                                        password = params$password)}

areas_conversion_factors_numtoweigth_ird <- st_read("data/areas_conversion_factors_numtoweigth_ird.csv")%>% select(-WKT) %>% rename(geom = geometry)



# geom_ird <- con %>% tbl(in_schema("area", "areas_conversion_factors_numtoweigth_ird")) %>% collect()


# query <- "SELECT  code,st_area(geom), geom from area.cwp_grid"
# world_sf <- st_make_valid(st_read(con, query = query))%>% filter(!st_is_empty(.))

JOINED_DATA <- st_join(world_sf, areas_conversion_factors_numtoweigth_ird) %>% dplyr::rename(GTA = code.x, IRD = code.y) %>% select(GTA, IRD) %>% mutate(IRD = ifelse(is.na(IRD), 0, IRD))
    
geographic_identifier_to_ird = function(x){
x <- x %>% full_join(JOINED_DATA, by = c("geographic_identifier" = "GTA")) %>% dplyr::rename(geographic_identifier_old = geographic_identifier, geographic_identifier = IRD ) %>% distinct() %>% mutate(Year = lubridate::year(time_start), Month = lubridate::month(time_start))
  return(x)
}

conversion_factor_level0 <- geographic_identifier_to_ird(conversion_factor_level0)
```

There are `r number_strata_loss_level0_level1` strata that are loss while raising on level 1 (`r percent_strata_loss_level0_level1` % of the total). This means there is no conversion done for those strata.


```{r}
pie_charts_multiple <- lapply(list_pie_chart,FUN = function_pie_chart_df, first= rdslevel0, topn = 6, title_yes_no = FALSE, titre_1 = "Rds_level0" )
```


```{r eval=TRUE}
figures<- lapply(pie_charts_multiple, function(x){x$pie_chart})
dimension_title_subfigures <- lapply(pie_charts_multiple, function(x){paste0("Distribution in value for the dimension : ", x$dimension)})


```


```{r echo=FALSE, eval=TRUE, fig.cap='Other dimensions', fig.subcap=c(unlist(gsub("_","..",dimension_title_subfigures))), fig.ncol = 2, out.width = "50%", fig.align = "center"}
for (i in figures){plot(i)}
```

```{r}
pie_chart_species <- pie_chart_2_default("species", spec_col = SPECIES_COL,first = conversion_factor_level0_too_high, topn = length(unique(conversion_factor_level0_too_high$species))+1, title_yes_no = FALSE, titre_1 = "")
```


\clearpage

# Comparison (for the data available) of the conversion factor of RFMOs and the conversion factor of Alain Fonteneau

```{r}

strata <- conversion_factors_num_to_weight_ird_2021 %>% select(-conversion_factor)

strata_level1_converted <- round((nrow(rds_level_1_strata) - nrow(strata_t))/nrow(strata))
```

The number of strata in the file of Alain Fonteneau, is `r nrow(strata)`. Which means each conversion factor is used for (in means) `r strata_level1_converted` strata.

```{r}
mapping_codelist <-readRDS("data/mapping_codelist/rds.rds")
mapping_codelist <- filtering_function(mapping_codelist, colnames_to_filter = colnames_to_filter, matchingList = matchingList)
 mapping_codelist <- geographic_identifier_to_ird(mapping_codelist)
MTNO <- mapping_codelist %>% filter(unit == "MTNO")
NOMT <- mapping_codelist %>% filter(unit == "NOMT")

setdiff <- setdiff(NOMT %>% select(-c(value, unit)), MTNO %>% select(-c(value, unit)))
setdiff2 <- setdiff(MTNO %>% select(-c(value, unit)), NOMT %>% select(-c(value, unit)))


```

<!-- There are `r nrow(conversion_factors_num_to_weight_ird_2021)` stratas converted in level 1. -->

```{r}
converted_level0 <- rbind(full_join(NOMT , MTNO, by = setdiff(colnames(MTNO), c("value", "unit") )) %>% dplyr::rename(NO =value.x, MT = value.y) %>% select(-c(unit.y, unit.x)), full_join(MTNO , NOMT, by = setdiff(colnames(MTNO), c("value", "unit") )) %>% dplyr::rename(MT =value.x, NO = value.y) %>% select(-c(unit.y, unit.x)))%>% distinct() %>% mutate(conversion_factor = MT/NO) #%>% full_join(test, by = c("geographic_identifier"="code")) %>% mutate(geographic_identifier = source) 

converted <- converted_level0 %>% select(setdiff(colnames(conversion_factors_num_to_weight_ird_2021),c("unit", "unit_target")))%>% distinct()
```

<!-- There are `r nrow(converted)` stratas converted in level 0. -->

```{r}

# converted_level0 <- geographic_identifier_to_ird(converted_level0)

comparison_converted_factor <- inner_join(conversion_factors_num_to_weight_ird_2021%>% dplyr::rename(conversion_factor_ird =conversion_factor) %>% mutate_if(is.Date, as.character),converted_level0%>% dplyr::rename(conversion_factor_gta =conversion_factor), by = c("gear", "Year", "Month", "geographic_identifier", "species", "source_authority")) 



```


We only analyse species having conversion factors from rfmos and Alain Fonteneau.

```{r}
# IOTC_data <- rbind(IOTC_DATASETS_CECoastal, IOTC_DATASETS_CELongline, IOTC_DATASETS_CESurface) %>% inner_join(IOTC_conversion_factor_3)
```

```{r}
conversion_factor_level0_to_compare <- conversion_factor_level0  %>% group_by(gear, Year, Month, geographic_identifier, species, source_authority) %>%
  summarise(NO = sum(NO), MT = sum(MT)) %>% mutate(conversion_factor = MT/NO) %>% distinct() 

# conversion_factor_level0_to_compare <- geographic_identifier_to_ird(conversion_factor_level0_to_compare)

```


```{r}
mean_conversion_factor <- conversion_factor_level0_to_compare %>% group_by(species, source_authority, gear, geographic_identifier, Year, Month) %>% mutate(conv_factor_mean = mean(conversion_factor), conv_factor_median = median(conversion_factor)) %>% mutate(ecart_a_la_moyenne = 100*(conversion_factor - conv_factor_mean)/conv_factor_mean,ecart_a_mediane = 100*(conversion_factor - conv_factor_median)/conv_factor_median)

```


```{r}
ird <- conversion_factors_num_to_weight_ird_2021 %>%  mutate(geographic_identifier = as.character(geographic_identifier))%>% select(setdiff(colnames(conversion_factors_num_to_weight_ird_2021), c("unit_target","unit"))) %>% mutate(time_start = as.character(time_start),time_end =as.character(time_end), geographic_identifier = as.character(geographic_identifier)) 
rfmo <- mean_conversion_factor%>%  select(setdiff(colnames(ird), c("time_start","time_end", "date")))#mutate(time_start = lubridate::year(time_start)) %>% mutate(time_end = lubridate::year(time_end))

combine <- gdata::combine(rfmo,ird )#%>% filter(species %in%params$species)
# combine <- combine %>% filter(species != "PBF") %>% mutate_if(is.character, as.factor)
ggplot(combine , aes( y = conversion_factor, col = source))+geom_boxplot()+facet_wrap("species", scales="free")





RFMO <- left_join(rfmo, ird %>% select(-conversion_factor),by = c("gear", "source_authority", "species","Year", "Month", "geographic_identifier"))


IRD <- left_join(ird, rfmo%>% select(-conversion_factor),by = c("gear", "source_authority", "species","Year", "Month", "geographic_identifier"))

double_data <- gdata::combine(RFMO,IRD ) %>%ungroup() %>%  group_by(species) %>% summarise(sum = n_distinct(source))#%>% filter(species %in%params$species)

combine <- gdata::combine(RFMO,IRD ) %>% inner_join(double_data) %>% filter(sum ==2)

if(!require(ggpubr)){
  install.packages("ggpubr")
  require(ggpubr)
}
if(!require(rstatix)){
  install.packages("rstatix")
  require(rstatix)
}
if(!require(tidyverse)){
  install.packages("tidyverse")
  require(tidyverse)
}


# combine <- full_join_2 %>% pivot_longer(cols =c(conversion_factor_ird, conversion_factor_rfmo), names_to = "source", values_drop_na = TRUE,values_to = "conversion_factor")

ggplot(combine  , aes( y = conversion_factor, col = source))+geom_boxplot()+facet_wrap("species", scales="free")





ggplot(combine%>% inner_join(max_conversion_factor) %>% filter(conversion_factor < max_weight)  , aes( y = conversion_factor, col = source))+geom_boxplot()+facet_wrap("species", scales="free")


espece_parametrage <- read_delim("data/espece_parametrage.csv", 
                                     delim = ";", escape_double = FALSE, trim_ws = TRUE)

espece_parametrage2 <- filtering_function(espece_parametrage %>%dplyr::rename(species =C_ESP_3L), colnames_to_filter = "species", matchingList = matchingList) %>% select(species, C_OCEA, V_POIDS_MOY)
```

## Atlantic

```{r}
moy_ob7 <- combine%>% inner_join(max_conversion_factor) %>% filter(conversion_factor < max_weight) %>% inner_join(espece_parametrage2) %>% filter(source_authority == "ICCAT" )


ggplot(moy_ob7, aes( y = conversion_factor, col = source))+geom_boxplot()+facet_grid(~species, scales="free")#+geom_hline(data = moy_ob7 %>% filter(C_OCEA == 1) ,         aes(yintercept = V_POIDS_MOY, colour= "Atlantic IRD Mean Value"),show.legend = "C_OCEA")

```

## Indian

```{r}
moy_ob7 <- combine%>% inner_join(max_conversion_factor) %>% filter(conversion_factor < max_weight) %>% inner_join(espece_parametrage2)%>% filter(source_authority == "IOTC" )


ggplot(moy_ob7, aes( y = conversion_factor, col = source))+geom_boxplot()+facet_grid(~species, scales="free")#+geom_hline(data = moy_ob7 %>% pivot_wider(names_from = "C_OCEA", values_from = "V_POIDS_MOY")%>% select(species = species, Atlantic = "1", Indian = "2") %>% distinct()  %>% ungroup(),         aes(yintercept = Indian, colour= "Indian IRD Mean Value"))
```


## Comparison 1 by 1 of the conversion factor for each strata

```{r}
full_joi <- full_join(RFMO, IRD%>% mutate(time_start = as.character(time_start),time_end =as.character(time_end), geographic_identifier = as.character(geographic_identifier)) ,by = c("gear", "source_authority", "species","Year", "Month", "geographic_identifier")) %>% dplyr::rename(conversion_factor_ird = conversion_factor.y, conversion_factor_rfmo=conversion_factor.x   )


to_comp <- full_joi %>% filter(!(is.na(conversion_factor_ird) || is.na(conversion_factor_rfmo)))
```


```{r}
to_comp <- to_comp %>% mutate(ecart_rfmo_ird_percent = 100*abs(conversion_factor_rfmo - conversion_factor_ird )/conversion_factor_rfmo)
```

```{r}
ggplot(to_comp) +
  aes(x = ecart_rfmo_ird_percent, group = species) +
  geom_histogram(bins = 30L, fill = "#112446") +
  theme_minimal() +
  facet_wrap(vars(species), scales = "free")
```


```{r}
ggplot( inner_join(to_comp %>% ungroup(), max_conversion_factor) %>% mutate(conform = case_when(conversion_factor_rfmo < min_weight ~ "Too low",  conversion_factor_rfmo > max_weight  ~ "Too high", TRUE ~ "INSIDE")) %>% filter(conform == "INSIDE")) +
  aes(x = ecart_rfmo_ird_percent, group = species) +
  geom_histogram(bins = 30L, fill = "#112446") +
  theme_minimal() +
  facet_wrap(vars(species), scales = "free")
```


```{r}

 new_table_longer <- inner_join(to_comp %>% ungroup(), max_conversion_factor) %>% mutate(conform = case_when(conversion_factor_rfmo < min_weight ~ "Too low",  conversion_factor_rfmo > max_weight  ~ "Too high", TRUE ~ "INSIDE")) %>% dplyr::rename(RFMO =conversion_factor_rfmo, IRD =conversion_factor_ird )%>% filter(conform == "INSIDE") %>% pivot_longer(cols = c("RFMO", "IRD"),names_to = "source" ) %>% dplyr::rename(time_start = time_start.x)
myplot <- ggboxplot(new_table_longer, x = "source", y = "value",
  fill = "source", palette = "npg", legend = "none",
  ggtheme = theme_pubr(border = TRUE)
  ) +
  facet_wrap(~species)
myplot

stat.test <- new_table_longer%>% inner_join(max_conversion_factor) %>% filter(value < max_weight)%>%
group_by(species) %>%
t_test(value ~ source) %>%
adjust_pvalue(method = "BH") %>%
add_significance()
res.aov <- new_table_longer%>% inner_join(max_conversion_factor) %>% filter(value < max_weight) %>%
anova_test(value ~ species*time_start)
stat.test <- stat.test %>% add_xy_position(x = "source")
myplot + stat_pvalue_manual(stat.test, label = "p.adj.signif")
```



# Annexes

```{r}
qflextable2(filtering_function(max_conversion_factor, colnames_to_filter = "species", matchingList = matchingList))
```



## IOTC Focuse


```{r}

#retrieve original data of IOTC
IOTC_DATASETS_CESurface <- read_csv("data/IOTC_data/drive-download-20220506T101442Z-001/IOTC-DATASETS-CESurface.csv") %>%
  dplyr::select(c(`Fleet`,      `Gear`,       `Year`,       `MonthStart`, `MonthEnd`,   `iGrid`,     `Grid`,       `Effort`,     `EffortUnits`,`QualityCode`,`Source`,     `CatchUnits`) | starts_with(c("YFT","BET","SWO", "ALB","SKJ","SBF")))#%>% filter(Fleet == `CHN`,& iGrid == `6215065`)
IOTC_DATASETS_CELongline <- read_csv("data/IOTC_data/drive-download-20220506T101442Z-001/IOTC-DATASETS-CELongline.csv")%>%
  dplyr::select(c(`Fleet`,      `Gear`,       `Year`,       `MonthStart`, `MonthEnd`,   `iGrid`,     `Grid`,       `Effort`,     `EffortUnits`,`QualityCode`,`Source`) | starts_with(c("YFT","BET","SWO", "ALB","SKJ", "SBF")))#%>% filter(Fleet == "CHN" & iGrid == "6215065")
IOTC_DATASETS_CECoastal <- read_csv("data/IOTC_data/drive-download-20220506T101442Z-001/IOTC-DATASETS-CECoastal.csv")%>%
  dplyr::select(c(`Fleet`,      `Gear`,       `Year`,       `MonthStart`, `MonthEnd`,   `iGrid`,     `Grid`,       `Effort`,     `EffortUnits`,`QualityCode`,`Source`) | starts_with(c("YFT","BET", "ALB","SKJ", "SBF","SWO"#, "LOT", "FRZ", "KAW", "COM", "BILL", "TUX", "SKH", "NTAD"
                                                                                                                                                                                       ))) #%>% filter(Fleet == "CHN" & iGrid == "6215065")


IOTC_interese <- read_csv("data/IOTC_data/drive-download-20220506T101442Z-001/IOTC-DATASETS-CELongline.csv")%>% select( c(Fleet, Gear, Year, MonthStart, MonthEnd, iGrid, Effort, EffortUnits, QualityCode, Source)| starts_with(c("YFT","BET","SWO", "ALB", "SKJ", "SBF"))) #%>% filter(Fleet == "CHN" & iGrid == "6215065") %>% filter(Year == 2011)
library(tidyr)
cleaned <- rbind(IOTC_DATASETS_CECoastal, IOTC_DATASETS_CELongline) %>% janitor::clean_names()%>% tidyr::pivot_longer(cols = starts_with(c("YFT","BET","SWO", "ALB", "SKJ", "SBF"#, "LOT", "FRZ", "KAW", "COM", "BILL", "TUX", "SKH", "NTAD"
                                                                                                                                           )), names_to = c("species"))%>% separate(species,into = c("species","unit"),sep= "_") %>% filter(!is.na(value))%>% mutate(species = toupper(species), unit = toupper(unit)) %>% dplyr::rename(fishingfleet = fleet)
```

```{r}
cleaned_fact_conv <- cleaned %>% pivot_wider(values_from = value, names_from = unit ) %>% filter(!is.na(NO) & !is.na(MT)) %>% mutate(fact_conv = MT/NO)

iotc_and_max <- cleaned_fact_conv %>% inner_join(max_conversion_factor) %>% filter(max_weight < fact_conv)

```

```{r}
error_costoal <- IOTC_DATASETS_CECoastal %>% janitor::clean_names()%>% inner_join(iotc_and_max )
```


```{r}
error_longline <- IOTC_DATASETS_CELongline %>% janitor::clean_names()%>% inner_join(iotc_and_max )

qflextable2(rbind(error_longline, error_costoal))



```

