---
title: "Summary of the treatment for a specific entity of the global tuna atlas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Bastien Grasset
    email: bastien.grasset@ird.fr
    affiliation:
      - IRD
      - MARBEC
  - name: Emmanuel Blondel
    affiliation: FAO
  - name: Julien Barde
    affiliation:
      - IRD
      - MARBEC
address:
  - code: IRD
    address: Institut de Recherche pour le Développement (IRD), av. Jean Monnet, CS 30171, 34203 Sète cédex, France
  - code: MARBEC
    address: MARBEC, University of Montpellier, CNRS, Ifremer, IRD, Sète, France
# bibliography: ["IOTC-2022-WPDCS18.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    extra_dependencies: ["adjustbox", "subfig", "flafter", "makecell"]
    toc: TRUE
    keep_tex: TRUE
    template: template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
    always_allow_html: TRUE
    keep_md: TRUE
  bookdown::html_document2: 
      # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
    keep_md: true
  bookdown::word_document2: 
      # Produces largely readable output, though some cross-referencing may fail. Useful for collaboration.
    toc: TRUE
lang: en # Main document language in BCP47 format
abstract: "This document is aimed to recap the impact on the data of the different treatment of the Global Tuna Atlas. \nFirst we present the data created by the GlobalTuna Atlas, as well as the comparison between this data and the rawdata provided by the RFMO’s before the global Tuna Atlas treatment. \nIn a second time we will present the process of the GlobalTuna Atlas in order to reference exactly the choices made."
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: FALSE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
fancyhdr:
  first:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 0pt
    #footleft: A left foot
    footrulewidth: 0pt
  subsequent:
    headleft: "Global Tuna Atlas outputs and process"
    headright: "2023"
    headrulewidth: 1pt
    footrulewidth: 0pt
---

\clearpage

```{r setuprmd, include = FALSE}

res <- knitr::knit_child("Setup_markdown.Rmd", quiet = TRUE, envir = environment())

```

```{r child-env-base, include = FALSE}

child_env_base <- new.env(parent = environment())
source("https://raw.githubusercontent.com/firms-gta/geoflow-tunaatlas/master/Analysis_markdown/functions/Functions_markdown.R", local = child_env_base)


```

```{r parameters-global, include=FALSE}

opts <- entity$data$actions[[1]]$options
con <- config$software$input$dbi

if(!exists("parameter_fact") | is.null("parameter_fact")){parameter_fact <-  opts$fact}

if(!is.null(opts$filtering)){parameter_filtering <-opts$filtering} else{ parameter_filtering <-list(species = NULL, fishingfleet = NULL)}

if(is.character(parameter_filtering)){
  parameter_filtering <- eval(parse(text=toString(parameter_filtering)))
}

parameters_child <- list(config = config, entity = entity,opts = opts ,action = action, parameter_filtering = parameter_filtering, parameter_fact = parameter_fact, parameter_con = con,  child = TRUE, no_recap_options = TRUE, plotting_type = "plot",parameter_colnames_to_keep = c("fishing_fleet",         "gear_type",                 "time_start",                 
                      "geographic_identifier","fishing_mode",           "species",                       
                      "measurement_unit",                 "measurement_value",                "source_authority", "Gear", "species_group"), treatment = FALSE, shapefile.fix = shapefile.fix, shape_without_geom = shape_without_geom, continent = continent, fig.path = fig.path, outputonly = FALSE, print_map = TRUE, coverage = FALSE, parameter_geographical_dimension = "geographic_identifier", parameter_geographical_dimension_groupping = "GRIDTYPE")

list2env(parameters_child, env = child_env_base)

```

\clearpage

# Summary of the final provided data : `r gsub("_", "", entity$identifiers[["id"]])`

```{r folder-with-rds, message=FALSE, warning=FALSE, include=FALSE}

sub_list_dir_2 <- paste0("Markdown/",list.dirs(path =paste0(getwd(),"/Markdown"), full.names = FALSE, recursive = FALSE))
details = file.info(paste0(sub_list_dir_2,"/Markdown"))
details = file.info(sub_list_dir_2)
details = details[with(details, order(as.POSIXct(mtime))), ]
sub_list_dir_2 = rownames(details)

```


```{r analysislast, include=FALSE, results='asis'}

child_env_last <- list2env(as.list(child_env_base), parent = child_env_base)

child_env_last$parameter_init <- paste0(sub_list_dir_2[length(sub_list_dir_2)], "/rds.rds")
child_env_last$parameter_titre_dataset_1 <- entity$identifiers[["id"]]
child_env_last$parameter_final <- NULL
child_env_last$child <- TRUE
child_env_last$no_recap_options <- TRUE
child_env_last$unique_analyse <- TRUE
if(exists("debugging") && debugging == TRUE){
  child_env_last$debugging <- TRUE
  } else {child_env_last$debugging <- FALSE}
child_env_last$parameter_mapped <- TRUE
child_env_last$last_to_first <- FALSE
child_env_last$parameter_short <- FALSE
child_env_last$child_header <- ""
child_env_last$coverage <- TRUE

filename <- paste0("Report_on_", entity$identifiers[["id"]], ".pdf")

new_path <- file.path(fig.path, filename)

dir.create(new_path, recursive = TRUE)

child_env_last$fig.path = new_path
# rmarkdown::render("comparison.Rmd", envir = child_env_last, quiet = TRUE)

res2 <- knitr::knit_child("comparison.Rmd", envir = child_env_last, quiet = TRUE)


rm(child_env_last)
gc()

```

```{r echo=FALSE, results='asis'}

cat(res2, sep =  "\\clearpage")

```



```{r include=FALSE}
rm(res2)
gc()
```


\clearpage

# Summary of the treatment done on the data (comparison between the Initial data and the data resulting from treatments)

```{r complastfirst, include=FALSE, results='asis'}

child_env_comp <- list2env(as.list(child_env_base), parent = child_env_base)


child_env_comp$parameter_init <- paste0(sub_list_dir_2[1], "/rds.rds")
child_env_comp$parameter_final <- paste0(sub_list_dir_2[length(sub_list_dir_2)], "/rds.rds")

child_env_comp$unique_analyse <- FALSE
child_env_comp$parameter_titre_dataset_1 <- "Initial_data"
child_env_comp$parameter_titre_dataset_2 <- entity$identifiers[["id"]]
child_env_comp$last_to_first <- TRUE
child_env_comp$parameter_mapped <- TRUE
child_env_comp$parameter_short <- FALSE
child_env_comp$child_header <- "#"

new_path <- file.path(fig.path, paste0("/Comparison/initfinal_",gsub("Markdown/","",sub_list_dir_2[1]),"_", gsub("Markdown/","",sub_list_dir_2[length(sub_list_dir_2)])))

dir.create(new_path, recursive = TRUE)

child_env_comp$fig.path = new_path

res <- knitr::knit_child("comparison.Rmd", envir = child_env_comp, quiet = TRUE)

rm(child_env_comp)
gc()

```


```{r complastfirstcat, echo=FALSE, results='asis'}

cat(res, sep =  "\\clearpage")

```


```{r include=FALSE}
rm(res)
gc()
```


\clearpage


# Analyse of the processing of the data

The following table recap all the treatment done on the data provided by tRFMOs.

```{r include=FALSE, eval=(parameter_fact == "catch")}

if(dir.exists("Markdown")){

    nominal_dataset <- readr::read_csv(paste0("data/global_nominal_catch_firms_level0.csv"))

nominal_dataset <- filtering_function(nominal_dataset, parameter_filtering = parameter_filtering)
}

if(exists("nominal_dataset")){nominal <- sum(nominal_dataset$measurement_value)} else {nominal <- 1}
df <- data.frame(matrix(ncol =13, nrow = 1))
      colnames(df) <- c(paste0(tail(str_split(paste0(sub_list_dir_2),"/")[[1]],n=1)),
                        "Explenation", "Fonctions", 
                        "Options", "Tons", "Number of fish", "Lines","Difference (in % of tons)","Difference in tons","Difference (in % of fish)", "Difference in number of fish", "Difference (in % of lines)", "Percentage of nominal"
      )
          main <- filtering_function(readRDS(paste0(sub_list_dir_2[1],"/rds.rds")), parameter_filtering = parameter_filtering)
          tons_init <- sum((main %>% dplyr::filter(measurement_unit%in%c("MTNO", "MT","t", "Tons")))$measurement_value)
          nofish_init <- sum((main %>% dplyr::filter(measurement_unit%in%c("NOMT", "NO","no", "Number of fish")))$measurement_value)
          lines_init <- nrow(main)
      for (i in sub_list_dir_2){
        Explenation <- readtext(paste0(i,"/explication.txt"))[2]
        Fonctions <- pull(readtext(paste0(i,"/fonctions.txt"))[2])
        if (file.exists(paste0(i,"/options_written.txt"))){
          Options <- pull(readtext(paste0(i,"/options_written.txt"))[2])} else {Options <- "Aucune"}
        if(isNullList(parameter_filtering)){
          sums <- read_csv(paste0(i,"/sums.csv"))
        sum_t <- sums$somme_t
        sum_no <- sums$somme_no
        nrow <- sums$lines
        } else {
          main <- filtering_function(readRDS(paste0(i,"/rds.rds")), parameter_filtering = parameter_filtering)
          sum_t <- sum((main %>% dplyr::filter(measurement_unit%in%c("MTNO", "MT","t", "Tons")))$measurement_value)
          sum_no <- sum((main %>% dplyr::filter(measurement_unit%in%c("NOMT", "NO","no", "Number of fish")))$measurement_value)
         nrow <- nrow(main)

        }

        step <- tail(str_split(paste0(i),"/")[[1]],n=1)
        Difference_percent <- -100*((tons_init - sum_t)/tons_init)
        Difference_tons <- -(tons_init - sum_t)
        Difference_no <- -(nofish_init - sum_no)
        Difference_percent_lines <- -100*((lines_init - nrow)/lines_init)
        Difference_percent_no <- -100*((nofish_init - sum_no)/nofish_init)
        percentage_of_nominal <- round((sum_t*100)/nominal, 1)
        sums <- as.data.frame(data.frame(sum_t, sum_no, nrow))
        data_i <- cbind(step,
                        Explenation, Fonctions, 
                        Options,
                        sums, Difference_percent,Difference_tons,Difference_percent_no,Difference_no, Difference_percent_lines, percentage_of_nominal)
        names(data_i) <- colnames(df)
        df <- rbind(df, data_i)
        tons_init <- sum_t
        nofish_init <- sum_no
        lines_init <- nrow
        
      }
      df2 = df[-1,]
      df2[df2 == -Inf] <- 0
      colnames(df2)[1] <- "Step"



     

      reduced <- df2 %>% dplyr::mutate(`Millions of tons` = `Tons` / 1000000, `Millions of fish` = `Number of fish` / 1000000)%>% dplyr::select(Step, `Millions of tons`, `Millions of fish`,# `Lines`,
                                                                "Difference (in % of tons)","Difference (in % of fish)", "Percentage of nominal")%>% dplyr::mutate(`Step number` = as.numeric(row_number()))
      reduced$Step <- factor(reduced$Step, levels = (reduced %>% dplyr::arrange(`Step number`))$Step)
      

coeff <- 3
temperatureColor <- "#69b3a2"
      
priceColor <- rgb(0.2, 0.6, 0.9, 1)

second_graf <- ggplot(reduced,aes(x = Step,group = 1))+ 
        geom_line( aes(y=`Millions of tons` ,group = 1), size=0.5, color=priceColor)+geom_point( aes(y=`Millions of tons` ,group = 1)) +
        geom_line( aes(y=`Millions of fish`/ coeff,group = 1), size=0.5, color=temperatureColor)+geom_point( aes(y=`Millions of fish`/ coeff))  +
        scale_y_continuous(
          
          # Features of the first axis
          name = "Tons",
          
          # Add a second axis and specify its features
          sec.axis = sec_axis(~.*coeff, name="Number of fish")
        ) + 
        
        
        
        theme(
          axis.title.y = element_text(color = priceColor, size=8),
          axis.title.y.right = element_text(color = temperatureColor, size=8)
        ) +
        
        ggtitle("Evolution of the repartition of captures depending on units and Steps")+
        theme(axis.text.x = element_text(angle = 90))

no_fish_plot <- ggplot(reduced,aes(x = Step,group = 1)) +geom_line( aes(y=`Millions of fish`,group = 1), size=0.5)+theme(axis.text.x = element_text(angle = 90))
      
tons_plot <- ggplot(reduced,aes(x = Step,group = 1)) +
geom_line( aes(y=`Millions of tons`,group = 1), size=0.5)+theme(axis.text.x = element_text(angle = 90))

```

```{r include=FALSE , eval=(parameter_fact == "effort")}

nominal <- 1
df <- data.frame(matrix(ncol =13, nrow = 1))
      colnames(df) <- c(paste0(tail(str_split(paste0(sub_list_dir_2),"/")[[1]],n=1)),
                        "Explenation", "Fonctions", 
                        "Options", "Sum in hooks", "Sum in fishing days", "Lines","Difference (in % of hooks)","Difference in hooks","Difference (in % of fishing days)","Difference in fishing days",  "Difference (in % of lines)", "Percentage of nominal"
      )

          main <- filtering_function(readRDS(paste0(sub_list_dir_2[1],"/rds.rds")), parameter_filtering = parameter_filtering)# %>% dplyr::filter(species%in%opts$species_filter)
          hooks_init <- sum((main %>% dplyr::filter(measurement_unit%in%c("HOOKS", "HOOKS")))$measurement_value)
          fishing_days_init <- sum((main %>% dplyr::filter(measurement_unit%in%c("FDAYS", "FDAYS")))$measurement_value)
          lines_init <- nrow(main)
        
      for (i in sub_list_dir_2){
        sums <- read.csv(paste0(i,"/sums.csv"))
        Explenation <- readtext(paste0(i,"/explication.txt"))[2]
        Fonctions <- pull(readtext(paste0(i,"/fonctions.txt"))[2])
        if (file.exists(paste0(i,"/options_written.txt"))){
          Options <- pull(readtext(paste0(i,"/options_written.txt"))[2])} else {Options <- "Aucune"}
          main <- filtering_function(readRDS(paste0(i,"/rds.rds")), parameter_filtering = parameter_filtering)# %>% dplyr::filter(species%in%filter_species)
          sum_hooks <- sum((main %>% dplyr::filter(measurement_unit%in%c("HOOKS", "HOOKS")))$measurement_value)
          sum_fdays <- sum((main %>% dplyr::filter(measurement_unit%in%c("FDAYS", "FDAYS")))$measurement_value)
          nrow <- nrow(main)
          

        step <- tail(str_split(paste0(i),"/")[[1]],n=1)
        Difference_percent <- -100*((hooks_init - sum_hooks)/hooks_init)
        Difference_hooks <- -(hooks_init - sum_fdays)
        Difference_fdays <- -(fishing_days_init - sum_hooks)
        Difference_percent_lines <- -100*((lines_init - nrow)/lines_init)
        Difference_percent_fdays <- -100*((fishing_days_init - sum_fdays)/fishing_days_init)
        percentage_of_nominal <- (sum_hooks*100)/nominal
        sums <- as.data.frame(data.frame(sum_hooks, sum_fdays, nrow))
        data_i <- cbind(step,
                        Explenation, Fonctions, 
                        Options,
                        sums, Difference_percent,Difference_hooks,Difference_percent_fdays,Difference_fdays, Difference_percent_lines, percentage_of_nominal)
        names(data_i) <- colnames(df)
        df <- rbind(df, data_i)
        hooks_init <- sum_hooks
        fishing_days_init <- sum_fdays
        lines_init <- nrow
        
      }
      df2 = df[-1,]
      df2[df2 == -Inf] <- 0
      colnames(df2)[1] <- "Step"
      # df <- df %>%  rename(Step = rawdata)



     

      reduced <- df2 %>% mutate(`Sum in millions of hooks` = `Sum in hooks` / 1000000, `Sum in millions of days` = `Sum in fishing days` / 1000000)%>% dplyr::select(Step, `Sum in millions of hooks`, `Sum in millions of days`, #`Lines`, 
                                                                                                                                                              "Difference (in % of hooks)","Difference (in % of fishing days)")%>% dplyr::mutate(`Step number` = as.numeric(row_number()))
      reduced$Step <- factor(reduced$Step, levels = (reduced %>% arrange(`Step number`))$Step)
      
      reduced <- ungroup()
      

coeff <- 3
temperatureColor <- "#69b3a2"
      
priceColor <- rgb(0.2, 0.6, 0.9, 1)

second_graf <- ggplot(reduced,aes(x = Step,group = 1))+ 
        geom_line( aes(y=`Sum in millions of hooks` ,group = 1), size=0.5, color=priceColor)+geom_point( aes(y=`Sum in millions of hooks` ,group = 1)) +
        geom_line( aes(y=`Sum in millions of days`/ coeff,group = 1), size=0.5, color=temperatureColor)+geom_point( aes(y=`Sum in millions of days`/ coeff))  +
        scale_y_continuous(
          
          # Features of the first axis
          name = "hooks",
          
          # Add a second axis and specify its features
          sec.axis = sec_axis(~.*coeff, name="fishing days")
        ) + 
        
        
        
        theme(
          axis.title.y = element_text(color = priceColor, size=8),
          axis.title.y.right = element_text(color = temperatureColor, size=8)
        ) +
        
        ggtitle("Evolution of the repartition of captures depending on units and Steps")+
        theme(axis.text.x = element_text(angle = 90))

fishing_days_plot <- ggplot(reduced,aes(x = Step,group = 1)) +geom_line( aes(y=`Sum in millions of days`,group = 1), size=0.5)+theme(axis.text.x = element_text(angle = 90))
      
hooks_plot <- ggplot(reduced,aes(x = Step,group = 1)) +
geom_line( aes(y=`Sum in millions of hooks`,group = 1), size=0.5)+theme(axis.text.x = element_text(angle = 90))

```

```{r echo=FALSE,  results='asis', eval=(parameter_fact == "catch")}

qflextable2(reduced, captionn ="Evolution of Tons and Number of fish and lines along the steps", pgwidth = 6, save_folder = "Summary", columns_to_color = c("Difference (in % of tons)","Difference (in % of fish)"), fig.pathinside = fig.path )

```

```{r echo=FALSE, results='asis', eval=(parameter_fact == "effort")}

qflextable2(reduced, captionn ="Evolution of Sum in hooks and Sum in fishing days and lines along the steps",  pgwidth = 6, save_folder = "Summary", fig.pathinside = fig.path )

```


```{r evolcapturesduringprocess,  echo=FALSE,results='asis', eval=(parameter_fact == "catch"), fig.cap = 'Evolution of captures in tons and number of fish during the process'}

plottonsfish <- cowplot::plot_grid(tons_plot, no_fish_plot)

save_image(title= knitr::opts_current$get("fig.cap"),plott = plottonsfish, folder = "Summary", fig.pathinside = fig.path)

```

```{r evolcapturesduringprocesseffort,  echo=FALSE,results='asis', eval=(parameter_fact == "effort"), fig.cap = 'Evolution of captures in fishing days and hooks during the process'}

plotfishingdayshooks <- cowplot::plot_grid(fishing_days_plot, hooks_plot)

save_image(title= knitr::opts_current$get("fig.cap"),plott = plotfishingdayshooks, folder = "Summary", fig.pathinside = fig.path)

```

\clearpage


# Detail of all the steps of treatment of the data

For each treatment on the data, we provide a summary of what has been added, removed or transformed. This allows to see where major changes occurs for specific data needs. 

```{r functionallcomp, echo=FALSE, results='hide'}

function_all_comp = function(counting, parameters = parameters_child, parameter_short = FALSE, sub_list_dir = sub_list_dir_2, child_env_baseinside =child_env_base){
  gc()
step_mapping <- sum(which(sub_list_dir=="Markdown/mapping_codelist"))

parameter_mapped <- ifelse(counting!=step_mapping, TRUE, FALSE)
child_env <- list2env(as.list(child_env_base), parent = child_env_base)
# source(knitr::purl("Functions_markdown.Rmd", quiet=TRUE),child_env)

child_env$parameter_init <- paste0(sub_list_dir[counting], "/rds.rds")
child_env$parameter_final <- paste0(sub_list_dir[counting+1], "/rds.rds")
child_env$parameter_titre_dataset_1 <- last_path(sub_list_dir[counting])
child_env$parameter_titre_dataset_2 <- last_path(sub_list_dir[counting+1])

child_env$child <- TRUE
child_env$treatment <- TRUE
child_env$parameter_mapped <- parameter_mapped
child_env$parameter_short <- parameter_short
child_env$step <- counting
child_env$step_title <- paste0(" Treatment : " ,last_path(sub_list_dir[counting+1]))
child_env$child_header <- "##"
child_env$unique_analyse <- FALSE

child_env$debug <- TRUE

new_path <- file.path(fig.path, "/Comparison/",gsub("Markdown/","",sub_list_dir[counting]),"_", gsub("Markdown/","",sub_list_dir[counting+1]))

dir.create(new_path, recursive = TRUE)

child_env$fig.path = new_path

if( identical(readRDS(child_env$parameter_init),readRDS(child_env$parameter_final) )){NA}else{
      # knitr::knit_child(text =c('```{r child="comparison.Rmd"}',
      #                       '',
      #                       '',
      #                       '',
      #                       '```'), envir = child_env, quiet= TRUE)
      res <- knitr::knit_child("comparison.Rmd", envir = child_env, quiet = TRUE)
        rm(child_env)
        return(res)
        gc()



    

    
}
}


final_step <- (length(sub_list_dir_2)-1)
# final_step <- 1
all <- lapply(1:final_step, function_all_comp)
all <- all[!is.na(all)]
```

```{r unlistallcomp,echo=FALSE, results='asis'}

cat(unlist(all), sep =  "\\clearpage")

```


\clearpage

# Annexe

```{r reviewimpact}
colnames(df2)[1] <-  "Treatment"

df2 <- df2 %>% dplyr::select(c("Treatment", "Explenation", "Functions" = "Fonctions"))%>% rowwise()%>% mutate_all(function(.)toString(.)) %>% mutate_all(function(x)gsub("_", "-", x)) %>%  dplyr::mutate_if(is.character,~str_replace_all( ., "_", "-" )) %>%  dplyr::mutate_if(is.character,~str_replace_all( ., "\n", " " )) %>%  dplyr::mutate_if(is.character,~str_replace_all( ., "  ", " " ))%>%mutate_if(is.character, toString)

df3 <- flextable(df2) %>% width("Explenation",width = 4)
df3 <- set_caption(df3, caption ="Review of all the impact and purpose of every the treatment done flextable")

```


```{r reviewimpactknitting, results='asis'}

knit_print(df3)

```


